### 2022-03-28 paper/paper.Rmd
initial commit

---
title: "Interactive dashboard to monitor the COVID-19 outbreak"
type: 
author:
  - name: Thiyanga S. Talagala
    affil: a
    email: ttalagala@sjp.ac.lk
  - name: Randi Shashikala
    affil: a
affiliation:
  - num: a
    address: |
      Department of Statistics, University of Sri Jayeardenepura
bibliography: interactcadsample.bib
appendix: appendix.tex
abstract: |
  As of September 20th, 2021, 221 countries and territories are infected by the COVID-19 worldwide pandemic. Dashboards are the most often used visualization method for visualizing COVID-19 data and informing the public. The main objective of this study is to identify different features, visualization methods & improvements that should be occurred by exploring the existing dashboards and to develop a dashbored to visualize COVID-19 outbreak in Sri Lanka. We explored 15 different dashboards. The most commonly used visualization methods in dashboard development are bar charts, line charts, and interactive maps. Dashboards that fit on a single screen are preferable than others.
keywords: |
  COVID-19, Dashboard, Visualization
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
output: rticles::tf_article
---

# Introduction

The COVID-19 pandemic is a global coronavirus illness outbreak that began in 2019 and is being caused by the Severe Acute Respiratory Syndrome Coronavirus 2 (SARS-CoV-2) virus. In December 2019, the first COVID-19-infected patient was discovered in Wuhan, China. According to Worldmeter data, there were 229835231 confirmed cases, 206515718 recovered cases, 18605877 active cases, and 4713636 deaths worldwide as of September 20, 2021. Currently, 99.5 percent of active patients are in a mild state, while 0.5 percent are in a critical state.

Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemic's spread and communication. There are a plethora of COVID-19 visualization dashboards that have been designed to visualize the pandemic's global and local status. Different software can be used to generate dashboards. We explored 15 dashboards designed to visualize COVID-19 data in the global and country levels. First, dashboards were compared to identify the various features, visualization approaches, and enhancements that should be implemented. Next, we developed an interactive dashboard to visualized COVID-19 outbreak in Sri Lanka.

# Literature Review

# Methodology

# Results

# Conclusions

### 2022-03-28 paper/paper.tex
initial commit

--- 2022-03-28 paper/paper.Rmd
@@
----
-title: "Interactive dashboard to monitor the COVID-19 outbreak"
-type: 
-author:
-  - name: Thiyanga S. Talagala
-    affil: a
-    email: ttalagala@sjp.ac.lk
-  - name: Randi Shashikala
-    affil: a
-affiliation:
-  - num: a
-    address: |
-      Department of Statistics, University of Sri Jayeardenepura
-bibliography: interactcadsample.bib
-appendix: appendix.tex
-abstract: |
-  As of September 20th, 2021, 221 countries and territories are infected by the COVID-19 worldwide pandemic.
-Dashboards are the most often used visualization method for visualizing COVID-19 data and informing the public.
-The main objective of this study is to identify different features, visualization methods & improvements that should be occurred by exploring the existing dashboards and to develop a dashbored to visualize COVID-19 outbreak in Sri Lanka.
-We explored 15 different dashboards.
-The most commonly used visualization methods in dashboard development are bar charts, line charts, and interactive maps.
-Dashboards that fit on a single screen are preferable than others.
-keywords: |
-  COVID-19, Dashboard, Visualization
-header-includes: |
+% interactcadsample.tex
+% v1.03 - April 2017
+
+\documentclass[]{interact}
+
+\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
+\usepackage{subfigure}% Support for small, `sub' figures and tables
+%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
+
+\usepackage{natbib}% Citation support using natbib.sty
+\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
+\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty
+
+\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{proposition}[theorem]{Proposition}
+
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+
+\theoremstyle{remark}
+\newtheorem{remark}{Remark}
+\newtheorem{notation}{Notation}
+
+% see https://stackoverflow.com/a/47122900
+
+% Pandoc citation processing
+
 \usepackage{hyperref}
@@
 \def\tightlist{}
-output: rticles::tf_article
----
 
-# Introduction
+\begin{document}
 
-The COVID-19 pandemic is a global coronavirus illness outbreak that began in 2019 and is being caused by the Severe Acute Respiratory Syndrome Coronavirus 2 (SARS-CoV-2) virus.
-In December 2019, the first COVID-19-infected patient was discovered in Wuhan, China.
-According to Worldmeter data, there were 229835231 confirmed cases, 206515718 recovered cases, 18605877 active cases, and 4713636 deaths worldwide as of September 20, 2021.
-Currently, 99.5 percent of active patients are in a mild state, while 0.5 percent are in a critical state.
+\articletype{}
 
-Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemic's spread and communication.
-There are a plethora of COVID-19 visualization dashboards that have been designed to visualize the pandemic's global and local status.
-Different software can be used to generate dashboards.
-We explored 15 dashboards designed to visualize COVID-19 data in the global and country levels.
-First, dashboards were compared to identify the various features, visualization approaches, and enhancements that should be implemented.
-Next, we developed an interactive dashboard to visualized COVID-19 outbreak in Sri Lanka.
+\title{Interactive dashboard to monitor the COVID-19 outbreak}
 
-# Literature Review
+\author{\name{Thiyanga S. Talagala$^{a}$, Randi Shashikala$^{a}$}
+\affil{$^{a}$Department of Statistics, University of Sri Jayeardenepura}
+}
 
-# Methodology
+\thanks{CONTACT Thiyanga S.
+Talagala. Email: \href{mailto:ttalagala@sjp.ac.lk}{\nolinkurl{ttalagala@sjp.ac.lk}}, Randi
+Shashikala. Email: }
 
-# Results
+\maketitle
 
-# Conclusions
+\begin{abstract}
+As of September 20th, 2021, 221 countries and territories are infected
+by the COVID-19 worldwide pandemic. Dashboards are the most often used
+visualization method for visualizing COVID-19 data and informing the
+public. The main objective of this study is to identify different
+features, visualization methods \& improvements that should be occurred
+by exploring the existing dashboards and to develop a dashbored to
+visualize COVID-19 outbreak in Sri Lanka. We explored 15 different
+dashboards. The most commonly used visualization methods in dashboard
+development are bar charts, line charts, and interactive maps.
+Dashboards that fit on a single screen are preferable than others.
+\end{abstract}
+
+\begin{keywords}
+COVID-19, Dashboard, Visualization
+\end{keywords}
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+The COVID-19 pandemic is a global coronavirus illness outbreak that
+began in 2019 and is being caused by the Severe Acute Respiratory
+Syndrome Coronavirus 2 (SARS-CoV-2) virus. In December 2019, the first
+COVID-19-infected patient was discovered in Wuhan, China. According to
+Worldmeter data, there were 229835231 confirmed cases, 206515718
+recovered cases, 18605877 active cases, and 4713636 deaths worldwide as
+of September 20, 2021. Currently, 99.5 percent of active patients are in
+a mild state, while 0.5 percent are in a critical state.
+
+Dashboards are one of the greatest visual interpretation methods for
+tracking the COVID-19 pandemic's spread and communication. There are a
+plethora of COVID-19 visualization dashboards that have been designed to
+visualize the pandemic's global and local status. Different software can
+be used to generate dashboards. We explored 15 dashboards designed to
+visualize COVID-19 data in the global and country levels. First,
+dashboards were compared to identify the various features, visualization
+approaches, and enhancements that should be implemented. Next, we
+developed an interactive dashboard to visualized COVID-19 outbreak in
+Sri Lanka.
+
+\hypertarget{literature-review}{%
+\section{Literature Review}\label{literature-review}}
+
+\hypertarget{methodology}{%
+\section{Methodology}\label{methodology}}
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\bibliographystyle{tfcad}
+\bibliography{interactcadsample.bib}
+
+\input{"appendix.tex"}
+
+\end{document}

### 2022-04-20 manuscript_covid19dashboard.Rmd
Updated the paper

---
title: "Interactive dashboard to monitor the COVID-19 outbreak"
output: pdf_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Thiyanga S. Talagala**^[Corresponding author, email: ttalagala@gmail.com]

Department of Statistics

Faculty of Applied Sciences

University of Sri Jayewardenepura, Sri Lanka, CO10230

\hspace{3cm}

**Randi Shashikala**

Department of Statistics, Faculty of Applied Sciences

University of Sri Jayewardenepura, Sri Lanka, CO 10230

# Abstract

As of September 20th, 2021, 221 countries and territories are infected by the COVID-19 worldwide pandemic. Dashboards are the most often used visualization method for visualizing COVID-19 data and informing the public. The main objective of this study is to identify different features, visualization methods & improvements that should be occurred by exploring the existing dashboards and to develop a dashbored to visualize COVID-19 outbreak in Sri Lanka. We explored 15 different dashboards. The most commonly used visualization methods in dashboard development are bar charts, line charts, and interactive maps. Dashboards that fit on a single screen are preferable than others.

**Keywords:** Visualization, multiple time series, heatmap, COVID-19 vaccine, flexdashboard

# Introduction

The COVID-19 pandemic is a global coronavirus illness outbreak that began in 2019 and is being caused by the Severe Acute Respiratory Syndrome Coronavirus 2 (SARS-CoV-2) virus. In December 2019, the first COVID-19-infected patient was discovered in Wuhan, China. According to Worldmeter data, there were 229835231 confirmed cases, 206515718 recovered cases, 18605877 active cases, and 4713636 deaths worldwide as of September 20, 2021. Currently, 99.5 percent of active patients are in a mild state, while 0.5 percent are in a critical state.

Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemic's spread and communication. There are a plethora of COVID-19 visualization dashboards that have been designed to visualize the pandemic's global and local status. Different software can be used to generate dashboards. We explored 15 dashboards designed to visualize COVID-19 data in the global and country levels. First, dashboards were compared to identify the various features, visualization approaches, and enhancements that should be implemented. Next, we developed an interactive dashboard to visualized COVID-19 outbreak in Sri Lanka.

# Literature Review

The 15 dashboards we used in the study are listed below (Table 1).

# Methodology

\newpage

# Results

The "Sri Lanka COVID-19 Dashboard" provides an overview of the COVID-19 pandemic for Sri Lanka. This dashboard has eight panels as follows.

**Table 01: Description of the panels**

| **Name of the Panel**|**Description of the Panel**| 
|:------|:----------------------------------------------|
|Overview          |Total count of COVID-19 confirmed, recovered, deaths, active cases & total vaccine doses administered.\hfill\break Provide an overview of daily COVID-19 confirmed, recovered & deaths by plots.\hfill\break|
|Cases by Wave| There are three tabs in this panel.\hfill\break  * Total COVID-19 confirmed cases - Cumulative count of COVID-19 confirmed cases with specific dates\hfill\break * COVID-19 Cases Distribution by Wave - Daily confirmed cases by wave\hfill\break * COVID-19 Deaths Distribution by Wave - Daily deaths by wave\hfill\break| 
|COVID-19 Patients Distribution | Provide an overview of confirmed cases district wise distribution. There are four tabs in this panel.\hfill\break * Total COVID-19 Patients Distribution in Sri Lanka - Total confirmed counts for each district is represented by tree map\hfill\break * Country Map - Total confirmed cases in each district represented by Sri Lanka country map\hfill\break * Distribution of Daily COVID-19 Patients for Last 30 Days - Visualize the daily confirmed cases distribution by districts in last 30 days\hfill\break * By Applying Min-Max Transformation - Visualize the details in the third tab by applying min-max transformation for each district\hfill\break|
|Vaccination Details | Provide an overview of COVID-19 vaccination in Sri Lanka. There are two tabs.\hfill\break * Total Vaccine Doses - Visualize the administered vaccine doses as first dose only & fully vaccinated\hfill\break * Total Administered Doses by Vaccine Name - Visualize the vaccination by vaccine names\hfill\break|
| Top 10 Countries| In this panel, compare the Sri Lanka confirmed & deaths with top 10 countries in the world (top 10 countries - The countries which have been reported highest number of confirmed cases as 31st of August 2021).\hfill\break There are two tabs.\hfill\break * Comparison of Cumulative Cases in Sri Lanka with Top 10 Countries - Compare the confirmed and deaths in Sri Lanka with top 10 countries by cumulative time series plots\hfill\break * Comparison of Log of Cumulative Cases in Sri Lanka with Top 10 Countries -  Compare the confirmed and deaths in Sri Lanka with top 10 countries by log cumulative time series plots\hfill\break (The data has been pulled from WHO)\hfill\break|
|Global Comparison | Compare the total confirmed & deaths in Sri lanka with top 10 countries in Global & Asia. There are two tabs.\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the World - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the world & compare the case fatality ratios\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the Asia - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the Asia & compare the case fatality ratios\hfill\break|
|Global Map| Visualize the distribution of confirmed, recovered & deaths in the world by world map.\hfill\break|
|About | This panel contains the details about the dashboard.|

# Discussion and Further research

# References

### 2022-04-20 manuscript_covid19dashboard.tex
Updated the paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Interactive dashboard to monitor the COVID-19 outbreak},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Interactive dashboard to monitor the COVID-19 outbreak}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\textbf{Thiyanga S. Talagala}\footnote{Corresponding author, email:
  \href{mailto:ttalagala@gmail.com}{\nolinkurl{ttalagala@gmail.com}}}

Department of Statistics

Faculty of Applied Sciences

University of Sri Jayewardenepura, Sri Lanka, CO10230

\hspace{3cm}

\textbf{Randi Shashikala}

Department of Statistics, Faculty of Applied Sciences

University of Sri Jayewardenepura, Sri Lanka, CO 10230

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

As of September 20th, 2021, 221 countries and territories are infected
by the COVID-19 worldwide pandemic. Dashboards are the most often used
visualization method for visualizing COVID-19 data and informing the
public. The main objective of this study is to identify different
features, visualization methods \& improvements that should be occurred
by exploring the existing dashboards and to develop a dashbored to
visualize COVID-19 outbreak in Sri Lanka. We explored 15 different
dashboards. The most commonly used visualization methods in dashboard
development are bar charts, line charts, and interactive maps.
Dashboards that fit on a single screen are preferable than others.

\textbf{Keywords:} Visualization, multiple time series, heatmap,
COVID-19 vaccine, flexdashboard

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The COVID-19 pandemic is a global coronavirus illness outbreak that
began in 2019 and is being caused by the Severe Acute Respiratory
Syndrome Coronavirus 2 (SARS-CoV-2) virus. In December 2019, the first
COVID-19-infected patient was discovered in Wuhan, China. According to
Worldmeter data, there were 229835231 confirmed cases, 206515718
recovered cases, 18605877 active cases, and 4713636 deaths worldwide as
of September 20, 2021. Currently, 99.5 percent of active patients are in
a mild state, while 0.5 percent are in a critical state.

Dashboards are one of the greatest visual interpretation methods for
tracking the COVID-19 pandemic's spread and communication. There are a
plethora of COVID-19 visualization dashboards that have been designed to
visualize the pandemic's global and local status. Different software can
be used to generate dashboards. We explored 15 dashboards designed to
visualize COVID-19 data in the global and country levels. First,
dashboards were compared to identify the various features, visualization
approaches, and enhancements that should be implemented. Next, we
developed an interactive dashboard to visualized COVID-19 outbreak in
Sri Lanka.

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

The 15 dashboards we used in the study are listed below (Table 1).

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

\newpage

\hypertarget{results}{%
\section{Results}\label{results}}

The ``Sri Lanka COVID-19 Dashboard'' provides an overview of the
COVID-19 pandemic for Sri Lanka. This dashboard has eight panels as
follows.

\textbf{Table 01: Description of the panels}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.13}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.87}}@{}}
\toprule
\textbf{Name of the Panel} & \textbf{Description of the Panel} \\
\midrule
\endhead
Overview & Total count of COVID-19 confirmed, recovered, deaths, active
cases \& total vaccine doses administered.\hfill\break Provide an
overview of daily COVID-19 confirmed, recovered \& deaths by
plots.\hfill\break \\
Cases by Wave & There are three tabs in this panel.\hfill\break  * Total
COVID-19 confirmed cases - Cumulative count of COVID-19 confirmed cases
with specific dates\hfill\break * COVID-19 Cases Distribution by Wave -
Daily confirmed cases by wave\hfill\break * COVID-19 Deaths Distribution
by Wave - Daily deaths by wave\hfill\break \\
COVID-19 Patients Distribution & Provide an overview of confirmed cases
district wise distribution. There are four tabs in this
panel.\hfill\break * Total COVID-19 Patients Distribution in Sri Lanka -
Total confirmed counts for each district is represented by tree
map\hfill\break * Country Map - Total confirmed cases in each district
represented by Sri Lanka country map\hfill\break * Distribution of Daily
COVID-19 Patients for Last 30 Days - Visualize the daily confirmed cases
distribution by districts in last 30 days\hfill\break * By Applying
Min-Max Transformation - Visualize the details in the third tab by
applying min-max transformation for each district\hfill\break \\
Vaccination Details & Provide an overview of COVID-19 vaccination in Sri
Lanka. There are two tabs.\hfill\break * Total Vaccine Doses - Visualize
the administered vaccine doses as first dose only \& fully
vaccinated\hfill\break * Total Administered Doses by Vaccine Name -
Visualize the vaccination by vaccine names\hfill\break \\
Top 10 Countries & In this panel, compare the Sri Lanka confirmed \&
deaths with top 10 countries in the world (top 10 countries - The
countries which have been reported highest number of confirmed cases as
31st of August 2021).\hfill\break There are two
tabs.\hfill\break * Comparison of Cumulative Cases in Sri Lanka with Top
10 Countries - Compare the confirmed and deaths in Sri Lanka with top 10
countries by cumulative time series plots\hfill\break * Comparison of
Log of Cumulative Cases in Sri Lanka with Top 10 Countries - Compare the
confirmed and deaths in Sri Lanka with top 10 countries by log
cumulative time series plots\hfill\break (The data has been pulled from
WHO)\hfill\break \\
Global Comparison & Compare the total confirmed \& deaths in Sri lanka
with top 10 countries in Global \& Asia. There are two
tabs.\hfill\break * Comparison of the sri Lanka with Top 10 Countries
Reporting the Most COVID-19 Cases in the World - Compare the total
confirmed \& deaths in Sri Lanka with top 10 countries in the world \&
compare the case fatality ratios\hfill\break * Comparison of the sri
Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the
Asia - Compare the total confirmed \& deaths in Sri Lanka with top 10
countries in the Asia \& compare the case fatality ratios\hfill\break \\
Global Map & Visualize the distribution of confirmed, recovered \&
deaths in the world by world map.\hfill\break \\
About & This panel contains the details about the dashboard. \\
\bottomrule
\end{longtable}

\hypertarget{discussion-and-further-research}{%
\section{Discussion and Further
research}\label{discussion-and-further-research}}

\hypertarget{references}{%
\section{References}\label{references}}

\end{document}

### 2022-04-20 paper/Panel Description.Rmd
Updated the paper

--- 2022-04-20 manuscript_covid19dashboard.Rmd
@@
 ---
-title: "Interactive dashboard to monitor the COVID-19 outbreak"
+title: "Dashboard Panel Description"
+author: "Randi Shashikala"
 output: pdf_document
-bibliography: references.bib
 ---
@@
 ```
-
-**Thiyanga S. Talagala**^[Corresponding author, email: ttalagala@gmail.com]
-
-Department of Statistics
-
-Faculty of Applied Sciences
-
-University of Sri Jayewardenepura, Sri Lanka, CO10230
-
-\hspace{3cm}
-
-**Randi Shashikala**
-
-Department of Statistics, Faculty of Applied Sciences
-
-University of Sri Jayewardenepura, Sri Lanka, CO 10230
-
-# Abstract
-
-As of September 20th, 2021, 221 countries and territories are infected by the COVID-19 worldwide pandemic.
-Dashboards are the most often used visualization method for visualizing COVID-19 data and informing the public.
-The main objective of this study is to identify different features, visualization methods & improvements that should be occurred by exploring the existing dashboards and to develop a dashbored to visualize COVID-19 outbreak in Sri Lanka.
-We explored 15 different dashboards.
-The most commonly used visualization methods in dashboard development are bar charts, line charts, and interactive maps.
-Dashboards that fit on a single screen are preferable than others.
-
-**Keywords:** Visualization, multiple time series, heatmap, COVID-19 vaccine, flexdashboard
-
-# Introduction
-
-The COVID-19 pandemic is a global coronavirus illness outbreak that began in 2019 and is being caused by the Severe Acute Respiratory Syndrome Coronavirus 2 (SARS-CoV-2) virus.
-In December 2019, the first COVID-19-infected patient was discovered in Wuhan, China.
-According to Worldmeter data, there were 229835231 confirmed cases, 206515718 recovered cases, 18605877 active cases, and 4713636 deaths worldwide as of September 20, 2021.
-Currently, 99.5 percent of active patients are in a mild state, while 0.5 percent are in a critical state.
-
-Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemic's spread and communication.
-There are a plethora of COVID-19 visualization dashboards that have been designed to visualize the pandemic's global and local status.
-Different software can be used to generate dashboards.
-We explored 15 dashboards designed to visualize COVID-19 data in the global and country levels.
-First, dashboards were compared to identify the various features, visualization approaches, and enhancements that should be implemented.
-Next, we developed an interactive dashboard to visualized COVID-19 outbreak in Sri Lanka.
-
-# Literature Review
-
-The 15 dashboards we used in the study are listed below (Table 1).
-
-# Methodology
-
-\newpage
-
-# Results
 
@@
 |About | This panel contains the details about the dashboard.|
-
-# Discussion and Further research
-
-# References

### 2022-04-20 manuscript_covid19dashboard.Rmd
added new manuscript format
Updated paper

@@
 title: "Interactive dashboard to monitor the COVID-19 outbreak"
-output: pdf_document
+output: 
+  pdf_document: 
+    keep_tex: true
 bibliography: references.bib
@@
 
-The 15 dashboards we used in the study are listed below (Table 1).
+Initially 15 dashboards were compared  to determine which data types are most frequently shown in dashboards. The 15 dashboards we used in the study are listed below (Table 1).
+
+**Table 01: Labeled of the Dashboards**
+
+|No |Name of the Dashboard|Reference|
+|:--:|:-------------------|------|
+|1 | COVID-19 dashboard created by the John Hopkins University Center for Systems Science & Engineering (JHU CSSE) ||
+|2 | WHO  COVID-19 Dashboard ||
+|3 | COVID-19 surveillance dashboard created by the University of Virginia ||
+|4 | Corona cases (COVID-19) per municipality in Belgium dashboard |@belgium|
+|5 | COVID-19 dashboard for England created by NHS providers |@nhs|
+|6 | NZ COVID-19 Dashboard ||
+|7 | Pakistan’s official COVID-19 dashboard ||
+|8 | COVID-19 Canada live dashboard ||
+|9 | India (COVID-19) Dashboard | @indiadash |
+|10 | Italy COVID-19 dashboard |@italy|
+|11 | Jamaica COVID-19 Dashboard ||
+|12 | GCI COVID-19 dashboard for Russia ||
+|13 | COVID-19 live situation analysis dashboard of Sri Lanka |@sl|
+|14 | COVID 19 ZA South Africa Dashboard ||
+|15 | COVID-19 dashboard for Germany ||
 

### 2022-04-20 manuscript_covid19dashboard.tex
Updated paper

@@
 \fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-The 15 dashboards we used in the study are listed below (Table 1).
+Initially 15 dashboards were compared to determine which data types are
+most frequently shown in dashboards. The 15 dashboards we used in the
+study are listed below (Table 1).
+
+\textbf{Table 01: Labeled of the Dashboards}
+
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.13}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.67}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.20}}@{}}
+\toprule
+No & Name of the Dashboard & Reference \\
+\midrule
+\endhead
+1 & COVID-19 dashboard created by the John Hopkins University Center for
+Systems Science \& Engineering (JHU CSSE) & \\
+2 & WHO COVID-19 Dashboard & \\
+3 & COVID-19 surveillance dashboard created by the University of
+Virginia & \\
+4 & Corona cases (COVID-19) per municipality in Belgium dashboard &
+{``{COVID 19 Dashboard -- Belgium}''} (2022) \\
+5 & COVID-19 dashboard for England created by NHS providers & {``{COVID
+19 Dashboard -- NHS providers}''} (2022) \\
+6 & NZ COVID-19 Dashboard & \\
+7 & Pakistan's official COVID-19 dashboard & \\
+8 & COVID-19 Canada live dashboard & \\
+9 & India (COVID-19) Dashboard & {``{COVID 19 Dashboard India -- ZOHO
+Analytics -- ZOHO}''} (2022) \\
+10 & Italy COVID-19 dashboard & {``{COVID-19 integrated surveillance
+data in Italy -- EpiCentro}''} (2022) \\
+11 & Jamaica COVID-19 Dashboard & \\
+12 & GCI COVID-19 dashboard for Russia & \\
+13 & COVID-19 live situation analysis dashboard of Sri Lanka &
+{``{COVID-19: Live situational Analysis Dashboard of Sri Lanka,}''}
+(2022) \\
+14 & COVID 19 ZA South Africa Dashboard & \\
+15 & COVID-19 dashboard for Germany & \\
+\bottomrule
+\end{longtable}
 
@@
 \hypertarget{references}{%
-\section{References}\label{references}}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-belgium}{}%
+{``{COVID 19 Dashboard -- Belgium}.''} 2022.
+\url{https://esribelux.maps.arcgis.com/apps/ops\%0Adashboard/index.html\#/e350724c87af49bb9ce29646f8a42344}.
+
+\leavevmode\hypertarget{ref-indiadash}{}%
+{``{COVID 19 Dashboard India -- ZOHO Analytics -- ZOHO}.''} 2022.
+\url{https://www.zoho.com/covid/india/}.
+
+\leavevmode\hypertarget{ref-nhs}{}%
+{``{COVID 19 Dashboard -- NHS providers}.''} 2022.
+\url{https://nhsproviders.org/topics/covid-\%0A19/tracking-covid-19-data/covid-19-dashboard}.
+
+\leavevmode\hypertarget{ref-italy}{}%
+{``{COVID-19 integrated surveillance data in Italy -- EpiCentro}.''}
+2022.
+\url{https://www.ep\%0Aicentro.iss.it/en/coronavirus/sars-cov-2-dashboard}.
+
+\leavevmode\hypertarget{ref-sl}{}%
+{``{COVID-19: Live situational Analysis Dashboard of Sri Lanka,}''}
+2022. \url{https:\%0A//hpb.health.gov.lk/covid19-dashboard/}.
+
+\end{CSLReferences}
 

### 2022-04-20 paper/Review paper/Literature Survey.Rmd
Updated paper

---
title: "Review of COVID-19 Dashboards"
author: "Randi Shashikala"
date: "9/20/2021"
output: 
  html_document:
  toc: TRUE
---
<style>
body{
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

There are 221 countries & territories which are affected by the COVID-19 global pandemic as 20th of September 2021.To visualize the COVID-19 data & inform the general public, dashboards are the most commonly used visualization tool. The main objective of this study is to identify different features, visualization methods & improvements that should be occurred by exploring the existing dashboards. 15 different dashboards that have been developed to visualize the COVID-19 data for different countries & global has been explored here. Bar charts, line charts & Interactive maps are the most frequently used visualization tools in developing the dashboards. Dashboards that are fitted on a single screen is better than others.

<b>Key words</b> : COVID-19, Dashboard, Visualization

# Introduction

The COVID-19 pandemic is the ongoing global pandemic of coronavirus disease 2019
caused by the Severe Acute Respiratory Syndrome Coronavirus 2 (SARS-CoV-2) virus. The first
patient infected with COVID-19 was identified in Wuhan, China, in December 2019. There have been reported 229835231 confirmed cases, 206515718 recovered cases, 18605877 active cases & 4713636 deaths in global as 20th September 2021. Currently 99.5% of active patients are mild condition and 0.5% of patients are in critical situation.

Dashboards are one of the best visual interpretation method of tracking the spread & communication of COVID-19 pandemic. There are so many COVID-19 visualization dashboards that have been created to visualize the global & local situation of pandemic. Real time updated COVID-19 dashboard created by the John Hopkins University Center for Systems Science & Engineering (JHU CSSE) has included total confirmed cases & total deaths by country, province/region/state (not for all countries)[6]. Weekly & daily global confirmed cases, deaths & vaccine doses have been visualized using bar plots related to the date in the side panels. Also, an interactive world map has included in the dashboard to represent total confirmed cases, incidence rate, fatality ratio & vaccine doses administered by country & state of US. All the links related to the data sets has been included in the dashboard. In this dashboard red, white, green colors has been used to indicate confirmed cases, deaths & vaccine doses respectively. The dashboard has fitted on a single screen.

<center>
![](1.PNG) <br/>
**Figure 01: Johns Hopkins University COVID-19 Dashboard**
</center>
<br/>

Live updated “WHO COVID-19 dashboard” contained four panels. The interactive world
map in the dashboard has visualized the distribution of COVID-19 confirmed cases, deaths &
vaccination by country [15]. Line chart & bar charts have been used as visualization plots. Blue, red, green colors have been used to represent confirmed cases, deaths & vaccination respectively. The links related to the underlying data was available in the dashboard. The dashboard was not fitted on a single screen.

<center>
![](2.PNG) <br/>
**Figure 02: WHO COVID-19 Dashboard**
</center>
<br/>

The “COVID-19 surveillance dashboard” created by the University of Virginia in
collaboration with Bio complexity Institute has included interactive world map with a time slide to visualize the confirmed cases, recovered cases, deaths, active cases & vaccine details [7]. Line charts, bar charts & data tables have been visualized in the dashboard. Red, blue, green & yellow colors have been used to identify confirmed, deceased, recovered & active cases respectively. The data behind the dashboard also available on the dashboard & can be downloaded. The dashboard has fitted on a single screen.

<center>
![](3.png)<br/>
**Figure 03: COVID-19 Surveillance Dashboard**
</center>
<br/>

Corona cases (COVID-19) per municipality in Belgium dashboard has included interactive
map for distribution of total confirmed cases per municipality, bar chart for number of cases per province, line charts for number of cases per day by region & municipality, dodge bar chart for deaths per region, per age & pie chart for deaths per age as visualization methods[3]. Also it included several value boxes for total cases, total deaths & hospital situations. There was no specific color theme used in this dashboard. This dashboard also fitted on a single screen.

<center>
![](4.png)<br/>
**Figure 04: Corona cases per Municipality in Belgium Dashboard**
</center>
<br/>

The COVID-19 dashboard for England created by NHS providers included daily totals, daily changes & weekly changes in new cases, hospital admission, patients in hospital, and patients
in ICU beds, deaths & vaccinations in England as counts [2]. Bar plots for daily new cases, hospital admissions & deaths for last 30 days have been added in this dashboard. Also it has included data table for total count, daily changes (as increase or decrease), weekly changes (as increase or decrease) of new cases & hospital admission by region. It can be seen that, three different colors have been used in bar plot as orange for new cases, green for hospital admissions & red for deaths. These plots were not interactive plots & there was only one panel. The dashboard was not fitted on a single screen.

<center>
![](5.png)<br/>
**Figure 05: COVID-19 dashboard for England**
</center>
<br/>

The Environmental Science and Research Institute COVID-19 dashboard for New Zealand (was developed & is maintained in collaboration with Epi-interactive Ltd) contained five panels
as overview, outbreak, source, international & ESR reports [10]. In this real time updated dashboard, they have included value boxes for total confirmed cases, recovered cases, deaths & interactive map has been added for distribution of confirmed & probable cases as incidence & count. Also they have included bar charts for confirmed cases distribution by gender, age group, & line charts for confirmed & deaths as daily & cumulative. The dashboard has not fitted on a single screen.

<center>
![](6.png)<br/>
**Figure 06: NZ COVID-19 Dashboard**
</center>
<br/>

Pakistan’s official COVID-19 dashboard is a real time updated dashboard with one panel. This dashboard contains value boxes for total confirmed cases, active cases, deaths, recovered
cases by country. Province wise cases have been represented by map & data table [11]. Line charts & bar charts have been added to visualize the distribution of cases related to time & province. In the value boxes blue, orange, pink & green colors have been applied for confirmed cases, active cases, deaths & recoveries respectively, but there was no specific color theme used when plotting the graphs. The dashboard cannot be seen on a single screen without grid overlay.

<center>
![](7.png)<br/>
**Figure 07: COVID-19 Dashboard for Pakistan**
</center>
<br/>

COVID-19 Canada live dashboard has been developed with three panels for provincial data, hospital resources & projections [14]. This dashboard contained interactive map for represented total reported cases & deaths by province. Data tables, line charts, bar charts were the most frequently used visualization types of this dashboard. More details were presented on this dashboard by province like ventilator counts, hospital beds per 100 residents, resource capacity based on critical case rate, effect of social distance, forecast of deaths & cases. There was no specific color theme applied for plotting the graphs. The dashboard also was not fitted on a single screen.

<center>
![](8.png)<br/>
**Figure 08: COVID-19 in Canada dashboard**
</center>
<br/>

“Zoho analytics COVID-19 live dashboard” for India contained three panels as insight by state, trend analysis & vaccination [1]. Interactive country map has been used to visualize the distribution of confirmed cases, active cases, deaths, recovered cases & vaccination by states in India. Line charts, bar charts & doughnut shape pie charts have been used as other visualization tools. Also this dashboard contained data table for overview of COVID-19 in India. This dashboard also was not fitted on a single screen.

<center>
![](9.png)<br/>
**Figure 09: India COVID-19 Dashboard**
</center>
<br/>

Italy COVID-19 dashboard created by the COVID-19 task force of the department of infection diseases and the IT service Istituto Superiore Sanita, contained interactive map, bar charts, doughnut shape pie charts & heat maps as visualization tools [4]. It provides data related to COVID-19 cases, cases distribution by age, clinical status, and weekly number of cases notified in Italy by region/province. This dashboard also was not fitted on a single screen.

<center>
![](10.png)<br/>
**Figure 10: Italy COVID-19 Dashboard**
</center>
<br/>

Live COVID-19 dashboard developed & contributed by an amber innovations for the government of Jamaica has added interactive country map & data table to represent the COVID-
19 cases by parish [9]. Line charts & bar charts have been added to the dashboard for cumulative & daily COVID-19 cases. Yellow, green & red colors have been used to represent confirmed, recovered & deaths respectively. Doughnut shape pie chart has been added to represent the distribution of recovered, confirmed & deaths by age group. This dashboard was not fitted on a single screen.

<center>
![](11.png)<br/>
**Figure 11: Jamaica COVID-19 Dashboard**
</center>
<br/>

The GCI COVID-19 dashboard for Russia includes an interactive map, bar charts & line charts to visualize the COVID-19 data [13]. Yellow, green, red & blue colors have been added to
identify the confirmed, recovered, deaths & active case respectively in plotting the charts.

<center>
![](12.png)<br/>
**Figure 12: GCI COVID-19 Dashboard for Russia**
</center>
<br/>

COVID-19 live situation analysis dashboard of Sri Lanka contains total & daily figures in Sri Lanka as counts [5]. Total cases & active cases have been visualized by using a line chart. Bar charts have been used to summarize the daily confirmed cases and recovered cases. As well fatality rate, recovery rate & daily investigations using PCR tests & rapid antigen tests have been compared by using the bar charts. Doughnut shape pie chart has been used to summarize the total cases. Blue, green & red colors have been applied to active, recovered & deaths respectively. This dashboard was also not fitted on a single screen.

<center>
![](13.png)<br/>
**Figure 13: COVID-19 Live Situational Analysis Dashboard of Sri Lanka**
</center>
<br/>

COVID-19 live situation analysis dashboard of Sri Lanka contains total & daily figures in Sri Lanka as counts [5]. Total cases & active cases have been visualized by using a line chart. Bar charts have been used to summarize the daily confirmed cases and recovered cases. As well fatality rate, recovery rate & daily investigations using PCR tests & rapid antigen tests have been compared by using the bar charts. Doughnut shape pie chart has been used to summarize the total cases. Blue, green & red colors have been applied to active, recovered & deaths respectively. This dashboard was also not fitted on a single screen.

<center>
![](14.png)<br/>
**Figure 14:COVID 19 ZA South Africa Dashboard**
</center>
<br/>

Robert Koch-Institute COVID-19 dashboard for German contains interactive map, line charts & bar charts as visualization tools of COVID-19 data [12]. The dashboard has been fitted
on a single screen. This dashboard format is similar to the John Hopkins dashboard.

<center>
![](15.png)<br/>
**Figure 15: Robert Koch-Institute COVID-19 dashboard for German**
</center>
<br/>

# Conclusion

Bar charts & line charts are the most frequently used tools for the visualization of total cases, daily cases & comparisons with respect to time. Some dashboards contained doughnut shape pie charts to summarize the total figures. In almost each and every dashboard, value boxes have been used to represent total figures. Some dashboards contained interactive maps & data tables to visualize the distribution of cases by country, province, region or state. The majority of the dashboards are daily updating real time dashboards. Gender, age groups and ethnicity can be identified as common breakdowns. The data sets & related links are available on most of the dashboards & can be downloaded. It can be seen that it is very easy, clear and user friendly to identify confirmed, recovered and deceased cases in dashboards which include one color theme for the whole dashboard. Dashboards with the dark background are comfortable to eyes than dashboards with light background & light colors. Also, it is better if the dashboard can be fitted on a single screen rather than adjusting through grid overlay.

# References

[1] COVID 19 Dashboard India – ZOHO Analytics – ZOHO, viewed 01 September 2021,
https://www.zoho.com/covid/india/

[2] COVID-19 dashboard – NHS providers, viewed 31 August 2021,
https://nhsproviders.org/topics/covid-19/tracking-covid-19-data/covid-19-dashboard

[3] COVID-19 in Belgium – Dashboard, viewed 31 August 2021,https://esribelux.maps.arcgis.com/apps/opsdashboard/index.html#/e350724c87af49bb9ce29646f8a42344

[4] COVID-19 integrated surveillance data in Italy – EpiCentro, viewed 02 September 2021,
https://www.epicentro.iss.it/en/coronavirus/sars-cov-2-dashboard

[5] COVID-19: Live situational Analysis Dashboard of Sri Lanka, viewed 08 September 2021,
https://hpb.health.gov.lk/covid19-dashboard/

[6] COVID-19 Map – Johns Hopkins Coronavirus Resource Center, viewed 31 August 2021,
https://coronavirus.jhu.edu/map.html

[7] COVID-19 Surveillance Dashboard – NSSAC Research, viewed 08 September 2021,
https://nssac.bii.virginia.edu/covid-19/dashboard/

[8] COVID 19 ZA Dashboard – Data Studio, viewed 08 September 2021,
https://datastudio.google.com/u/0/reporting/1b60bdc7-bec7-44c9-ba29-be0e043d8534/page/hrUIB

[9] Ministry of Health and Wellness, viewed 02 September 2021, https://jamcovid19.moh.gov.jm/

[10] New Zealand COVID-19 Surveillance Dashboard, viewed 27 August 2021,
https://nzcoviddashboard.esr.cri.nz/#!/

[11] Pakistan’s Official COVID-19 Dashboard – Shifa International Hospitals Ltd, viewed 01
September 2021, https://www.shifa.com.pk/covid-19-pakistan/

[12] RKI COVID-19 Germany – ArcGIS Experience, viewed 08 September 2021,
https://experience.arcgis.com/experience/478220a4c454480e823b17327b2bf1d4

[13] The Global COVId-19 Index (GCI) – Russia Dashboard – PEMANDU Associates, viewed
31 August 2021, https://covid19.pemandu.org/Russia

[14] Track COVID-19 Across Canada Using Our Interactive Dashboards, viewed 01 September
2021, https://samples.dundas.com/Dashboard/62cef916-5488-4d3b-9c57-204092d01813?e=false&vo=viewonly

[15] WHO Coronavirus (COVID-19) Dashboard, viewed 01 September 2021,
https://covid19.who.int/?gclid=CjwKCAjwvuGJBhB1EiwACU1Aia9pAaNLOkBMS2KVlSqXptskt1t46b7_bNv6vpTt_HjvXfrnDS-rWxoCE3sQAvD_BwE

### 2022-04-20 paper/Review paper/Review Paper.Rmd
Updated paper

---
title: "Review of COVID-19 Dashboards"
author: "Randi Shashikala"
date: "10/4/2021"
output: pdf_document
---

# 1. Abstract

There are 221 countries & territories which are affected by the COVID-19 global pandemic as the 20th of September 2021. To visualize the COVID-19 data & inform the public, dashboards are the most commonly used visualization tool. The main objective of this study is to identify different features, visualization methods & improvements that should be occurred by exploring the existing dashboards. 15 different dashboards that have been developed to visualize the COVID-19 data for different countries & globally has been explored here. Bar charts, line charts & Interactive maps are the most frequently used visualization tools in developing the dashboards. Dashboards that are fitted on a single screen is better than others.

**Key words** : COVID-19, Dashboard, Visualization

# 2. Introduction

The COVID-19 pandemic is the ongoing global pandemic of coronavirus disease 2019
caused by the Severe Acute Respiratory Syndrome Coronavirus 2 (SARS-CoV-2) virus. The first
patient infected with COVID-19 was identified in Wuhan, China, in December 2019. There have been reported 229835231 confirmed cases, 206515718 recovered cases, 18605877 active cases & 4713636 deaths in global as of 20th September 2021 according to the Worldmeter records. Currently, 99.5% of active patients are mild condition and 0.5% of patients are in critical situation.

Dashboards are one of the best visual interpretation method of tracking the spread & communication of COVID-19 pandemic. There are so many COVID-19 visualization dashboards that have been created to visualize the global & the local situation of the pandemic. Dashboards can be created using different software. There have been explored 15 dashboards which are developed to visualize the COVID-19 data in global & countries. In this paper has been compared dashboards to identify the different features, visualization methods & improvements that should be occurred.

### 2.1 Data represent in the Dashboards

Initially dashboards have been compared with the type of data to get an idea about which type of data is most frequently visualized in the dashboards.
The dashboards are labeled as follows;

**Table 01: Labeled of the Dashboards**

|No |Name of the Dashboard|
|:--:|:-------------------|
|1 | COVID-19 dashboard created by the John Hopkins University Center for Systems Science & Engineering (JHU CSSE) 
|2 | WHO  COVID-19 Dashboard 
|3 | COVID-19 surveillance dashboard created by the University of Virginia 
|4 | Corona cases (COVID-19) per municipality in Belgium dashboard 
|5 | COVID-19 dashboard for England created by NHS providers 
|6 | NZ COVID-19 Dashboard 
|7 | Pakistan’s official COVID-19 dashboard 
|8 | COVID-19 Canada live dashboard
|9 | India (COVID-19) Dashboard  
|10 | Italy COVID-19 dashboard 
|11 | Jamaica COVID-19 Dashboard 
|12 | GCI COVID-19 dashboard for Russia 
|13 | COVID-19 live situation analysis dashboard of Sri Lanka
|14 | COVID 19 ZA South Africa Dashboard 
|15 | COVID-19 dashboard for German

**Table 02: Summary of data represent in the dashboards**

| Name of the Dashboard| Location (Represented) | Confirmed Cases| Recovered Cases| Deaths | Vaccination Details| Tests| Global Comparison|
|:------------:|:------------|:------------:|:------------:|:------------:| :------------:|:------------:|:------------:|  
| 1                     | Global | \checkmark        |   \checkmark       |\checkmark  | \checkmark             |   | \checkmark 
| 2                     | Global |\checkmark        |   \checkmark        |\checkmark  | \checkmark              |   | \checkmark 
| 3                     | Global  |   \checkmark        |\checkmark  | \checkmark              |   | \checkmark 
| 4                     | Belgium |\checkmark          |   \checkmark        |\checkmark  |                     |  |  
| 5                     | England |\checkmark          | \checkmark        |\checkmark  | \checkmark              |  |
| 6                     | New Zealand | \checkmark          |   \checkmark       |\checkmark |                     |  |\checkmark 
| 7                     | Pakistan | \checkmark         |  \checkmark        |\checkmark  |                     |\checkmark  |
| 8                     | Canada | \checkmark          |  \checkmark        |\checkmark  |                     |  | |
| 9                     | India | \checkmark         |   \checkmark        |\checkmark  | \checkmark              | | \checkmark 
| 10                    | Italy | \checkmark         |   \checkmark        |\checkmark  |                     |\checkmark  |
| 11                    | Jamaica | \checkmark        |   \checkmark        |\checkmark  | \checkmark            | | 
| 12                    | Russia |\checkmark         |  \checkmark       |\checkmark  |                     | | 
| 13                    | Sri Lanka |\checkmark          |   \checkmark       |\checkmark  |                     | \checkmark  | \checkmark 
| 14                    | South Africa |\checkmark        |   \checkmark        |\checkmark  | \checkmark              | \checkmark  |
| 15                    | German |\checkmark          |  \checkmark        |\checkmark  | \checkmark              | |

As shown in table 02 all dashboards which are considered in this paper represent the data related to COVID-19 confirmed cases, recovered cases & deaths. There have been contained vaccination details, 8 dashboards out of 15 dashboards. In almost each & every dashboard, value boxes have been used to represent these total figures. Bar charts & line charts (trend lines) are most frequently used to visualize these data (confirmed cases, recovered cases, deaths & vaccination details) with respect to time. A Majority of dashboards presented data by daily or weekly. The mapping is used to track the spatial distribution of COVID-19 cases by country/ provincial/ regional etc. When visualizing the data by map color code system & circles with respect to the size of the cases have been used to visualize the variation in size. The considerable number of dashboards has been used doughnut shape pie charts to represent total COVID-19 confirmed cases, recovered cases, active cases & deaths as a percentage. Also, region, gender, age group & ethnicity can be identified as common breakdowns of COVID-19 cases. In some dashboard has been added data tables for representing cases distribution by province/region.  Only very few dashboards have been visualized in COVID-19 test details. Only 6 dashboards have been compared to global situations. In addition, fatality rate, incidence rate, ICU beds, stage of the patients & hospitalize details have been contained in the several dashboards. These are the most common data types that have been contained in the dashboards. The main tools that have been used for different purpose can be shown in the following table.\hfill\break

**Table 03: Summary of tools which are used for different purpose**

|Purpose | Bar chart | Line chart | Pie chart | Dot plot| Heat map| Mapping| Data table|
|:-------|:---------:|:----------:|:---------:|:--------:|:--------:|:-------:|:-------:|
| COVID-19 confirmed 
cases | \checkmark  | \checkmark  |  | \checkmark  |  | \checkmark  |\checkmark 
| COVID-19 deaths | \checkmark  | \checkmark  |  |  | | \checkmark  | \checkmark 
| COVID-19 recovered 
cases | \checkmark  | \checkmark  | | | | \checkmark  | \checkmark
| COVID-19 vaccination | | \checkmark  |  | | | \checkmark  | \checkmark
| COVID-19 test 
conducted | \checkmark  | \checkmark  | | | | |
| Clinical status | \checkmark  | | | | | | 
| Cases distribution by 
age  | \checkmark  | | \checkmark  | | |  |
| Cases distribution by 
gender | \checkmark  | | | | | |
| Cases distribution by area
(Province/state/region) | \checkmark  | \checkmark  |  | | \checkmark  | \checkmark  | \checkmark
| To compare the cases | | |  \checkmark  | | |  |  \checkmark 
| Global comparison |  \checkmark  |  \checkmark  | | | |  \checkmark  | \checkmark

### 2.2 Comparison of Dashboards

Before developing a dashboard, it is necessary to think about which visualization tools & features that should be contained in the dashboards. What are the most suitable plots, how many panels in the dashboard, what data should be included, how to fit dashboard on a screen, colors & is it real time updated or not are the common things that should be considered before develop the dashboards. In table 02 has been compared these visualization tools & features under the following categories.

  * Number of panels - How many panels which are included in the dashboard.
  * Visualization tools – what are the graphical representations of data which are contained in the dashboards. 
  * Fitted on a single screen – whether the dashboard fits on a single screen or not (users can see the whole dashboard on a single screen without adjusting through grid overlay or not).
  * Color theme – is there a unique color used for one data type in the whole dashboard (i.e.: one color scale for one data type everywhere on the dashboard). 
  * Dark background – background color of the dashboard is dark or light.
  * Data available – whether users can be downloaded/available the data set which has been visualized on the dashboard. 
  * Real time updated – whether the dashboard is updated daily/ specific time (live dashboard) or not.

\hfill\break  
\hfill\break  
\hfill\break  
\hfill\break \hfill\break  
\hfill\break  
\hfill\break  
\hfill\break 
\hfill\break  
\hfill\break  
\hfill\break  
\hfill\break

**Table 04: Comparison of visualization tools & features of dashboard**

| Name of the Dashboard | Number of panels|  Visualization tools | Fitted on a single screen | Color theme | Dark background | Data available | Real time updated|
|:-------: |:-------: |:-------:  |:-------:|:-------:   |:-------:  |:-------:   |:-------:  |
| 1                |  1   | Bar chart\hfill\break  Interactive map\hfill\break |\checkmark         |   \checkmark      |\checkmark | \checkmark   | \checkmark
| 2             | 4   | Line chart\hfill\break  Interactive map\hfill\break  Data table\hfill\break|         |   \checkmark       |    |     \checkmark        |\checkmark 
| 3         |2         |Line chart\hfill\break  Bar chart\hfill\break Interactive map\hfill\break Data table\hfill\break  | \checkmark         |   \checkmark       |\checkmark | \checkmark             |\checkmark
|4 | 1|Line chart\hfill\break Bar chart\hfill\break  Pie chart\hfill\break Interactive map\hfill\break |\checkmark |   | \checkmark |\checkmark| \checkmark
|5 | 1| Line chart\hfill\break Bar chart\hfill\break Data table\hfill\break |   | \checkmark |   | \checkmark | \checkmark
|6 | 5| Line chart\hfill\break Bar chart\hfill\break Dot plot\hfill\break  Interactive country map\hfill\break |   |   |   | \checkmark | \checkmark
|7 | 1|  Line chart\hfill\break Bar chart\hfill\break  Country map\hfill\break Data table\hfill\break |    | \checkmark  |    |    | \checkmark
|8 | 3| Line chart\hfill\break Bar chart\hfill\break Data table\hfill\break Interactive map\hfill\break |    |    | \checkmark |   | \checkmark
|9 |3 | Line chart\hfill\break Bar chart\hfill\break Doughnut shape pie chart\hfill\break Data table\hfill\break Interactive country map\hfill\break |  | \checkmark |  | \checkmark | \checkmark 
|10  | 2| Bar chart\hfill\break Doughnut shape pie chart\hfill\break Heat map\hfill\break Interactive country map\hfill\break |   |   |    |    | \checkmark
|11 | 1| Line chart\hfill\break Bar chart\hfill\break Doughnut shape pie chart\hfill\break Data table\hfill\break Interactive country map\hfill\break |  | \checkmark |   |   | \checkmark
|12 | 1| Line chart\hfill\break Bar chart\hfill\break  Interactive map\hfill\break |   | \checkmark | \checkmark |   | \checkmark
|13 | 1| Line chart\hfill\break Bar chart\hfill\break Doughnut shape pie chart\hfill\break |  |\checkmark |   |   | \checkmark
|14| 2| Line chart\hfill\break Bar chart\hfill\break  Interactive Country map\hfill\break |  | \checkmark|   | \checkmark | \checkmark
|15 | 1| Line chart\hfill\break Bar chart\hfill\break  Data table\hfill\break Interactive map\hfill\break | \checkmark| \checkmark| \checkmark|  | \checkmark

As shown in table 04 almost each & every dashboard, line charts & bar charts have been used to visualize the data. Heat map & dot plot has been used only one dashboard. Only four dashboards have been fitted with a single screen. The majority of dashboards have used color theme on the whole dashboard. That is, dashboards have been applied different colors for different type of data (i.e. One specific color for confirmed cases, another color for deaths, etc.) in the whole dashboard. The data set & related links have available on some dashboards & users can download these data sets. There are 6 dashboards with dark background while others have been used light background. Last updated time & date of the latest available data  has been reported at the top or bottom of the first panel in the dashboard. Like, half of dashboards included all data in one panel.

# 3. Conclusions & Discussion

Bar charts & line charts are the most frequently used tools for the visualization of total cases, daily cases & comparisons with respect to time. Some dashboards contained doughnut shape pie charts to summarize the total figures. In almost each and every dashboard, value boxes have been used to represent total figures. Some dashboards contained interactive maps & data tables to visualize the distribution of cases by country, province, region or state. All dashboards are daily updating real time dashboards. Gender, age groups and ethnicity can be identified as common breakdowns. The data sets & related links are available on most of the dashboards & can be downloaded. It can be seen that it is very easy, clear and user friendly to identify confirmed, recovered and deceased cases in dashboards which include one color theme for the whole dashboard. Dashboards with the dark background are more comfortable to the eyes than dashboards with light background & light colors. Also, it is better if the dashboard can be fitted on a single screen rather than adjusting through a grid overlay.

# 4. References

[1] COVID 19 Dashboard India – ZOHO Analytics – ZOHO, viewed 01 September 2021,
https://www.zoho.com/covid/india/

[2] COVID-19 dashboard – NHS providers, viewed 31 August 2021,
https://nhsproviders.org/topics/covid-19/tracking-covid-19-data/covid-19-dashboard

[3] COVID-19 in Belgium – Dashboard, viewed 31 August 2021,https://esribelux.maps.arcgis.com/apps/opsdashboard/index.html#/e350724c87af49bb9ce29646f8a42344

[4] COVID-19 integrated surveillance data in Italy – EpiCentro, viewed 02 September 2021,
https://www.epicentro.iss.it/en/coronavirus/sars-cov-2-dashboard

[5] COVID-19: Live situational Analysis Dashboard of Sri Lanka, viewed 08 September 2021,
https://hpb.health.gov.lk/covid19-dashboard/

[6] COVID-19 Map – Johns Hopkins Coronavirus Resource Center, viewed 31 August 2021,
https://coronavirus.jhu.edu/map.html

[7] COVID-19 Surveillance Dashboard – NSSAC Research, viewed 08 September 2021,
https://nssac.bii.virginia.edu/covid-19/dashboard/

[8] COVID 19 ZA Dashboard – Data Studio, viewed 08 September 2021,
https://datastudio.google.com/u/0/reporting/1b60bdc7-bec7-44c9-ba29-be0e043d8534/page/hrUIB

[9] Ministry of Health and Wellness, viewed 02 September 2021, https://jamcovid19.moh.gov.jm/

[10] New Zealand COVID-19 Surveillance Dashboard, viewed 27 August 2021,
https://nzcoviddashboard.esr.cri.nz/#!/

[11] Pakistan’s Official COVID-19 Dashboard – Shifa International Hospitals Ltd, viewed 01
September 2021, https://www.shifa.com.pk/covid-19-pakistan/

[12] RKI COVID-19 Germany – ArcGIS Experience, viewed 08 September 2021,
https://experience.arcgis.com/experience/478220a4c454480e823b17327b2bf1d4

[13] The Global COVId-19 Index (GCI) – Russia Dashboard – PEMANDU Associates, viewed
31 August 2021, https://covid19.pemandu.org/Russia

[14] Track COVID-19 Across Canada Using Our Interactive Dashboards, viewed 01 September
2021, https://samples.dundas.com/Dashboard/62cef916-5488-4d3b-9c57-204092d01813?e=false&vo=viewonly

[15] WHO Coronavirus (COVID-19) Dashboard, viewed 01 September 2021,
https://covid19.who.int/?gclid=CjwKCAjwvuGJBhB1EiwACU1Aia9pAaNLOkBMS2KVlSqXptskt1t46b7_bNv6vpTt_HjvXfrnDS-rWxoCE3sQAvD_BwE

### 2022-04-20 manuscript_covid19dashboard.Rmd
updated

--- 2022-04-20 manuscript_covid19dashboard.Rmd
@@
 title: "Interactive dashboard to monitor the COVID-19 outbreak"
-output: pdf_document
+output: 
+  pdf_document: 
+    keep_tex: true
+    number_sections: true
 bibliography: references.bib
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
@@
 
-Department of Statistics
-
-Faculty of Applied Sciences
+Department of Statistics, Faculty of Applied Sciences
 
@@
 
-# Abstract
+**Abstract**
 
@@
 
-The 15 dashboards we used in the study are listed below (Table 1).
+Initially 15 dashboards were compared  to determine which data types are most frequently shown in dashboards. The 15 dashboards we used in the study are listed below (Table 1).
+
+**Table 01: Labeled of the Dashboards**
+
+|No |Name of the Dashboard|Reference|
+|:--:|:-------------------|------|
+|1 | COVID-19 dashboard created by the John Hopkins University Center for Systems Science & Engineering (JHU CSSE) |@jh|
+|2 | WHO  COVID-19 Dashboard |@who|
+|3 | COVID-19 surveillance dashboard created by the University of Virginia |@nssac|
+|4 | Corona cases (COVID-19) per municipality in Belgium dashboard |@belgium|
+|5 | COVID-19 dashboard for England created by NHS providers |@nhs|
+|6 | NZ COVID-19 Dashboard |@nz|
+|7 | Pakistan’s official COVID-19 dashboard |@pakistan|
+|8 | COVID-19 Canada live dashboard |@canada|
+|9 | India (COVID-19) Dashboard | @indiadash |
+|10 | Italy COVID-19 dashboard |@italy|
+|11 | Jamaica COVID-19 Dashboard |@jamaica|
+|12 | GCI COVID-19 dashboard for Russia |@russia|
+|13 | COVID-19 live situation analysis dashboard of Sri Lanka |@sl|
+|14 | COVID 19 ZA South Africa Dashboard |@za|
+|15 | COVID-19 dashboard for Germany |@rki|
+
+### Summary of data representation on dashboards
+
+**Table 02: Summary of data represent in the dashboards**
+
+| Name of the Dashboard| Location (Represented) | Confirmed Cases| Recovered Cases| Deaths | Vaccination Details| Tests| Global Comparison|
+|:------------:|:------------|:------------:|:------------:|:------------:| :------------:|:------------:|:------------:|  
+| 1                     | Global | \checkmark        |   \checkmark       |\checkmark  | \checkmark             |   | \checkmark 
+| 2                     | Global |\checkmark        |   \checkmark        |\checkmark  | \checkmark              |   | \checkmark 
+| 3                     | Global  |   \checkmark        |\checkmark  | \checkmark              |   | \checkmark 
+| 4                     | Belgium |\checkmark          |   \checkmark        |\checkmark  |                     |  |  
+| 5                     | England |\checkmark          | \checkmark        |\checkmark  | \checkmark              |  |
+| 6                     | New Zealand | \checkmark          |   \checkmark       |\checkmark |                     |  |\checkmark 
+| 7                     | Pakistan | \checkmark         |  \checkmark        |\checkmark  |                     |\checkmark  |
+| 8                     | Canada | \checkmark          |  \checkmark        |\checkmark  |                     |  | |
+| 9                     | India | \checkmark         |   \checkmark        |\checkmark  | \checkmark              | | \checkmark 
+| 10                    | Italy | \checkmark         |   \checkmark        |\checkmark  |                     |\checkmark  |
+| 11                    | Jamaica | \checkmark        |   \checkmark        |\checkmark  | \checkmark            | | 
+| 12                    | Russia |\checkmark         |  \checkmark       |\checkmark  |                     | | 
+| 13                    | Sri Lanka |\checkmark          |   \checkmark       |\checkmark  |                     | \checkmark  | \checkmark 
+| 14                    | South Africa |\checkmark        |   \checkmark        |\checkmark  | \checkmark              | \checkmark  |
+| 15                    | German |\checkmark          |  \checkmark        |\checkmark  | \checkmark              | |
+
+As shown in table 02 all dashboards which are considered in this paper represent the data related to COVID-19 confirmed cases, recovered cases & deaths.
+There have been contained vaccination details, 8 dashboards out of 15 dashboards.
+In almost each & every dashboard, value boxes have been used to represent these total figures.
+Bar charts & line charts (trend lines) are most frequently used to visualize these data (confirmed cases, recovered cases, deaths & vaccination details) with respect to time.
+A Majority of dashboards presented data by daily or weekly.
+The mapping is used to track the spatial distribution of COVID-19 cases by country/ provincial/ regional etc.
+When visualizing the data by map color code system & circles with respect to the size of the cases have been used to visualize the variation in size.
+The considerable number of dashboards has been used doughnut shape pie charts to represent total COVID-19 confirmed cases, recovered cases, active cases & deaths as a percentage.
+Also, region, gender, age group & ethnicity can be identified as common breakdowns of COVID-19 cases.
+In some dashboard has been added data tables for representing cases distribution by province/region.
+Only very few dashboards have been visualized in COVID-19 test details.
+Only 6 dashboards have been compared to global situations.
+In addition, fatality rate, incidence rate, ICU beds, stage of the patients & hospitalize details have been contained in the several dashboards.
+These are the most common data types that have been contained in the dashboards.
+The main tools that have been used for different purpose can be shown in the following table.\hfill\break
+
+**Table 03: Summary of tools which are used for different purpose**
+
+|Purpose | Bar chart | Line chart | Pie chart | Dot plot| Heat map| Mapping| Data table|
+|:-------|:---------:|:----------:|:---------:|:--------:|:--------:|:-------:|:-------:|
+| COVID-19 confirmed 
+cases | \checkmark  | \checkmark  |  | \checkmark  |  | \checkmark  |\checkmark 
+| COVID-19 deaths | \checkmark  | \checkmark  |  |  | | \checkmark  | \checkmark 
+| COVID-19 recovered 
+cases | \checkmark  | \checkmark  | | | | \checkmark  | \checkmark
+| COVID-19 vaccination | | \checkmark  |  | | | \checkmark  | \checkmark
+| COVID-19 test 
+conducted | \checkmark  | \checkmark  | | | | |
+| Clinical status | \checkmark  | | | | | | 
+| Cases distribution by 
+age  | \checkmark  | | \checkmark  | | |  |
+| Cases distribution by 
+gender | \checkmark  | | | | | |
+| Cases distribution by area
+(Province/state/region) | \checkmark  | \checkmark  |  | | \checkmark  | \checkmark  | \checkmark
+| To compare the cases | | |  \checkmark  | | |  |  \checkmark 
+| Global comparison |  \checkmark  |  \checkmark  | | | |  \checkmark  | \checkmark
+
+### 2.2 Comparison of Dashboards
+
+Before developing a dashboard, it is necessary to think about which visualization tools & features that should be contained in the dashboards.
+What are the most suitable plots, how many panels in the dashboard, what data should be included, how to fit dashboard on a screen, colors & is it real time updated or not are the common things that should be considered before develop the dashboards.
+In table 02 has been compared these visualization tools & features under the following categories.
+
+  * Number of panels - How many panels which are included in the dashboard.
+  * Visualization tools – what are the graphical representations of data which are contained in the dashboards. 
+  * Fitted on a single screen – whether the dashboard fits on a single screen or not (users can see the whole dashboard on a single screen without adjusting through grid overlay or not).
+  * Color theme – is there a unique color used for one data type in the whole dashboard (i.e.: one color scale for one data type everywhere on the dashboard). 
+  * Dark background – background color of the dashboard is dark or light.
+  * Data available – whether users can be downloaded/available the data set which has been visualized on the dashboard. 
+  * Real time updated – whether the dashboard is updated daily/ specific time (live dashboard) or not.
+
+**Table 04: Comparison of visualization tools & features of dashboard**
+
+| Name of the Dashboard | Number of panels|  Visualization tools | Fitted on a single screen | Color theme | Dark background | Data available | Real time updated|
+|:-------: |:-------: |:-------:  |:-------:|:-------:   |:-------:  |:-------:   |:-------:  |
+| 1                |  1   | Bar chart\hfill\break  Interactive map\hfill\break |\checkmark         |   \checkmark      |\checkmark | \checkmark   | \checkmark
+| 2             | 4   | Line chart\hfill\break  Interactive map\hfill\break  Data table\hfill\break|         |   \checkmark       |    |     \checkmark        |\checkmark 
+| 3         |2         |Line chart\hfill\break  Bar chart\hfill\break Interactive map\hfill\break Data table\hfill\break  | \checkmark         |   \checkmark       |\checkmark | \checkmark             |\checkmark
+|4 | 1|Line chart\hfill\break Bar chart\hfill\break  Pie chart\hfill\break Interactive map\hfill\break |\checkmark |   | \checkmark |\checkmark| \checkmark
+|5 | 1| Line chart\hfill\break Bar chart\hfill\break Data table\hfill\break |   | \checkmark |   | \checkmark | \checkmark
+|6 | 5| Line chart\hfill\break Bar chart\hfill\break Dot plot\hfill\break  Interactive country map\hfill\break |   |   |   | \checkmark | \checkmark
+|7 | 1|  Line chart\hfill\break Bar chart\hfill\break  Country map\hfill\break Data table\hfill\break |    | \checkmark  |    |    | \checkmark
+|8 | 3| Line chart\hfill\break Bar chart\hfill\break Data table\hfill\break Interactive map\hfill\break |    |    | \checkmark |   | \checkmark
+|9 |3 | Line chart\hfill\break Bar chart\hfill\break Doughnut shape pie chart\hfill\break Data table\hfill\break Interactive country map\hfill\break |  | \checkmark |  | \checkmark | \checkmark 
+|10  | 2| Bar chart\hfill\break Doughnut shape pie chart\hfill\break Heat map\hfill\break Interactive country map\hfill\break |   |   |    |    | \checkmark
+|11 | 1| Line chart\hfill\break Bar chart\hfill\break Doughnut shape pie chart\hfill\break Data table\hfill\break Interactive country map\hfill\break |  | \checkmark |   |   | \checkmark
+|12 | 1| Line chart\hfill\break Bar chart\hfill\break  Interactive map\hfill\break |   | \checkmark | \checkmark |   | \checkmark
+|13 | 1| Line chart\hfill\break Bar chart\hfill\break Doughnut shape pie chart\hfill\break |  |\checkmark |   |   | \checkmark
+|14| 2| Line chart\hfill\break Bar chart\hfill\break  Interactive Country map\hfill\break |  | \checkmark|   | \checkmark | \checkmark
+|15 | 1| Line chart\hfill\break Bar chart\hfill\break  Data table\hfill\break Interactive map\hfill\break | \checkmark| \checkmark| \checkmark|  | \checkmark
+
+As shown in table 04 almost each & every dashboard, line charts & bar charts have been used to visualize the data.
+Heat map & dot plot has been used only one dashboard.
+Only four dashboards have been fitted with a single screen.
+The majority of dashboards have used color theme on the whole dashboard.
+That is, dashboards have been applied different colors for different type of data (i.e.
+One specific color for confirmed cases, another color for deaths, etc.) in the whole dashboard.
+The data set & related links have available on some dashboards & users can download these data sets.
+There are 6 dashboards with dark background while others have been used light background.
+Last updated time & date of the latest available data  has been reported at the top or bottom of the first panel in the dashboard.
+Like, half of dashboards included all data in one panel.
 
@@
 
-**Table 01: Description of the panels**
+**Table 05: Description of the panels**
 
@@
 
+Bar charts & line charts are the most frequently used tools for the visualization of total cases, daily cases & comparisons with respect to time.
+Some dashboards contained doughnut shape pie charts to summarize the total figures.
+In almost each and every dashboard, value boxes have been used to represent total figures.
+Some dashboards contained interactive maps & data tables to visualize the distribution of cases by country, province, region or state.
+All dashboards are daily updating real time dashboards.
+Gender, age groups and ethnicity can be identified as common breakdowns.
+The data sets & related links are available on most of the dashboards & can be downloaded.
+It can be seen that it is very easy, clear and user friendly to identify confirmed, recovered and deceased cases in dashboards which include one color theme for the whole dashboard.
+Dashboards with the dark background are more comfortable to the eyes than dashboards with light background & light colors.
+Also, it is better if the dashboard can be fitted on a single screen rather than adjusting through a grid overlay.
+
+# Appendix: Overview of reviewed dashboard
+
+Real time updated COVID-19 dashboard created by the John Hopkins University Center for Systems Science & Engineering (JHU CSSE) has included total confirmed cases & total deaths by country, province/region/state (not for all countries)[6].
+Weekly & daily global confirmed cases, deaths & vaccine doses have been visualized using bar plots related to the date in the side panels.
+Also, an interactive world map has included in the dashboard to represent total confirmed cases, incidence rate, fatality ratio & vaccine doses administered by country & state of US.
+All the links related to the data sets has been included in the dashboard.
+In this dashboard red, white, green colors has been used to indicate confirmed cases, deaths & vaccine doses respectively.
+The dashboard has fitted on a single screen.
+
+```{r, fig.cap="Johns Hopkins University COVID-19 Dashboard"}
+knitr::include_graphics("Images/1.PNG")
+```
+
+Live updated “WHO COVID-19 dashboard” contained four panels. The interactive world
+map in the dashboard has visualized the distribution of COVID-19 confirmed cases, deaths &
+vaccination by country [15].
+Line chart & bar charts have been used as visualization plots.
+Blue, red, green colors have been used to represent confirmed cases, deaths & vaccination respectively.
+The links related to the underlying data was available in the dashboard.
+The dashboard was not fitted on a single screen.
+
+```{r, fig.cap="WHO COVID-19 Dashboard"}
+knitr::include_graphics("Images/2.PNG")
+```
+
+The “COVID-19 surveillance dashboard” created by the University of Virginia in
+collaboration with Bio complexity Institute has included interactive world map with a time slide to visualize the confirmed cases, recovered cases, deaths, active cases & vaccine details [7].
+Line charts, bar charts & data tables have been visualized in the dashboard.
+Red, blue, green & yellow colors have been used to identify confirmed, deceased, recovered & active cases respectively.
+The data behind the dashboard also available on the dashboard & can be downloaded.
+The dashboard has fitted on a single screen.
+
+```{r, fig.cap="COVID-19 Surveillance Dashboard"}
+knitr::include_graphics("Images/3.PNG")
+```
+
+Corona cases (COVID-19) per municipality in Belgium dashboard has included interactive
+map for distribution of total confirmed cases per municipality, bar chart for number of cases per province, line charts for number of cases per day by region & municipality, dodge bar chart for deaths per region, per age & pie chart for deaths per age as visualization methods[3].
+Also it included several value boxes for total cases, total deaths & hospital situations.
+There was no specific color theme used in this dashboard.
+This dashboard also fitted on a single screen.
+
+```{r, fig.cap="Corona cases per Municipality in Belgium Dashboard"}
+knitr::include_graphics("Images/4.PNG")
+```
+
+The COVID-19 dashboard for England created by NHS providers included daily totals, daily changes & weekly changes in new cases, hospital admission, patients in hospital, and patients
+in ICU beds, deaths & vaccinations in England as counts [2].
+Bar plots for daily new cases, hospital admissions & deaths for last 30 days have been added in this dashboard.
+Also it has included data table for total count, daily changes (as increase or decrease), weekly changes (as increase or decrease) of new cases & hospital admission by region.
+It can be seen that, three different colors have been used in bar plot as orange for new cases, green for hospital admissions & red for deaths.
+These plots were not interactive plots & there was only one panel.
+The dashboard was not fitted on a single screen.
+
+```{r, fig.cap="COVID-19 dashboard for England"}
+knitr::include_graphics("Images/5.PNG")
+```
+
+The Environmental Science and Research Institute COVID-19 dashboard for New Zealand (was developed & is maintained in collaboration with Epi-interactive Ltd) contained five panels
+as overview, outbreak, source, international & ESR reports [10].
+In this real time updated dashboard, they have included value boxes for total confirmed cases, recovered cases, deaths & interactive map has been added for distribution of confirmed & probable cases as incidence & count.
+Also they have included bar charts for confirmed cases distribution by gender, age group, & line charts for confirmed & deaths as daily & cumulative.
+The dashboard has not fitted on a single screen.
+
+```{r, fig.cap="New Zealand COVID-19 Dashboard"}
+knitr::include_graphics("Images/6.PNG")
+```
+
+Pakistan’s official COVID-19 dashboard is a real time updated dashboard with one panel. This dashboard contains value boxes for total confirmed cases, active cases, deaths, recovered
+cases by country.
+Province wise cases have been represented by map & data table [11].
+Line charts & bar charts have been added to visualize the distribution of cases related to time & province.
+In the value boxes blue, orange, pink & green colors have been applied for confirmed cases, active cases, deaths & recoveries respectively, but there was no specific color theme used when plotting the graphs.
+The dashboard cannot be seen on a single screen without grid overlay.
+
+```{r, fig.cap="COVID-19 Dashboard for Pakistan"}
+knitr::include_graphics("Images/7.PNG")
+```
+
+COVID-19 Canada live dashboard has been developed with three panels for provincial data, hospital resources & projections [14].
+This dashboard contained interactive map for represented total reported cases & deaths by province.
+Data tables, line charts, bar charts were the most frequently used visualization types of this dashboard.
+More details were presented on this dashboard by province like ventilator counts, hospital beds per 100 residents, resource capacity based on critical case rate, effect of social distance, forecast of deaths & cases.
+There was no specific color theme applied for plotting the graphs.
+The dashboard also was not fitted on a single screen.
+
+```{r, fig.cap="COVID-19 in Canada dashboard"}
+knitr::include_graphics("Images/8.PNG")
+```
+
+“Zoho analytics COVID-19 live dashboard” for India contained three panels as insight by state, trend analysis & vaccination [1].
+Interactive country map has been used to visualize the distribution of confirmed cases, active cases, deaths, recovered cases & vaccination by states in India.
+Line charts, bar charts & doughnut shape pie charts have been used as other visualization tools.
+Also this dashboard contained data table for overview of COVID-19 in India.
+This dashboard also was not fitted on a single screen.
+
+```{r, fig.cap="India COVID-19 Dashboard"}
+knitr::include_graphics("Images/9.PNG")
+```
+
+Italy COVID-19 dashboard created by the COVID-19 task force of the department of infection diseases and the IT service Istituto Superiore Sanita, contained interactive map, bar charts, doughnut shape pie charts & heat maps as visualization tools [4].
+It provides data related to COVID-19 cases, cases distribution by age, clinical status, and weekly number of cases notified in Italy by region/province.
+This dashboard also was not fitted on a single screen.
+
+```{r, fig.cap="Italy COVID-19 Dashboard"}
+knitr::include_graphics("Images/10.PNG")
+```
+
+Live COVID-19 dashboard developed & contributed by an amber innovations for the government of Jamaica has added interactive country map & data table to represent the COVID-
+19 cases by parish [9].
+Line charts & bar charts have been added to the dashboard for cumulative & daily COVID-19 cases.
+Yellow, green & red colors have been used to represent confirmed, recovered & deaths respectively.
+Doughnut shape pie chart has been added to represent the distribution of recovered, confirmed & deaths by age group.
+This dashboard was not fitted on a single screen.
+
+```{r, fig.cap="Jamaica COVID-19 Dashboard"}
+knitr::include_graphics("Images/11.PNG")
+```
+
+The GCI COVID-19 dashboard for Russia includes an interactive map, bar charts & line charts to visualize the COVID-19 data [13]. Yellow, green, red & blue colors have been added to
+identify the confirmed, recovered, deaths & active case respectively in plotting the charts.
+
+```{r, fig.cap="GCI COVID-19 Dashboard for Russia"}
+knitr::include_graphics("Images/12.PNG")
+```
+
+COVID-19 live situation analysis dashboard of Sri Lanka contains total & daily figures in Sri Lanka as counts [5].
+Total cases & active cases have been visualized by using a line chart.
+Bar charts have been used to summarize the daily confirmed cases and recovered cases.
+As well fatality rate, recovery rate & daily investigations using PCR tests & rapid antigen tests have been compared by using the bar charts.
+Doughnut shape pie chart has been used to summarize the total cases.
+Blue, green & red colors have been applied to active, recovered & deaths respectively.
+This dashboard was also not fitted on a single screen.
+
+```{r, fig.cap="COVID-19 Live Situational Analysis Dashboard of Sri Lanka"}
+knitr::include_graphics("Images/13.PNG")
+```
+
+COVID-19 live situation analysis dashboard of Sri Lanka contains total & daily figures in Sri Lanka as counts [5].
+Total cases & active cases have been visualized by using a line chart.
+Bar charts have been used to summarize the daily confirmed cases and recovered cases.
+As well fatality rate, recovery rate & daily investigations using PCR tests & rapid antigen tests have been compared by using the bar charts.
+Doughnut shape pie chart has been used to summarize the total cases.
+Blue, green & red colors have been applied to active, recovered & deaths respectively.
+This dashboard was also not fitted on a single screen.
+
+```{r, fig.cap="COVID 19 ZA South Africa Dashboard"}
+knitr::include_graphics("Images/14.PNG")
+```
+
+Robert Koch-Institute COVID-19 dashboard for German contains interactive map, line charts & bar charts as visualization tools of COVID-19 data [12]. The dashboard has been fitted
+on a single screen. This dashboard format is similar to the John Hopkins dashboard.
+
+```{r, fig.cap="Robert Koch-Institute COVID-19 dashboard for German"}
+knitr::include_graphics("Images/15.PNG")
+```
+
 # References

### 2022-04-20 manuscript_covid19dashboard.tex
updated

@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{5}
 \ifluatex
@@
 
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
+\textbf{Abstract}
 
@@
 1 & COVID-19 dashboard created by the John Hopkins University Center for
-Systems Science \& Engineering (JHU CSSE) & \\
-2 & WHO COVID-19 Dashboard & \\
+Systems Science \& Engineering (JHU CSSE) & {``{COVID-19 Map -- Johns
+Hopkins Coronavirus Resource Center}''} (2022) \\
+2 & WHO COVID-19 Dashboard & {``{WHO Coronavirus (COVID-19)
+Dashboard}''} (2021) \\
 3 & COVID-19 surveillance dashboard created by the University of
-Virginia & \\
+Virginia & {``{COVID-19 Surveillance Dashboard -- NSSAC Research}''}
+(2022) \\
 4 & Corona cases (COVID-19) per municipality in Belgium dashboard &
@@
 19 Dashboard -- NHS providers}''} (2022) \\
-6 & NZ COVID-19 Dashboard & \\
-7 & Pakistan's official COVID-19 dashboard & \\
-8 & COVID-19 Canada live dashboard & \\
+6 & NZ COVID-19 Dashboard & {``{New Zealand COVID-19 Surveillance
+Dashboard}''} (2021) \\
+7 & Pakistan's official COVID-19 dashboard & {``{Pakistan's Official
+COVID-19 Dashboard -- Shifa International Hospitals Ltd}''} (2021) \\
+8 & COVID-19 Canada live dashboard & {``{Track COVID-19 Across Canada
+Using Our Interactive Dashboards}''} (2021) \\
 9 & India (COVID-19) Dashboard & {``{COVID 19 Dashboard India -- ZOHO
@@
 data in Italy -- EpiCentro}''} (2022) \\
-11 & Jamaica COVID-19 Dashboard & \\
-12 & GCI COVID-19 dashboard for Russia & \\
+11 & Jamaica COVID-19 Dashboard & {``{COVID-19 Jamaica - Ministry of
+Health and Wellness}''} (2021) \\
+12 & GCI COVID-19 dashboard for Russia & {``{The Global COVId-19 Index
+(GCI) -- Russia Dashboard -- PEMANDU Associates}''} (2021) \\
 13 & COVID-19 live situation analysis dashboard of Sri Lanka &
-{``{COVID-19: Live situational Analysis Dashboard of Sri Lanka,}''}
+{``{COVID-19: Live situational Analysis Dashboard of Sri Lanka}''}
 (2022) \\
-14 & COVID 19 ZA South Africa Dashboard & \\
-15 & COVID-19 dashboard for Germany & \\
+14 & COVID 19 ZA South Africa Dashboard & {``{COVID-19 ZA Dashboard -
+Data Studio}''} (2021) \\
+15 & COVID-19 dashboard for Germany & {``{RKI COVID-19 Germany -- ArcGIS
+Experience}''} (2021) \\
 \bottomrule
 \end{longtable}
+
+\hypertarget{summary-of-data-representation-on-dashboards}{%
+\subsubsection{Summary of data representation on
+dashboards}\label{summary-of-data-representation-on-dashboards}}
+
+\textbf{Table 02: Summary of data represent in the dashboards}
+
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.13}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.12}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.13}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.13}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.13}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.13}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.13}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.13}}@{}}
+\toprule
+Name of the Dashboard & Location (Represented) & Confirmed Cases &
+Recovered Cases & Deaths & Vaccination Details & Tests & Global
+Comparison \\
+\midrule
+\endhead
+1 & Global & \checkmark & \checkmark & \checkmark & \checkmark & &
+\checkmark \\
+2 & Global & \checkmark & \checkmark & \checkmark & \checkmark & &
+\checkmark \\
+3 & Global & \checkmark & \checkmark & \checkmark & & \checkmark & \\
+4 & Belgium & \checkmark & \checkmark & \checkmark & & & \\
+5 & England & \checkmark & \checkmark & \checkmark & \checkmark & & \\
+6 & New Zealand & \checkmark & \checkmark & \checkmark & & &
+\checkmark \\
+7 & Pakistan & \checkmark & \checkmark & \checkmark & & \checkmark & \\
+8 & Canada & \checkmark & \checkmark & \checkmark & & & \\
+9 & India & \checkmark & \checkmark & \checkmark & \checkmark & &
+\checkmark \\
+10 & Italy & \checkmark & \checkmark & \checkmark & & \checkmark & \\
+11 & Jamaica & \checkmark & \checkmark & \checkmark & \checkmark & & \\
+12 & Russia & \checkmark & \checkmark & \checkmark & & & \\
+13 & Sri Lanka & \checkmark & \checkmark & \checkmark & & \checkmark &
+\checkmark \\
+14 & South Africa & \checkmark & \checkmark & \checkmark & \checkmark &
+\checkmark & \\
+15 & German & \checkmark & \checkmark & \checkmark & \checkmark & & \\
+\bottomrule
+\end{longtable}
+
+As shown in table 02 all dashboards which are considered in this paper
+represent the data related to COVID-19 confirmed cases, recovered cases
+\& deaths. There have been contained vaccination details, 8 dashboards
+out of 15 dashboards. In almost each \& every dashboard, value boxes
+have been used to represent these total figures. Bar charts \& line
+charts (trend lines) are most frequently used to visualize these data
+(confirmed cases, recovered cases, deaths \& vaccination details) with
+respect to time. A Majority of dashboards presented data by daily or
+weekly. The mapping is used to track the spatial distribution of
+COVID-19 cases by country/ provincial/ regional etc. When visualizing
+the data by map color code system \& circles with respect to the size of
+the cases have been used to visualize the variation in size. The
+considerable number of dashboards has been used doughnut shape pie
+charts to represent total COVID-19 confirmed cases, recovered cases,
+active cases \& deaths as a percentage. Also, region, gender, age group
+\& ethnicity can be identified as common breakdowns of COVID-19 cases.
+In some dashboard has been added data tables for representing cases
+distribution by province/region. Only very few dashboards have been
+visualized in COVID-19 test details. Only 6 dashboards have been
+compared to global situations. In addition, fatality rate, incidence
+rate, ICU beds, stage of the patients \& hospitalize details have been
+contained in the several dashboards. These are the most common data
+types that have been contained in the dashboards. The main tools that
+have been used for different purpose can be shown in the following
+table.\hfill\break
+
+\textbf{Table 03: Summary of tools which are used for different purpose}
+
+\begin{longtable}[]{@{}lccccccc@{}}
+\toprule
+Purpose & Bar chart & Line chart & Pie chart & Dot plot & Heat map &
+Mapping & Data table \\
+\midrule
+\endhead
+COVID-19 confirmed & & & & & & & \\
+cases & \checkmark & \checkmark & & \checkmark & & \checkmark &
+\checkmark \\
+COVID-19 deaths & \checkmark & \checkmark & & & & \checkmark &
+\checkmark \\
+COVID-19 recovered & & & & & & & \\
+cases & \checkmark & \checkmark & & & & \checkmark & \checkmark \\
+COVID-19 vaccination & & \checkmark & & & & \checkmark & \checkmark \\
+COVID-19 test & & & & & & & \\
+conducted & \checkmark & \checkmark & & & & & \\
+Clinical status & \checkmark & & & & & & \\
+Cases distribution by & & & & & & & \\
+age & \checkmark & & \checkmark & & & & \\
+Cases distribution by & & & & & & & \\
+gender & \checkmark & & & & & & \\
+Cases distribution by area & & & & & & & \\
+(Province/state/region) & \checkmark & \checkmark & & & \checkmark &
+\checkmark & \checkmark \\
+To compare the cases & & & \checkmark & & & & \checkmark \\
+Global comparison & \checkmark & \checkmark & & & & \checkmark &
+\checkmark \\
+\bottomrule
+\end{longtable}
+
+\hypertarget{comparison-of-dashboards}{%
+\subsubsection{2.2 Comparison of
+Dashboards}\label{comparison-of-dashboards}}
+
+Before developing a dashboard, it is necessary to think about which
+visualization tools \& features that should be contained in the
+dashboards. What are the most suitable plots, how many panels in the
+dashboard, what data should be included, how to fit dashboard on a
+screen, colors \& is it real time updated or not are the common things
+that should be considered before develop the dashboards. In table 02 has
+been compared these visualization tools \& features under the following
+categories.
+
+\begin{itemize}
+\tightlist
+\item
+  Number of panels - How many panels which are included in the
+  dashboard.
+\item
+  Visualization tools -- what are the graphical representations of data
+  which are contained in the dashboards.
+\item
+  Fitted on a single screen -- whether the dashboard fits on a single
+  screen or not (users can see the whole dashboard on a single screen
+  without adjusting through grid overlay or not).
+\item
+  Color theme -- is there a unique color used for one data type in the
+  whole dashboard (i.e.: one color scale for one data type everywhere on
+  the dashboard).
+\item
+  Dark background -- background color of the dashboard is dark or light.
+\item
+  Data available -- whether users can be downloaded/available the data
+  set which has been visualized on the dashboard.
+\item
+  Real time updated -- whether the dashboard is updated daily/ specific
+  time (live dashboard) or not.
+\end{itemize}
+
+\textbf{Table 04: Comparison of visualization tools \& features of
+dashboard}
+
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.12}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.12}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.12}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.12}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.12}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.12}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.12}}
+  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.12}}@{}}
+\toprule
+Name of the Dashboard & Number of panels & Visualization tools & Fitted
+on a single screen & Color theme & Dark background & Data available &
+Real time updated \\
+\midrule
+\endhead
+1 & 1 & Bar chart\hfill\break  Interactive map\hfill\break & \checkmark
+& \checkmark & \checkmark & \checkmark & \checkmark \\
+2 & 4 & Line chart\hfill\break  Interactive map\hfill\break  Data
+table\hfill\break & & \checkmark & & \checkmark & \checkmark \\
+3 & 2 & Line chart\hfill\break  Bar chart\hfill\break Interactive
+map\hfill\break Data table\hfill\break & \checkmark & \checkmark &
+\checkmark & \checkmark & \checkmark \\
+4 & 1 & Line chart\hfill\break Bar chart\hfill\break  Pie
+chart\hfill\break Interactive map\hfill\break & \checkmark & &
+\checkmark & \checkmark & \checkmark \\
+5 & 1 & Line chart\hfill\break Bar chart\hfill\break Data
+table\hfill\break & & \checkmark & & \checkmark & \checkmark \\
+6 & 5 & Line chart\hfill\break Bar chart\hfill\break Dot
+plot\hfill\break  Interactive country map\hfill\break & & & & \checkmark
+& \checkmark \\
+7 & 1 & Line chart\hfill\break Bar chart\hfill\break  Country
+map\hfill\break Data table\hfill\break & & \checkmark & & &
+\checkmark \\
+8 & 3 & Line chart\hfill\break Bar chart\hfill\break Data
+table\hfill\break Interactive map\hfill\break & & & \checkmark & &
+\checkmark \\
+9 & 3 & Line chart\hfill\break Bar chart\hfill\break Doughnut shape pie
+chart\hfill\break Data table\hfill\break Interactive country
+map\hfill\break & & \checkmark & & \checkmark & \checkmark \\
+10 & 2 & Bar chart\hfill\break Doughnut shape pie chart\hfill\break Heat
+map\hfill\break Interactive country map\hfill\break & & & & &
+\checkmark \\
+11 & 1 & Line chart\hfill\break Bar chart\hfill\break Doughnut shape pie
+chart\hfill\break Data table\hfill\break Interactive country
+map\hfill\break & & \checkmark & & & \checkmark \\
+12 & 1 & Line chart\hfill\break Bar chart\hfill\break  Interactive
+map\hfill\break & & \checkmark & \checkmark & & \checkmark \\
+13 & 1 & Line chart\hfill\break Bar chart\hfill\break Doughnut shape pie
+chart\hfill\break & & \checkmark & & & \checkmark \\
+14 & 2 & Line chart\hfill\break Bar chart\hfill\break  Interactive
+Country map\hfill\break & & \checkmark & & \checkmark & \checkmark \\
+15 & 1 & Line chart\hfill\break Bar chart\hfill\break  Data
+table\hfill\break Interactive map\hfill\break & \checkmark & \checkmark
+& \checkmark & & \checkmark \\
+\bottomrule
+\end{longtable}
+
+As shown in table 04 almost each \& every dashboard, line charts \& bar
+charts have been used to visualize the data. Heat map \& dot plot has
+been used only one dashboard. Only four dashboards have been fitted with
+a single screen. The majority of dashboards have used color theme on the
+whole dashboard. That is, dashboards have been applied different colors
+for different type of data (i.e.~One specific color for confirmed cases,
+another color for deaths, etc.) in the whole dashboard. The data set \&
+related links have available on some dashboards \& users can download
+these data sets. There are 6 dashboards with dark background while
+others have been used light background. Last updated time \& date of the
+latest available data has been reported at the top or bottom of the
+first panel in the dashboard. Like, half of dashboards included all data
+in one panel.
 
@@
 
-\textbf{Table 01: Description of the panels}
+\textbf{Table 05: Description of the panels}
 
@@
 
+Bar charts \& line charts are the most frequently used tools for the
+visualization of total cases, daily cases \& comparisons with respect to
+time. Some dashboards contained doughnut shape pie charts to summarize
+the total figures. In almost each and every dashboard, value boxes have
+been used to represent total figures. Some dashboards contained
+interactive maps \& data tables to visualize the distribution of cases
+by country, province, region or state. All dashboards are daily updating
+real time dashboards. Gender, age groups and ethnicity can be identified
+as common breakdowns. The data sets \& related links are available on
+most of the dashboards \& can be downloaded. It can be seen that it is
+very easy, clear and user friendly to identify confirmed, recovered and
+deceased cases in dashboards which include one color theme for the whole
+dashboard. Dashboards with the dark background are more comfortable to
+the eyes than dashboards with light background \& light colors. Also, it
+is better if the dashboard can be fitted on a single screen rather than
+adjusting through a grid overlay.
+
+\hypertarget{appendix-overview-of-reviewed-dashboard}{%
+\section{Appendix: Overview of reviewed
+dashboard}\label{appendix-overview-of-reviewed-dashboard}}
+
+Real time updated COVID-19 dashboard created by the John Hopkins
+University Center for Systems Science \& Engineering (JHU CSSE) has
+included total confirmed cases \& total deaths by country,
+province/region/state (not for all countries){[}6{]}. Weekly \& daily
+global confirmed cases, deaths \& vaccine doses have been visualized
+using bar plots related to the date in the side panels. Also, an
+interactive world map has included in the dashboard to represent total
+confirmed cases, incidence rate, fatality ratio \& vaccine doses
+administered by country \& state of US. All the links related to the
+data sets has been included in the dashboard. In this dashboard red,
+white, green colors has been used to indicate confirmed cases, deaths \&
+vaccine doses respectively. The dashboard has fitted on a single screen.
+
+\begin{figure}
+\includegraphics[width=8.42in]{Images/1} \caption{Johns Hopkins University COVID-19 Dashboard}\label{fig:unnamed-chunk-1}
+\end{figure}
+
+Live updated ``WHO COVID-19 dashboard'' contained four panels. The
+interactive world map in the dashboard has visualized the distribution
+of COVID-19 confirmed cases, deaths \& vaccination by country {[}15{]}.
+Line chart \& bar charts have been used as visualization plots. Blue,
+red, green colors have been used to represent confirmed cases, deaths \&
+vaccination respectively. The links related to the underlying data was
+available in the dashboard. The dashboard was not fitted on a single
+screen.
+
+\begin{figure}
+\includegraphics[width=8.35in]{Images/2} \caption{WHO COVID-19 Dashboard}\label{fig:unnamed-chunk-2}
+\end{figure}
+
+The ``COVID-19 surveillance dashboard'' created by the University of
+Virginia in collaboration with Bio complexity Institute has included
+interactive world map with a time slide to visualize the confirmed
+cases, recovered cases, deaths, active cases \& vaccine details {[}7{]}.
+Line charts, bar charts \& data tables have been visualized in the
+dashboard. Red, blue, green \& yellow colors have been used to identify
+confirmed, deceased, recovered \& active cases respectively. The data
+behind the dashboard also available on the dashboard \& can be
+downloaded. The dashboard has fitted on a single screen.
+
+\begin{figure}
+\includegraphics[width=8.42in]{Images/3} \caption{COVID-19 Surveillance Dashboard}\label{fig:unnamed-chunk-3}
+\end{figure}
+
+Corona cases (COVID-19) per municipality in Belgium dashboard has
+included interactive map for distribution of total confirmed cases per
+municipality, bar chart for number of cases per province, line charts
+for number of cases per day by region \& municipality, dodge bar chart
+for deaths per region, per age \& pie chart for deaths per age as
+visualization methods{[}3{]}. Also it included several value boxes for
+total cases, total deaths \& hospital situations. There was no specific
+color theme used in this dashboard. This dashboard also fitted on a
+single screen.
+
+\begin{figure}
+\includegraphics[width=8.39in]{Images/4} \caption{Corona cases per Municipality in Belgium Dashboard}\label{fig:unnamed-chunk-4}
+\end{figure}
+
+The COVID-19 dashboard for England created by NHS providers included
+daily totals, daily changes \& weekly changes in new cases, hospital
+admission, patients in hospital, and patients in ICU beds, deaths \&
+vaccinations in England as counts {[}2{]}. Bar plots for daily new
+cases, hospital admissions \& deaths for last 30 days have been added in
+this dashboard. Also it has included data table for total count, daily
+changes (as increase or decrease), weekly changes (as increase or
+decrease) of new cases \& hospital admission by region. It can be seen
+that, three different colors have been used in bar plot as orange for
+new cases, green for hospital admissions \& red for deaths. These plots
+were not interactive plots \& there was only one panel. The dashboard
+was not fitted on a single screen.
+
+\begin{figure}
+\includegraphics[width=8.46in]{Images/5} \caption{COVID-19 dashboard for England}\label{fig:unnamed-chunk-5}
+\end{figure}
+
+The Environmental Science and Research Institute COVID-19 dashboard for
+New Zealand (was developed \& is maintained in collaboration with
+Epi-interactive Ltd) contained five panels as overview, outbreak,
+source, international \& ESR reports {[}10{]}. In this real time updated
+dashboard, they have included value boxes for total confirmed cases,
+recovered cases, deaths \& interactive map has been added for
+distribution of confirmed \& probable cases as incidence \& count. Also
+they have included bar charts for confirmed cases distribution by
+gender, age group, \& line charts for confirmed \& deaths as daily \&
+cumulative. The dashboard has not fitted on a single screen.
+
+\begin{figure}
+\includegraphics[width=8.47in]{Images/6} \caption{New Zealand COVID-19 Dashboard}\label{fig:unnamed-chunk-6}
+\end{figure}
+
+Pakistan's official COVID-19 dashboard is a real time updated dashboard
+with one panel. This dashboard contains value boxes for total confirmed
+cases, active cases, deaths, recovered cases by country. Province wise
+cases have been represented by map \& data table {[}11{]}. Line charts
+\& bar charts have been added to visualize the distribution of cases
+related to time \& province. In the value boxes blue, orange, pink \&
+green colors have been applied for confirmed cases, active cases, deaths
+\& recoveries respectively, but there was no specific color theme used
+when plotting the graphs. The dashboard cannot be seen on a single
+screen without grid overlay.
+
+\begin{figure}
+\includegraphics[width=8.33in]{Images/7} \caption{COVID-19 Dashboard for Pakistan}\label{fig:unnamed-chunk-7}
+\end{figure}
+
+COVID-19 Canada live dashboard has been developed with three panels for
+provincial data, hospital resources \& projections {[}14{]}. This
+dashboard contained interactive map for represented total reported cases
+\& deaths by province. Data tables, line charts, bar charts were the
+most frequently used visualization types of this dashboard. More details
+were presented on this dashboard by province like ventilator counts,
+hospital beds per 100 residents, resource capacity based on critical
+case rate, effect of social distance, forecast of deaths \& cases. There
+was no specific color theme applied for plotting the graphs. The
+dashboard also was not fitted on a single screen.
+
+\begin{figure}
+\includegraphics[width=8.47in]{Images/8} \caption{COVID-19 in Canada dashboard}\label{fig:unnamed-chunk-8}
+\end{figure}
+
+``Zoho analytics COVID-19 live dashboard'' for India contained three
+panels as insight by state, trend analysis \& vaccination {[}1{]}.
+Interactive country map has been used to visualize the distribution of
+confirmed cases, active cases, deaths, recovered cases \& vaccination by
+states in India. Line charts, bar charts \& doughnut shape pie charts
+have been used as other visualization tools. Also this dashboard
+contained data table for overview of COVID-19 in India. This dashboard
+also was not fitted on a single screen.
+
+\begin{figure}
+\includegraphics[width=8.42in]{Images/9} \caption{India COVID-19 Dashboard}\label{fig:unnamed-chunk-9}
+\end{figure}
+
+Italy COVID-19 dashboard created by the COVID-19 task force of the
+department of infection diseases and the IT service Istituto Superiore
+Sanita, contained interactive map, bar charts, doughnut shape pie charts
+\& heat maps as visualization tools {[}4{]}. It provides data related to
+COVID-19 cases, cases distribution by age, clinical status, and weekly
+number of cases notified in Italy by region/province. This dashboard
+also was not fitted on a single screen.
+
+\begin{figure}
+\includegraphics[width=8.12in]{Images/10} \caption{Italy COVID-19 Dashboard}\label{fig:unnamed-chunk-10}
+\end{figure}
+
+Live COVID-19 dashboard developed \& contributed by an amber innovations
+for the government of Jamaica has added interactive country map \& data
+table to represent the COVID- 19 cases by parish {[}9{]}. Line charts \&
+bar charts have been added to the dashboard for cumulative \& daily
+COVID-19 cases. Yellow, green \& red colors have been used to represent
+confirmed, recovered \& deaths respectively. Doughnut shape pie chart
+has been added to represent the distribution of recovered, confirmed \&
+deaths by age group. This dashboard was not fitted on a single screen.
+
+\begin{figure}
+\includegraphics[width=8.35in]{Images/11} \caption{Jamaica COVID-19 Dashboard}\label{fig:unnamed-chunk-11}
+\end{figure}
+
+The GCI COVID-19 dashboard for Russia includes an interactive map, bar
+charts \& line charts to visualize the COVID-19 data {[}13{]}. Yellow,
+green, red \& blue colors have been added to identify the confirmed,
+recovered, deaths \& active case respectively in plotting the charts.
+
+\begin{figure}
+\includegraphics[width=8.25in]{Images/12} \caption{GCI COVID-19 Dashboard for Russia}\label{fig:unnamed-chunk-12}
+\end{figure}
+
+COVID-19 live situation analysis dashboard of Sri Lanka contains total
+\& daily figures in Sri Lanka as counts {[}5{]}. Total cases \& active
+cases have been visualized by using a line chart. Bar charts have been
+used to summarize the daily confirmed cases and recovered cases. As well
+fatality rate, recovery rate \& daily investigations using PCR tests \&
+rapid antigen tests have been compared by using the bar charts. Doughnut
+shape pie chart has been used to summarize the total cases. Blue, green
+\& red colors have been applied to active, recovered \& deaths
+respectively. This dashboard was also not fitted on a single screen.
+
+\begin{figure}
+\includegraphics[width=8.4in]{Images/13} \caption{COVID-19 Live Situational Analysis Dashboard of Sri Lanka}\label{fig:unnamed-chunk-13}
+\end{figure}
+
+COVID-19 live situation analysis dashboard of Sri Lanka contains total
+\& daily figures in Sri Lanka as counts {[}5{]}. Total cases \& active
+cases have been visualized by using a line chart. Bar charts have been
+used to summarize the daily confirmed cases and recovered cases. As well
+fatality rate, recovery rate \& daily investigations using PCR tests \&
+rapid antigen tests have been compared by using the bar charts. Doughnut
+shape pie chart has been used to summarize the total cases. Blue, green
+\& red colors have been applied to active, recovered \& deaths
+respectively. This dashboard was also not fitted on a single screen.
+
+\begin{figure}
+\includegraphics[width=8.46in]{Images/14} \caption{COVID 19 ZA South Africa Dashboard}\label{fig:unnamed-chunk-14}
+\end{figure}
+
+Robert Koch-Institute COVID-19 dashboard for German contains interactive
+map, line charts \& bar charts as visualization tools of COVID-19 data
+{[}12{]}. The dashboard has been fitted on a single screen. This
+dashboard format is similar to the John Hopkins dashboard.
+
+\begin{figure}
+\includegraphics[width=8.33in]{Images/15} \caption{Robert Koch-Institute COVID-19 dashboard for German}\label{fig:unnamed-chunk-15}
+\end{figure}
+
 \hypertarget{references}{%
@@
 
+\leavevmode\hypertarget{ref-jamaica}{}%
+{``{COVID-19 Jamaica - Ministry of Health and Wellness}.''} 2021.
+\url{https://jamcovid19.moh.gov.jm/}.
+
 \leavevmode\hypertarget{ref-sl}{}%
-{``{COVID-19: Live situational Analysis Dashboard of Sri Lanka,}''}
+{``{COVID-19: Live situational Analysis Dashboard of Sri Lanka}.''}
 2022. \url{https:\%0A//hpb.health.gov.lk/covid19-dashboard/}.
+
+\leavevmode\hypertarget{ref-jh}{}%
+{``{COVID-19 Map -- Johns Hopkins Coronavirus Resource Center}.''} 2022.
+\url{https://coronavi\%0Arus.jhu.edu/map.html}.
+
+\leavevmode\hypertarget{ref-nssac}{}%
+{``{COVID-19 Surveillance Dashboard -- NSSAC Research}.''} 2022.
+\url{https://nssac.bii.virg\%0Ainia.edu/covid-19/dashboard/}.
+
+\leavevmode\hypertarget{ref-za}{}%
+{``{COVID-19 ZA Dashboard - Data Studio}.''} 2021.
+\url{https://datastudio.google.com/u/0/\%0Areporting/1b60bdc7-bec7-44c9-ba29-be0e043d8534/page/hrUIB}.
+
+\leavevmode\hypertarget{ref-nz}{}%
+{``{New Zealand COVID-19 Surveillance Dashboard}.''} 2021.
+\url{https://nzcoviddashboard.esr.c\%0Ari.nz/\#!/}.
+
+\leavevmode\hypertarget{ref-pakistan}{}%
+{``{Pakistan's Official COVID-19 Dashboard -- Shifa International
+Hospitals Ltd}.''} 2021.
+\url{https://www.shifa.com.pk/covid-19-pakistan/}.
+
+\leavevmode\hypertarget{ref-rki}{}%
+{``{RKI COVID-19 Germany -- ArcGIS Experience}.''} 2021.
+\url{https://experience.arcgis.co\%0Am/experience/478220a4c454480e823b17327b2bf1d4}.
+
+\leavevmode\hypertarget{ref-russia}{}%
+{``{The Global COVId-19 Index (GCI) -- Russia Dashboard -- PEMANDU
+Associates}.''} 2021. \url{https://covid19.pemandu.org/Russia}.
+
+\leavevmode\hypertarget{ref-canada}{}%
+{``{Track COVID-19 Across Canada Using Our Interactive Dashboards}.''}
+2021.
+\url{https:\%0A//samples.dundas.com/Dashboard/62cef916-5488-4d3b-9c57-204092d01813?e=false\&vo=viewonly}.
+
+\leavevmode\hypertarget{ref-who}{}%
+{``{WHO Coronavirus (COVID-19) Dashboard}.''} 2021.
+\url{https://covid19.who.int/?gcl\%0Aid=CjwKCAjwvuGJBhB1EiwACU1Aia9pAaNLOkBMS2KVlSqXptskt1t46b7_bNv6vpTt_HjvXfrnDSrWxoCE3sQAvD_\%0ABwE}.
 

### 2022-04-20 manuscript_covid19dashboard.Rmd
Updated

@@
 University of Sri Jayewardenepura, Sri Lanka, CO 10230
+
+\hspace{3cm}
 
@@
 # Methodology
+
+## Data
+
+We obtained data from COVID-19 situation reports published by Epidemiology Unit, Ministry of Health Sri Lanka.
+We obtained data corresponds to number of death cases, number of reported cases, number of recovered cases and COVID-19 vaccinated counts in Sri Lanka.
+The data are made available through an open-source R package covid19srilanka [@talagala].
+
+## Software
+
+R software was used for data cleaning and analysis. The flexdashboard [@flexdashboard] package was used to build the data visualization dashboard.
+
+## Data visualization
+
+Inorder, to compare patterns across districts min-max transformation is used to normalize data into 0-1 scale.
 

### 2022-04-20 manuscript_covid19dashboard.tex
Updated

@@
 \section{Methodology}\label{methodology}}
+
+\hypertarget{data}{%
+\subsection{Data}\label{data}}
+
+We obtained data from COVID-19 situation reports published by
+Epidemiology Unit, Ministry of Health Sri Lanka. We obtained data
+corresponds to number of death cases, number of reported cases, number
+of recovered cases and COVID-19 vaccinated counts in Sri Lanka. The data
+are made available through an open-source R package covid19srilanka
+(Talagala 2021).
+
+\hypertarget{software}{%
+\subsection{Software}\label{software}}
+
+R software was used for data cleaning and analysis. The flexdashboard
+(Iannone, Allaire, and Borges 2020) package was used to build the data
+visualization dashboard.
+
+\hypertarget{data-visualization}{%
+\subsection{Data visualization}\label{data-visualization}}
+
+Inorder, to compare patterns across districts min-max transformation is
+used to normalize data into 0-1 scale.
 
@@
 
+\leavevmode\hypertarget{ref-flexdashboard}{}%
+Iannone, Richard, JJ Allaire, and Barbara Borges. 2020.
+\emph{Flexdashboard: R Markdown Format for Flexible Dashboards}.
+\url{https://CRAN.R-project.org/package=flexdashboard}.
+
 \leavevmode\hypertarget{ref-nz}{}%
@@
 \url{https://experience.arcgis.co\%0Am/experience/478220a4c454480e823b17327b2bf1d4}.
+
+\leavevmode\hypertarget{ref-talagala}{}%
+Talagala, Thiyanga S. 2021. \emph{Covid19srilanka: The 2019 Novel
+Coronavirus COVID-19 (2019-nCoV) Data in Sri Lanka}.
+\url{https://github.com/thiyangt/covid19srilanka}.
 

### 2022-04-20 manuscript_covid19dashboard.Rmd
Working on the methodology section

@@
 We obtained data from COVID-19 situation reports published by Epidemiology Unit, Ministry of Health Sri Lanka.
-We obtained data corresponds to number of death cases, number of reported cases, number of recovered cases and COVID-19 vaccinated counts in Sri Lanka.
+We obtained data corresponds to number of death cases, number of hospitalized cases, number of recovered cases and COVID-19 vaccinated counts in Sri Lanka.
 The data are made available through an open-source R package covid19srilanka [@talagala].
@@
 
-R software was used for data cleaning and analysis. The flexdashboard [@flexdashboard] package was used to build the data visualization dashboard.
+R software was used for data cleaning and analysis.
+The flexdashboard [@flexdashboard] package was used to build the data visualization dashboard.
+The initial layout for the dashboard was prepared based on @rami.
 
@@
 
-Inorder, to compare patterns across districts min-max transformation is used to normalize data into 0-1 scale.
+Table 5 provides an overview of methods that have been used to visualize data.
+
+**Table 5: Data visualization approaches used to visualize data**
+
+| **Data**|**Type of graphics**| 
+|:-------------------|:----------------|
+|Daily COVID-19 confirmed          |Time series plots \hfill\break
+|Daily COVID-19 recovered cases by time         |Time series plots \hfill\break
+|Daily COVID-19 death cases by time         |Time series plots \hfill\break
+|Total COVID-19 confirmed cases by time and wave         |Time series plots annotated with vertical lines to denote significant milestones \hfill\break
+|Total COVID-19 death cases by time and wave         | Histogram \hfill\break
+|Distribution of COVID-19 patients by districts        | Tile map, Choropleth maps, Dorling Cartogram, heatmaps \hfill\break
+|Total vaccination by first dose and second dose        | Time series plot \hfill\break
+|Total administrated does by vaccine name       | Stacked bar chart \hfill\break
+|Total administrated does by vaccine name       | Stacked bar chart \hfill\break
+|Total administrated does by vaccine name       | Stacked bar chart \hfill\break
 
@@
 
-**Table 05: Description of the panels**
+**Table 6: Description of the panels**
 
@@
 |About | This panel contains the details about the dashboard.|
+
+Inorder, to compare patterns across districts min-max transformation is used to normalize data into 0-1 scale.
 

### 2022-04-20 manuscript_covid19dashboard.tex
Working on the methodology section

@@
 Epidemiology Unit, Ministry of Health Sri Lanka. We obtained data
-corresponds to number of death cases, number of reported cases, number
-of recovered cases and COVID-19 vaccinated counts in Sri Lanka. The data
-are made available through an open-source R package covid19srilanka
-(Talagala 2021).
+corresponds to number of death cases, number of hospitalized cases,
+number of recovered cases and COVID-19 vaccinated counts in Sri Lanka.
+The data are made available through an open-source R package
+covid19srilanka (Talagala 2021).
 
@@
 (Iannone, Allaire, and Borges 2020) package was used to build the data
-visualization dashboard.
+visualization dashboard. The initial layout for the dashboard was
+prepared based on Krispin (2021).
 
@@
 
-Inorder, to compare patterns across districts min-max transformation is
-used to normalize data into 0-1 scale.
+Table 5 provides an overview of methods that have been used to visualize
+data.
+
+\textbf{Table 5: Data visualization approaches used to visualize data}
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.54}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.46}}@{}}
+\toprule
+\textbf{Data} & \textbf{Type of graphics} \\
+\midrule
+\endhead
+Daily COVID-19 confirmed & Time series plots \hfill\break \\
+Daily COVID-19 recovered cases by time & Time series plots
+\hfill\break \\
+Daily COVID-19 death cases by time & Time series plots \hfill\break \\
+Total COVID-19 confirmed cases by time and wave & Time series plots
+annotated with vertical lines to denote significant milestones
+\hfill\break \\
+Total COVID-19 death cases by time and wave & Histogram \hfill\break \\
+Distribution of COVID-19 patients by districts & Tile map, Choropleth
+maps, Dorling Cartogram, heatmaps \hfill\break \\
+Total vaccination by first dose and second dose & Time series plot
+\hfill\break \\
+Total administrated does by vaccine name & Stacked bar chart
+\hfill\break \\
+Total administrated does by vaccine name & Stacked bar chart
+\hfill\break \\
+Total administrated does by vaccine name & Stacked bar chart
+\hfill\break \\
+\bottomrule
+\end{longtable}
 
@@
 
-\textbf{Table 05: Description of the panels}
+\textbf{Table 6: Description of the panels}
 
@@
 \end{longtable}
+
+Inorder, to compare patterns across districts min-max transformation is
+used to normalize data into 0-1 scale.
 
@@
 
+\leavevmode\hypertarget{ref-rami}{}%
+Krispin, Rami. 2021. {``{The Coronavirus Dashboard}.''}
+\url{https://github.com/RamiKrispin/coronavirus_dashboard}.
+
 \leavevmode\hypertarget{ref-nz}{}%

### 2022-04-20 manuscript_covid19dashboard.Rmd
Updated the paper

--- 2022-04-20 manuscript_covid19dashboard.Rmd
@@
 
-Inorder, to compare patterns across districts min-max transformation is used to normalize data into 0-1 scale.
+Table 5 provides an overview of methods that have been used to visualize data.
+
+**Table 5: Data visualization approaches used to visualize data**
+
+| **Data**|**Type of graphics**| 
+|:-------------------|:----------------|
+|Daily COVID-19 confirmed          |Time series plots \hfill\break
+|Daily COVID-19 recovered cases by time         |Time series plots \hfill\break
+|Daily COVID-19 death cases by time         |Time series plots \hfill\break
+|Total COVID-19 confirmed cases by time and wave         |Time series plots annotated with vertical lines to denote significant milestones \hfill\break
+|Total COVID-19 death cases by time and wave         | Histogram \hfill\break
+|Distribution of COVID-19 patients by districts        | Tile map, Choropleth maps, Dorling Cartogram, heatmaps \hfill\break
+|Total vaccination by first dose and second dose        | Time series plot \hfill\break
+|Total administrated does by vaccine name       | Stacked bar chart \hfill\break
+|Total administrated does by vaccine name       | Stacked bar chart \hfill\break
+|Total administrated does by vaccine name       | Stacked bar chart \hfill\break
+|Comparison of cases with in Sri Lanka with Top 10 countries       | Cummulative cases by time, Log of cummulative cases by time, stacked bar chart \hfill\break
+|Spread of COVID-19 around the world      | Choropleth maps \hfill\break
 

### 2022-04-20 manuscript_covid19dashboard.tex
Updated the paper
Updated paper

@@
 \hfill\break \\
+Comparison of cases with in Sri Lanka with Top 10 countries &
+Cummulative cases by time, Log of cummulative cases by time, stacked bar
+chart \hfill\break \\
+Spread of COVID-19 around the world & Choropleth maps \hfill\break \\
 \bottomrule
@@
 adjusting through a grid overlay.
+
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-belgium}{}%
+{``{COVID 19 Dashboard -- Belgium}.''} 2022.
+\url{https://esribelux.maps.arcgis.com/apps/ops\%0Adashboard/index.html\#/e350724c87af49bb9ce29646f8a42344}.
+
+\leavevmode\hypertarget{ref-indiadash}{}%
+{``{COVID 19 Dashboard India -- ZOHO Analytics -- ZOHO}.''} 2022.
+\url{https://www.zoho.com/covid/india/}.
+
+\leavevmode\hypertarget{ref-nhs}{}%
+{``{COVID 19 Dashboard -- NHS providers}.''} 2022.
+\url{https://nhsproviders.org/topics/covid-\%0A19/tracking-covid-19-data/covid-19-dashboard}.
+
+\leavevmode\hypertarget{ref-italy}{}%
+{``{COVID-19 integrated surveillance data in Italy -- EpiCentro}.''}
+2022.
+\url{https://www.ep\%0Aicentro.iss.it/en/coronavirus/sars-cov-2-dashboard}.
+
+\leavevmode\hypertarget{ref-jamaica}{}%
+{``{COVID-19 Jamaica - Ministry of Health and Wellness}.''} 2021.
+\url{https://jamcovid19.moh.gov.jm/}.
+
+\leavevmode\hypertarget{ref-sl}{}%
+{``{COVID-19: Live situational Analysis Dashboard of Sri Lanka}.''}
+2022. \url{https:\%0A//hpb.health.gov.lk/covid19-dashboard/}.
+
+\leavevmode\hypertarget{ref-jh}{}%
+{``{COVID-19 Map -- Johns Hopkins Coronavirus Resource Center}.''} 2022.
+\url{https://coronavi\%0Arus.jhu.edu/map.html}.
+
+\leavevmode\hypertarget{ref-nssac}{}%
+{``{COVID-19 Surveillance Dashboard -- NSSAC Research}.''} 2022.
+\url{https://nssac.bii.virg\%0Ainia.edu/covid-19/dashboard/}.
+
+\leavevmode\hypertarget{ref-za}{}%
+{``{COVID-19 ZA Dashboard - Data Studio}.''} 2021.
+\url{https://datastudio.google.com/u/0/\%0Areporting/1b60bdc7-bec7-44c9-ba29-be0e043d8534/page/hrUIB}.
+
+\leavevmode\hypertarget{ref-flexdashboard}{}%
+Iannone, Richard, JJ Allaire, and Barbara Borges. 2020.
+\emph{Flexdashboard: R Markdown Format for Flexible Dashboards}.
+\url{https://CRAN.R-project.org/package=flexdashboard}.
+
+\leavevmode\hypertarget{ref-rami}{}%
+Krispin, Rami. 2021. {``{The Coronavirus Dashboard}.''}
+\url{https://github.com/RamiKrispin/coronavirus_dashboard}.
+
+\leavevmode\hypertarget{ref-nz}{}%
+{``{New Zealand COVID-19 Surveillance Dashboard}.''} 2021.
+\url{https://nzcoviddashboard.esr.c\%0Ari.nz/\#!/}.
+
+\leavevmode\hypertarget{ref-pakistan}{}%
+{``{Pakistan's Official COVID-19 Dashboard -- Shifa International
+Hospitals Ltd}.''} 2021.
+\url{https://www.shifa.com.pk/covid-19-pakistan/}.
+
+\leavevmode\hypertarget{ref-rki}{}%
+{``{RKI COVID-19 Germany -- ArcGIS Experience}.''} 2021.
+\url{https://experience.arcgis.co\%0Am/experience/478220a4c454480e823b17327b2bf1d4}.
+
+\leavevmode\hypertarget{ref-talagala}{}%
+Talagala, Thiyanga S. 2021. \emph{Covid19srilanka: The 2019 Novel
+Coronavirus COVID-19 (2019-nCoV) Data in Sri Lanka}.
+\url{https://github.com/thiyangt/covid19srilanka}.
+
+\leavevmode\hypertarget{ref-russia}{}%
+{``{The Global COVId-19 Index (GCI) -- Russia Dashboard -- PEMANDU
+Associates}.''} 2021. \url{https://covid19.pemandu.org/Russia}.
+
+\leavevmode\hypertarget{ref-canada}{}%
+{``{Track COVID-19 Across Canada Using Our Interactive Dashboards}.''}
+2021.
+\url{https:\%0A//samples.dundas.com/Dashboard/62cef916-5488-4d3b-9c57-204092d01813?e=false\&vo=viewonly}.
+
+\leavevmode\hypertarget{ref-who}{}%
+{``{WHO Coronavirus (COVID-19) Dashboard}.''} 2021.
+\url{https://covid19.who.int/region/amro/country/us}.
+
+\end{CSLReferences}
+
+\newpage
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-belgium}{}%
-{``{COVID 19 Dashboard -- Belgium}.''} 2022.
-\url{https://esribelux.maps.arcgis.com/apps/ops\%0Adashboard/index.html\#/e350724c87af49bb9ce29646f8a42344}.
-
-\leavevmode\hypertarget{ref-indiadash}{}%
-{``{COVID 19 Dashboard India -- ZOHO Analytics -- ZOHO}.''} 2022.
-\url{https://www.zoho.com/covid/india/}.
-
-\leavevmode\hypertarget{ref-nhs}{}%
-{``{COVID 19 Dashboard -- NHS providers}.''} 2022.
-\url{https://nhsproviders.org/topics/covid-\%0A19/tracking-covid-19-data/covid-19-dashboard}.
-
-\leavevmode\hypertarget{ref-italy}{}%
-{``{COVID-19 integrated surveillance data in Italy -- EpiCentro}.''}
-2022.
-\url{https://www.ep\%0Aicentro.iss.it/en/coronavirus/sars-cov-2-dashboard}.
-
-\leavevmode\hypertarget{ref-jamaica}{}%
-{``{COVID-19 Jamaica - Ministry of Health and Wellness}.''} 2021.
-\url{https://jamcovid19.moh.gov.jm/}.
-
-\leavevmode\hypertarget{ref-sl}{}%
-{``{COVID-19: Live situational Analysis Dashboard of Sri Lanka}.''}
-2022. \url{https:\%0A//hpb.health.gov.lk/covid19-dashboard/}.
-
-\leavevmode\hypertarget{ref-jh}{}%
-{``{COVID-19 Map -- Johns Hopkins Coronavirus Resource Center}.''} 2022.
-\url{https://coronavi\%0Arus.jhu.edu/map.html}.
-
-\leavevmode\hypertarget{ref-nssac}{}%
-{``{COVID-19 Surveillance Dashboard -- NSSAC Research}.''} 2022.
-\url{https://nssac.bii.virg\%0Ainia.edu/covid-19/dashboard/}.
-
-\leavevmode\hypertarget{ref-za}{}%
-{``{COVID-19 ZA Dashboard - Data Studio}.''} 2021.
-\url{https://datastudio.google.com/u/0/\%0Areporting/1b60bdc7-bec7-44c9-ba29-be0e043d8534/page/hrUIB}.
-
-\leavevmode\hypertarget{ref-flexdashboard}{}%
-Iannone, Richard, JJ Allaire, and Barbara Borges. 2020.
-\emph{Flexdashboard: R Markdown Format for Flexible Dashboards}.
-\url{https://CRAN.R-project.org/package=flexdashboard}.
-
-\leavevmode\hypertarget{ref-rami}{}%
-Krispin, Rami. 2021. {``{The Coronavirus Dashboard}.''}
-\url{https://github.com/RamiKrispin/coronavirus_dashboard}.
-
-\leavevmode\hypertarget{ref-nz}{}%
-{``{New Zealand COVID-19 Surveillance Dashboard}.''} 2021.
-\url{https://nzcoviddashboard.esr.c\%0Ari.nz/\#!/}.
-
-\leavevmode\hypertarget{ref-pakistan}{}%
-{``{Pakistan's Official COVID-19 Dashboard -- Shifa International
-Hospitals Ltd}.''} 2021.
-\url{https://www.shifa.com.pk/covid-19-pakistan/}.
-
-\leavevmode\hypertarget{ref-rki}{}%
-{``{RKI COVID-19 Germany -- ArcGIS Experience}.''} 2021.
-\url{https://experience.arcgis.co\%0Am/experience/478220a4c454480e823b17327b2bf1d4}.
-
-\leavevmode\hypertarget{ref-talagala}{}%
-Talagala, Thiyanga S. 2021. \emph{Covid19srilanka: The 2019 Novel
-Coronavirus COVID-19 (2019-nCoV) Data in Sri Lanka}.
-\url{https://github.com/thiyangt/covid19srilanka}.
-
-\leavevmode\hypertarget{ref-russia}{}%
-{``{The Global COVId-19 Index (GCI) -- Russia Dashboard -- PEMANDU
-Associates}.''} 2021. \url{https://covid19.pemandu.org/Russia}.
-
-\leavevmode\hypertarget{ref-canada}{}%
-{``{Track COVID-19 Across Canada Using Our Interactive Dashboards}.''}
-2021.
-\url{https:\%0A//samples.dundas.com/Dashboard/62cef916-5488-4d3b-9c57-204092d01813?e=false\&vo=viewonly}.
-
-\leavevmode\hypertarget{ref-who}{}%
-{``{WHO Coronavirus (COVID-19) Dashboard}.''} 2021.
-\url{https://covid19.who.int/?gcl\%0Aid=CjwKCAjwvuGJBhB1EiwACU1Aia9pAaNLOkBMS2KVlSqXptskt1t46b7_bNv6vpTt_HjvXfrnDSrWxoCE3sQAvD_\%0ABwE}.
-
-\end{CSLReferences}
-
 \end{document}

### 2022-04-22 manuscript_covid19dashboard.Rmd
Updated paper
Updated manuscript

@@
 
-The COVID-19 pandemic is a global coronavirus illness outbreak that began in 2019 and is being caused by the Severe Acute Respiratory Syndrome Coronavirus 2 (SARS-CoV-2) virus.
-In December 2019, the first COVID-19-infected patient was discovered in Wuhan, China.
-According to Worldmeter data, there were 229835231 confirmed cases, 206515718 recovered cases, 18605877 active cases, and 4713636 deaths worldwide as of September 20, 2021.
-Currently, 99.5 percent of active patients are in a mild state, while 0.5 percent are in a critical state.
+COVID-19 has expanded over the globe, having a significant impact on our daily lives and work.
+In the fight against COVID-19, early responses and timely decisions and actions are critical to save communities and economy worldwide.
+In order to make effective decisions, data is essential.
+Data-driven information gide the decision-making process and also evaluate the effectiveness of strategies taken.
+
+Given such available data, it is important to develop tools for exploratory analysis for
 
@@
 
-\newpage
-
 # Results
@@
 Also, it is better if the dashboard can be fitted on a single screen rather than adjusting through a grid overlay.
+
+# References
+
+<div id="refs"></div>
+
+\newpage
 
@@
 ```
-
-# References

### 2022-04-22 manuscript_covid19dashboard.tex
Updated paper
Updated manuscript

@@
 
-The COVID-19 pandemic is a global coronavirus illness outbreak that
-began in 2019 and is being caused by the Severe Acute Respiratory
-Syndrome Coronavirus 2 (SARS-CoV-2) virus. In December 2019, the first
-COVID-19-infected patient was discovered in Wuhan, China. According to
-Worldmeter data, there were 229835231 confirmed cases, 206515718
-recovered cases, 18605877 active cases, and 4713636 deaths worldwide as
-of September 20, 2021. Currently, 99.5 percent of active patients are in
-a mild state, while 0.5 percent are in a critical state.
+COVID-19 has expanded over the globe, having a significant impact on our
+daily lives and work. In the fight against COVID-19, early responses and
+timely decisions and actions are critical to save communities and
+economy worldwide. In order to make effective decisions, data is
+essential. Data-driven information gide the decision-making process and
+also evaluate the effectiveness of strategies taken.
+
+Given such available data, it is important to develop tools for
+exploratory analysis for
 
@@
 \end{longtable}
-
-\newpage
 

### 2022-04-22 manuscript_covid19dashboard.Rmd
Updated introduction

@@
 
-Given such available data, it is important to develop tools for exploratory analysis for
+In the response to the COVID-19 pandemic, massive amounts of data are being generated.
+Given such available data, it is critical to create tools for exploratory analysis for policymakers, health officials, and the general public.
 

### 2022-04-22 manuscript_covid19dashboard.tex
Updated introduction

@@
 
-Given such available data, it is important to develop tools for
-exploratory analysis for
+In the response to the COVID-19 pandemic, massive amounts of data are
+being generated. Given such available data, it is critical to create
+tools for exploratory analysis for policymakers, health officials, and
+the general public.
 

### 2022-04-22 manuscript_covid19dashboard.Rmd
Updated introduction
Completed introduction

@@
 Given such available data, it is critical to create tools for exploratory analysis for policymakers, health officials, and the general public.
+Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemic's spread and communication.
+The main purpose of this research is to create a dashboard to visualize COVID-19 data in Sri Lanka.
+Our
 
-Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemic's spread and communication.
 There are a plethora of COVID-19 visualization dashboards that have been designed to visualize the pandemic's global and local status.
@@
 
-# Literature Review
+The rest of the paper is organized as follows: the [Section 2](#litreview) of dashboards created using data related to COVID-19 pandemic.
+[Section 3](#methods) provides the methodology; [Section 4](#results) presents results; and [Section 5](#conclusion) offers some concluding remarks.
+
+# Literature Review {#litreview}
 
@@
 
-# Methodology
+# Methodology {#method}
 
@@
 |Total COVID-19 death cases by time and wave         | Histogram \hfill\break
-|Distribution of COVID-19 patients by districts        | Tile map, Choropleth maps, Dorling Cartogram, heatmaps \hfill\break
+|Distribution of COVID-19 patients by districts        | Tree map, Choropleth maps, Dorling Cartogram, heatmaps \hfill\break
 |Total vaccination by first dose and second dose        | Time series plot \hfill\break
@@
 
-# Results
+# Results {#results}
 
@@
 
-# Discussion and Further research
+# Discussion and Further research {#conclusions}
 

### 2022-04-22 manuscript_covid19dashboard.tex
Updated introduction
Completed introduction

@@
 tools for exploratory analysis for policymakers, health officials, and
-the general public.
+the general public. Dashboards are one of the greatest visual
+interpretation methods for tracking the COVID-19 pandemic's spread and
+communication. The main purpose of this research is to create a
+dashboard to visualize COVID-19 data in Sri Lanka. Our
 
-Dashboards are one of the greatest visual interpretation methods for
-tracking the COVID-19 pandemic's spread and communication. There are a
-plethora of COVID-19 visualization dashboards that have been designed to
-visualize the pandemic's global and local status. Different software can
-be used to generate dashboards. We explored 15 dashboards designed to
-visualize COVID-19 data in the global and country levels. First,
-dashboards were compared to identify the various features, visualization
-approaches, and enhancements that should be implemented. Next, we
-developed an interactive dashboard to visualized COVID-19 outbreak in
-Sri Lanka.
+There are a plethora of COVID-19 visualization dashboards that have been
+designed to visualize the pandemic's global and local status. Different
+software can be used to generate dashboards. We explored 15 dashboards
+designed to visualize COVID-19 data in the global and country levels.
+First, dashboards were compared to identify the various features,
+visualization approaches, and enhancements that should be implemented.
+Next, we developed an interactive dashboard to visualized COVID-19
+outbreak in Sri Lanka.
 
-\hypertarget{literature-review}{%
-\section{Literature Review}\label{literature-review}}
+The rest of the paper is organized as follows: the
+\protect\hyperlink{litreview}{Section 2} of dashboards created using
+data related to COVID-19 pandemic. \protect\hyperlink{methods}{Section
+3} provides the methodology; \protect\hyperlink{results}{Section 4}
+presents results; and \protect\hyperlink{conclusion}{Section 5} offers
+some concluding remarks.
+
+\hypertarget{litreview}{%
+\section{Literature Review}\label{litreview}}
 
@@
 
-\hypertarget{methodology}{%
-\section{Methodology}\label{methodology}}
+\hypertarget{method}{%
+\section{Methodology}\label{method}}
 
@@
 Total COVID-19 death cases by time and wave & Histogram \hfill\break \\
-Distribution of COVID-19 patients by districts & Tile map, Choropleth
+Distribution of COVID-19 patients by districts & Tree map, Choropleth
 maps, Dorling Cartogram, heatmaps \hfill\break \\
@@
 
-\hypertarget{discussion-and-further-research}{%
-\section{Discussion and Further
-research}\label{discussion-and-further-research}}
+\hypertarget{conclusions}{%
+\section{Discussion and Further research}\label{conclusions}}
 

### 2022-04-22 manuscript_covid19dashboard.Rmd
Updated paper

@@
 
-University of Sri Jayewardenepura, Sri Lanka, CO10230
+University of Sri Jayewardenepura, Sri Lanka, CO 10230
 
@@
 The rest of the paper is organized as follows: the [Section 2](#litreview) of dashboards created using data related to COVID-19 pandemic.
-[Section 3](#methods) provides the methodology; [Section 4](#results) presents results; and [Section 5](#conclusion) offers some concluding remarks.
+[Section 3](#methods) provides the methodology and basic design concept; [Section 4](#results) presents results; and [Section 5](#conclusion) offers some concluding remarks.
 

### 2022-04-22 manuscript_covid19dashboard.tex
Updated paper

@@
 data related to COVID-19 pandemic. \protect\hyperlink{methods}{Section
-3} provides the methodology; \protect\hyperlink{results}{Section 4}
-presents results; and \protect\hyperlink{conclusion}{Section 5} offers
-some concluding remarks.
+3} provides the methodology and basic design concept;
+\protect\hyperlink{results}{Section 4} presents results; and
+\protect\hyperlink{conclusion}{Section 5} offers some concluding
+remarks.
 

### 2022-04-22 manuscript_covid19dashboard.Rmd
Updated

@@
 
-### Summary of data representation on dashboards
+## Summary of data representation on dashboards
 
@@
 
-### 2.2 Comparison of Dashboards
+## 2.2 Comparison of Dashboards
 
@@
 
-As shown in table 04 almost each & every dashboard, line charts & bar charts have been used to visualize the data.
+As shown in Table 4 almost each & every dashboard, line charts & bar charts have been used to visualize the data.
 Heat map & dot plot has been used only one dashboard.
@@
 
+In this study, we have daily COVID-19 data related to deaths, recovered and confirmed cases in all 25 districts in Sri Lanka.
+This structure generates a multiple time series collection.
+Visualizing these time series data is useful to identify similarities and dissimilarities between districts and their general trends.
+There are two approaches to visualizing these time series: (i) drawing individual time series plots for each district, and (ii) simultaneously plots all time series on a single panel.
+Plotting separate panels for each district is not effective.
+The reason is that it is hard to compare across 25 different panels at once.
+Plotting all time series simultaneously is also not possible due to overlapping time series and scale differences.
+We use heatmaps  to visualize global and local similarities and dissimilarities across districts (@peng2008method) along with Tree Maps, Choropleth Maps, and  Dorling Cartogram.
+The cumulative number of confirmed deaths, and confirmed cases (both on log scale and linear scale) were used to compare the status of Sri Lanka against the top-10 countries affected by COVID-19.
+Furthermore, we use color-blind friendly colour palettes for the graphics.
+To represent qualitative data, a diverging colour palette was used, and to represent numeric variables, a sequential colour theme was used.
 Table 5 provides an overview of methods that have been used to visualize data.

### 2022-04-22 manuscript_covid19dashboard.tex
Updated

@@
 \hypertarget{summary-of-data-representation-on-dashboards}{%
-\subsubsection{Summary of data representation on
+\subsection{Summary of data representation on
 dashboards}\label{summary-of-data-representation-on-dashboards}}
@@
 \hypertarget{comparison-of-dashboards}{%
-\subsubsection{2.2 Comparison of
+\subsection{2.2 Comparison of
 Dashboards}\label{comparison-of-dashboards}}
@@
 
-As shown in table 04 almost each \& every dashboard, line charts \& bar
+As shown in Table 4 almost each \& every dashboard, line charts \& bar
 charts have been used to visualize the data. Heat map \& dot plot has
@@
 
-Table 5 provides an overview of methods that have been used to visualize
-data.
+In this study, we have daily COVID-19 data related to deaths, recovered
+and confirmed cases in all 25 districts in Sri Lanka. This structure
+generates a multiple time series collection. Visualizing these time
+series data is useful to identify similarities and dissimilarities
+between districts and their general trends. There are two approaches to
+visualizing these time series: (i) drawing individual time series plots
+for each district, and (ii) simultaneously plots all time series on a
+single panel. Plotting separate panels for each district is not
+effective. The reason is that it is hard to compare across 25 different
+panels at once. Plotting all time series simultaneously is also not
+possible due to overlapping time series and scale differences. We use
+heatmaps ~to visualize global and local similarities and dissimilarities
+across districts (Peng (2008)) along with Tree Maps, Choropleth Maps,
+and ~Dorling Cartogram. The cumulative number of confirmed deaths, and
+confirmed cases (both on log scale and linear scale) were used to
+compare the status of Sri Lanka against the top-10 countries affected by
+COVID-19. Furthermore, we use color-blind friendly colour palettes for
+the graphics. To represent qualitative data, a diverging colour palette
+was used, and to represent numeric variables, a sequential colour theme
+was used. Table 5 provides an overview of methods that have been used to
+visualize data.
 
@@
 \url{https://www.shifa.com.pk/covid-19-pakistan/}.
+
+\leavevmode\hypertarget{ref-peng2008method}{}%
+Peng, Roger D. 2008. {``A Method for Visualizing Multivariate Time
+Series Data.''}
 

### 2022-05-06 manuscript_covid19dashboard.Rmd
Updated

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+```{r setup, include=FALSE, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+library(ceylon)
+library(tidyverse)
+library(sf)
+#install.packages("cartogram")
+library(cartogram)
 ```
@@
 
+```{r, comment=NA, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of COVID19 Cases by Districts Plotted on a Single Panel"}
+library(covid19srilanka)
+library(tidyverse)
+library(patchwork)
+data("district.wise.cases")
+f1 <- ggplot(district.wise.cases, aes(
+  x=Date, y=Count,  col=District)) + geom_line() + ggtitle("A")
+
+```
+
+```{r, comment=NA, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of COVID19 Cases by Districts Plotted on a Single Panel", fig.height=10, fig.width=10}
+library(covid19srilanka)
+library(tidyverse)
+data("district.wise.cases")
+f2 <- ggplot(district.wise.cases, aes(
+  x=Date, y=Count)) + geom_line() + 
+  facet_wrap(vars(District), scale="free", ncol=5) + ggtitle("B")
+f1/f2
+```
+
+```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
+library(viridis)
+library(patchwork)
+# Get District wise data 
+District_Covid <- district.wise.cases
+District_Covid <-District_Covid[order(as.Date(District_Covid$Date, format="%Y-%m-%d")), ]
+
+# Get total count for each district
+Last_day <- District_Covid %>%
+  filter(Date == max(District_Covid$Date))
+
+Covid <- Last_day %>%
+  mutate(District = case_when(
+    District == "Kurunagala" ~ "Kurunegala",
+    TRUE ~ District
+  ))
+
+Covidnew <- Covid %>%
+  mutate(parents = "Confirmed")
+
+i <- 1:length(District_Covid$District)
+District_Covid$daily <- District_Covid$Count[i+26] - District_Covid$Count[i]
+District_Covid$Date <- as.Date(District_Covid$Date)
+District_Covid <- head(District_Covid, -26)
+
+District_Covid_new <- District_Covid %>%
+  arrange(Count) %>%
+  mutate(District = fct_inorder(factor(District, ordered = TRUE)))
+
+p1 <- ggplot(District_Covid_new, aes(Date, District)) + 
+           geom_tile(aes(fill = daily))+
+           scale_fill_viridis(option = "magma")+ 
+           scale_x_date(date_breaks = '1 week', date_labels = "%d-%b-%Y") + ggtitle("A")
+```
+
+```{r, echo=FALSE,fig.cap="(A) Global View of Distribution of COVID19 Cases by DIstricts, (B) Local View of Distribution of COVID19 Cases by Districts", fig.width=10, fig.height=10}
+
+cov_District <- as.data.frame(District_Covid)
+
+District <- c(District_Covid$District[1:26])
+data <- cov_District
+
+# Apply min-max transformation for each district separately
+covid <- function(data, District){
+  results <- list(1:26)
+  for(i in 1: length(District)){
+    x <- filter(data, District == District[i])
+    j <- 1:(nrow(data)/26)
+    y <- x$daily
+    x$min_max <-(y[j] - min(y))/(max(y) - min(y))
+    results[[i]] <- x
+  }
+  return(results)
+}
+
+Data <- covid(data, District)
+District_data <- do.call(rbind, Data)
+
+District_Covid_new_1 <- District_data %>%
+  arrange(Count) %>%
+  mutate(District = fct_inorder(factor(District, ordered = TRUE)))
+
+p2 <- ggplot(District_Covid_new_1, aes(Date, District)) + geom_tile(aes(fill = min_max)) + scale_fill_viridis(option = "magma") + scale_x_date(date_breaks = '1 week', date_labels = "%d-%b-%Y") + ggtitle("B")
+
+p1/p2
+```
+
+```{r, comment=NA, message=FALSE, warning=FALSE}
+
+Last_day <- District_Covid %>%
+  filter(Date == max(District_Covid$Date))
+
+Covid <- Last_day %>%
+  mutate(District = case_when(
+    District == "Kurunagala" ~ "Kurunegala",
+    TRUE ~ District
+  ))
+
+Covidnew <- Covid %>%
+  mutate(parents = "Confirmed")
+
+SL <- district
+SL$DISTRICT <- str_to_title(SL$DISTRICT)
+colnames(SL) <- c("geometry", "District", "status", "population")
+
+Covid_SL <- left_join(SL, Covid, by = c("District"))
+
+Ampara <- Covid%>%
+  filter(District %in% c("Kalmunai","Ampara" ))
+newcount <- Ampara$Count %>%
+  sum()
+
+Covid_SL <- Covid_SL %>%
+  filter(!(District == "[unknown]")) %>%
+  filter(!(District == "Kalmunai")) %>%
+  mutate(Count = ifelse(District == "Ampara", newcount, Count))
+
+m1 <- ggplot(Covid_SL) + geom_sf(aes(fill = Count, text = paste('District:', District)), show.legend = TRUE) + 
+  scale_fill_viridis()+
+  theme(axis.text.x = element_blank(), 
+        axis.text.y = element_blank(),
+        axis.ticks = element_blank(),
+        rect = element_blank()) + ggtitle("A")
+
+Covid_SL <- Covid_SL[-c(1), ]
+dorl <- cartogram_dorling(Covid_SL, weight = "population")%>%
+  st_as_sf()
+
+m2 <- ggplot(dorl)+
+  geom_sf(data = Covid_SL)+
+  geom_sf(aes(fill = Count, text = paste('District:', District)))+
+   scale_fill_viridis()+
+  theme(axis.text.x = element_blank(), 
+        axis.text.y = element_blank(),
+        axis.ticks = element_blank(),
+        rect = element_blank()) + ggtitle("B")
+
+m1|m2
+
+```
+
 # Discussion and Further research {#conclusions}

### 2022-05-06 manuscript_covid19dashboard.tex
Updated

@@
 
+\begin{figure}
+\centering
+\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{Distribution of COVID19 Cases by Districts Plotted on a Single
+Panel}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{(A) Global View of Distribution of COVID19 Cases by DIstricts,
+(B) Local View of Distribution of COVID19 Cases by Districts}
+\end{figure}
+
+\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-5-1.pdf}
+
 \hypertarget{conclusions}{%
@@
 \begin{figure}
-\includegraphics[width=8.42in]{Images/1} \caption{Johns Hopkins University COVID-19 Dashboard}\label{fig:unnamed-chunk-1}
+\includegraphics[width=8.42in]{Images/1} \caption{Johns Hopkins University COVID-19 Dashboard}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.35in]{Images/2} \caption{WHO COVID-19 Dashboard}\label{fig:unnamed-chunk-2}
+\includegraphics[width=8.35in]{Images/2} \caption{WHO COVID-19 Dashboard}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.42in]{Images/3} \caption{COVID-19 Surveillance Dashboard}\label{fig:unnamed-chunk-3}
+\includegraphics[width=8.42in]{Images/3} \caption{COVID-19 Surveillance Dashboard}\label{fig:unnamed-chunk-8}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.39in]{Images/4} \caption{Corona cases per Municipality in Belgium Dashboard}\label{fig:unnamed-chunk-4}
+\includegraphics[width=8.39in]{Images/4} \caption{Corona cases per Municipality in Belgium Dashboard}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.46in]{Images/5} \caption{COVID-19 dashboard for England}\label{fig:unnamed-chunk-5}
+\includegraphics[width=8.46in]{Images/5} \caption{COVID-19 dashboard for England}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.47in]{Images/6} \caption{New Zealand COVID-19 Dashboard}\label{fig:unnamed-chunk-6}
+\includegraphics[width=8.47in]{Images/6} \caption{New Zealand COVID-19 Dashboard}\label{fig:unnamed-chunk-11}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.33in]{Images/7} \caption{COVID-19 Dashboard for Pakistan}\label{fig:unnamed-chunk-7}
+\includegraphics[width=8.33in]{Images/7} \caption{COVID-19 Dashboard for Pakistan}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.47in]{Images/8} \caption{COVID-19 in Canada dashboard}\label{fig:unnamed-chunk-8}
+\includegraphics[width=8.47in]{Images/8} \caption{COVID-19 in Canada dashboard}\label{fig:unnamed-chunk-13}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.42in]{Images/9} \caption{India COVID-19 Dashboard}\label{fig:unnamed-chunk-9}
+\includegraphics[width=8.42in]{Images/9} \caption{India COVID-19 Dashboard}\label{fig:unnamed-chunk-14}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.12in]{Images/10} \caption{Italy COVID-19 Dashboard}\label{fig:unnamed-chunk-10}
+\includegraphics[width=8.12in]{Images/10} \caption{Italy COVID-19 Dashboard}\label{fig:unnamed-chunk-15}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.35in]{Images/11} \caption{Jamaica COVID-19 Dashboard}\label{fig:unnamed-chunk-11}
+\includegraphics[width=8.35in]{Images/11} \caption{Jamaica COVID-19 Dashboard}\label{fig:unnamed-chunk-16}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.25in]{Images/12} \caption{GCI COVID-19 Dashboard for Russia}\label{fig:unnamed-chunk-12}
+\includegraphics[width=8.25in]{Images/12} \caption{GCI COVID-19 Dashboard for Russia}\label{fig:unnamed-chunk-17}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.4in]{Images/13} \caption{COVID-19 Live Situational Analysis Dashboard of Sri Lanka}\label{fig:unnamed-chunk-13}
+\includegraphics[width=8.4in]{Images/13} \caption{COVID-19 Live Situational Analysis Dashboard of Sri Lanka}\label{fig:unnamed-chunk-18}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.46in]{Images/14} \caption{COVID 19 ZA South Africa Dashboard}\label{fig:unnamed-chunk-14}
+\includegraphics[width=8.46in]{Images/14} \caption{COVID 19 ZA South Africa Dashboard}\label{fig:unnamed-chunk-19}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.33in]{Images/15} \caption{Robert Koch-Institute COVID-19 dashboard for German}\label{fig:unnamed-chunk-15}
+\includegraphics[width=8.33in]{Images/15} \caption{Robert Koch-Institute COVID-19 dashboard for German}\label{fig:unnamed-chunk-20}
 \end{figure}

### 2022-05-06 manuscript_covid19dashboard.Rmd
Updated

@@
         axis.ticks = element_blank(),
-        rect = element_blank()) + ggtitle("A")
+        rect = element_blank()) + ggtitle("A: Choropleth map")
 
@@
         axis.ticks = element_blank(),
-        rect = element_blank()) + ggtitle("B")
+        rect = element_blank()) + ggtitle("B: Dorling cartogram")
 

### 2022-05-08 manuscript_covid19dashboard.Rmd
Updated paper

@@
 ---
-title: "Interactive dashboard to monitor the COVID-19 outbreak"
+title: "Interactive dashboard to monitor the COVID-19 outbreak and Vaccine Administration"
 output: 
@@
 Dashboards are the most often used visualization method for visualizing COVID-19 data and informing the public.
-The main objective of this study is to identify different features, visualization methods & improvements that should be occurred by exploring the existing dashboards and to develop a dashbored to visualize COVID-19 outbreak in Sri Lanka.
 We explored 15 different dashboards.
-The most commonly used visualization methods in dashboard development are bar charts, line charts, and interactive maps.
-Dashboards that fit on a single screen are preferable than others.
 
@@
 In order to make effective decisions, data is essential.
-Data-driven information gide the decision-making process and also evaluate the effectiveness of strategies taken.
+Data-driven information guide the decision-making process and also evaluate the effectiveness of strategies taken.
 
 In the response to the COVID-19 pandemic, massive amounts of data are being generated.
-Given such available data, it is critical to create tools for exploratory analysis for policymakers, health officials, and the general public.
-Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemic's spread and communication.
-The main purpose of this research is to create a dashboard to visualize COVID-19 data in Sri Lanka.
-Our
+Given such available data, it is critical to create tools for exploratory analysis for policy makers, health officials, and the general public.
+Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemics spread and vaccine administration.
+Dashboards allow user to quickly interact with combination of exploratory visualizations and gain quick overview of the data.
+This paper describes the development and implementation of dashboard for COVID-19 outbreak and vaccine administration data  in Sri Lanka.
 
@@
 First, dashboards were compared to identify the various features, visualization approaches, and enhancements that should be implemented.
-Next, we developed an interactive dashboard to visualized COVID-19 outbreak in Sri Lanka.
+Next, we developed an interactive dashboard to visualized COVID-19 outbreak and vaccination information in Sri Lanka.
+This dashboard provides front-line health officers a situational awareness of spread of COVID-19 and status of the vaccination program.
 
@@
 
-Initially 15 dashboards were compared  to determine which data types are most frequently shown in dashboards. The 15 dashboards we used in the study are listed below (Table 1).
+Dashboards are one of the best visual interpretation method of tracking the spread & communication of COVID-19 pandemic.
+Initially 15 dashboards were compared  to determine which data types are most frequently shown in dashboards.
+The 15 dashboards we used in the study are listed below (Table 1).
+We compared dashboards to identify the different features, visualization
+methods and improvements that should be occurred.
 
@@
 
-## Summary of data representation on dashboards
+First, dashboards were compared to determine which data types are most frequently shown in dashboards. The results are shown in Table 02.
 
@@
 
-## Software
+## Design and development
 
@@
 The initial layout for the dashboard was prepared based on @rami.
-
-## Data visualization
-
-In this study, we have daily COVID-19 data related to deaths, recovered and confirmed cases in all 25 districts in Sri Lanka.
-This structure generates a multiple time series collection.
-Visualizing these time series data is useful to identify similarities and dissimilarities between districts and their general trends.
-There are two approaches to visualizing these time series: (i) drawing individual time series plots for each district, and (ii) simultaneously plots all time series on a single panel.
-Plotting separate panels for each district is not effective.
-The reason is that it is hard to compare across 25 different panels at once.
-Plotting all time series simultaneously is also not possible due to overlapping time series and scale differences.
-We use heatmaps  to visualize global and local similarities and dissimilarities across districts (@peng2008method) along with Tree Maps, Choropleth Maps, and  Dorling Cartogram.
-The cumulative number of confirmed deaths, and confirmed cases (both on log scale and linear scale) were used to compare the status of Sri Lanka against the top-10 countries affected by COVID-19.
-Furthermore, we use color-blind friendly colour palettes for the graphics.
+Data visualizations are generated using ggplot2 [@ggplot2] and plotly [@plotly] packages in R.
+we used color-blind friendly colour palettes for the graphics.
 To represent qualitative data, a diverging colour palette was used, and to represent numeric variables, a sequential colour theme was used.
@@
 
+The source code to reprouduce the results are available in a public GitHub repository and can be access online via https://github.com/thiyangt/covid19srilanka
+
 # Results {#results}
 
-The "Sri Lanka COVID-19 Dashboard" provides an overview of the COVID-19 pandemic for Sri Lanka. This dashboard has eight panels as follows.
+The "Sri Lanka COVID-19 Dashboard" provides an overview of the COVID-19 pandemic and administration of vaacine information in Sri Lanka.
+Figure 1 shows an screenshot of the dashboard.
+This dashboard has eight panels as listed in Table 6.
+The deployed dashboard is made accessible to public through https://thiyanga.netlify.app/post/covid19srilankadb/
+
+```{r, echo=FALSE, fig.cap="Overview Panel: Screeshot of the Dashboard", out.width="80%", fig.align='center'}
+knitr::include_graphics("images/dashboard.png")
+```
+
+\newpage
 
@@
 
-Inorder, to compare patterns across districts min-max transformation is used to normalize data into 0-1 scale.
-
 ```{r, comment=NA, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of COVID19 Cases by Districts Plotted on a Single Panel"}
@@
 f1 <- ggplot(district.wise.cases, aes(
-  x=Date, y=Count,  col=District)) + geom_line() + ggtitle("A")
+  x=Date, y=Count,  col=District)) + geom_line() + ggtitle("A") + theme(legend.position="bottom")
 
@@
 
-```{r, comment=NA, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of COVID19 Cases by Districts Plotted on a Single Panel", fig.height=10, fig.width=10}
+```{r, comment=NA, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of COVID19 Cases by Districts: (A) Plotted on a Single Panel, (B) Plotted on Separate Panels.", fig.height=12, fig.width=10}
 library(covid19srilanka)
@@
 
-```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
+```{r heat, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
 library(viridis)
@@
 
-```{r, comment=NA, message=FALSE, warning=FALSE}
+We now describe the novel visualization approaches we included in out dashboard.
+To effectively distribute the vaccine and to support situational awareness and inform policy makers decision making it is important to know the district-wise spread of COVID-19 cases.
+We have daily COVID-19 data related to confirmed cases in all 25 districts in Sri Lanka.
+This structure generates a multiple time series collection.
+Visualizing these time series data is useful to identify similarities and dissimilarities between districts and their general trends.
+There are two approaches to visualizing these time series: (i) drawing individual time series plots for each district (as shown in Figure 2), and (ii) simultaneously plots all time series on a single panel (as shown in Figure 3).
+Plotting all time series simultaneously is also not possible due to overlapping time series and scale differences.
+Plotting separate panels for each district is not effective.
+The reason is that it is hard to compare across 25 different panels at once.
+In order to overcome these problems in multiple time series visualization,  we use heat maps ((@peng2008method)) to visualize global and local similarities and dissimilarities across districts.
+The associated results are shown in Figure 3.
+Here two heatmaps are used to show the global variations (Figure 3: A) and local variation (Figure 3: B) in the time series collection.
+Figure 3-A cell colours represent the actual counts of the COVID-19 confirmed cases.
+This is useful to get an idea about the differences in absolute values.
+Figure 3-B cell colours represent the normalized values created applying the min-max transformation.
+Min-max transformation is applied to each district time series by using the corresponding district minimum and maximum value of the time series.
+This helps us to get an idea about patterns within districts.
+For example, according to Figure 3A we can see Colombo, Gampaha and Kalutara districs COVID-19 cases are significantly higher than other districts.
+According to Figure 3B all districts show an increasing trend pattern as the right-hand side of the cells are lighter than the left-hand side cells in the heat map.
+Furthermore according to Figure 3B all districts reported high number of cases on  19, 24, 29 August, 2021.
+Figure 3B is useful for identifying these local outlying behaviours.
+As shown in Figure 4, we also use Choropleth map and Dorling cartogram to visualize spatial distribution of COVID-19 cases.
+The vaccination information are visualize through interactive time series plots and bar charts.
+A screenshot of associated panels are shown in Figure 5 and Figure 6.
+
+```{r, comment=NA, message=FALSE, warning=FALSE, fig.cap="Spatial DIstribution of COVID1-19 Cases by Districts"}
 
@@
 
+```{r, echo=FALSE, fig.cap="Time Series Plot of Distribution of First Dose and Second Dose", out.width="80%", fig.align='center'}
+knitr::include_graphics("Images/d3.png")
+```
+
+```{r, echo=FALSE, fig.cap="Distribution of Vaccines by Different Types and Dose", out.width="80%", fig.align='center'}
+knitr::include_graphics("Images/d2.png")
+```
+
 # Discussion and Further research {#conclusions}
 
-Bar charts & line charts are the most frequently used tools for the visualization of total cases, daily cases & comparisons with respect to time.
+Bar charts and line charts are the most frequently used tools for the visualization of total cases, daily cases and comparisons with respect to time.
 Some dashboards contained doughnut shape pie charts to summarize the total figures.

### 2022-05-08 manuscript_covid19dashboard.tex
Updated paper

@@
 \hypersetup{
-  pdftitle={Interactive dashboard to monitor the COVID-19 outbreak},
+  pdftitle={Interactive dashboard to monitor the COVID-19 outbreak and Vaccine Administration},
   hidelinks,
@@
 
-\title{Interactive dashboard to monitor the COVID-19 outbreak}
+\title{Interactive dashboard to monitor the COVID-19 outbreak and
+Vaccine Administration}
 \author{}
@@
 visualization method for visualizing COVID-19 data and informing the
-public. The main objective of this study is to identify different
-features, visualization methods \& improvements that should be occurred
-by exploring the existing dashboards and to develop a dashbored to
-visualize COVID-19 outbreak in Sri Lanka. We explored 15 different
-dashboards. The most commonly used visualization methods in dashboard
-development are bar charts, line charts, and interactive maps.
-Dashboards that fit on a single screen are preferable than others.
+public. We explored 15 different dashboards.
 
@@
 economy worldwide. In order to make effective decisions, data is
-essential. Data-driven information gide the decision-making process and
+essential. Data-driven information guide the decision-making process and
 also evaluate the effectiveness of strategies taken.
@@
 being generated. Given such available data, it is critical to create
-tools for exploratory analysis for policymakers, health officials, and
+tools for exploratory analysis for policy makers, health officials, and
 the general public. Dashboards are one of the greatest visual
-interpretation methods for tracking the COVID-19 pandemic's spread and
-communication. The main purpose of this research is to create a
-dashboard to visualize COVID-19 data in Sri Lanka. Our
+interpretation methods for tracking the COVID-19 pandemics spread and
+vaccine administration. Dashboards allow user to quickly interact with
+combination of exploratory visualizations and gain quick overview of the
+data. This paper describes the development and implementation of
+dashboard for COVID-19 outbreak and vaccine administration data in Sri
+Lanka.
 
@@
 Next, we developed an interactive dashboard to visualized COVID-19
-outbreak in Sri Lanka.
+outbreak and vaccination information in Sri Lanka. This dashboard
+provides front-line health officers a situational awareness of spread of
+COVID-19 and status of the vaccination program.
 
@@
 
-Initially 15 dashboards were compared to determine which data types are
-most frequently shown in dashboards. The 15 dashboards we used in the
-study are listed below (Table 1).
+Dashboards are one of the best visual interpretation method of tracking
+the spread \& communication of COVID-19 pandemic. Initially 15
+dashboards were compared to determine which data types are most
+frequently shown in dashboards. The 15 dashboards we used in the study
+are listed below (Table 1). We compared dashboards to identify the
+different features, visualization methods and improvements that should
+be occurred.
 
@@
 
-\hypertarget{summary-of-data-representation-on-dashboards}{%
-\subsection{Summary of data representation on
-dashboards}\label{summary-of-data-representation-on-dashboards}}
+First, dashboards were compared to determine which data types are most
+frequently shown in dashboards. The results are shown in Table 02.
 
@@
 
-\hypertarget{software}{%
-\subsection{Software}\label{software}}
+\hypertarget{design-and-development}{%
+\subsection{Design and development}\label{design-and-development}}
 
@@
 visualization dashboard. The initial layout for the dashboard was
-prepared based on Krispin (2021).
-
-\hypertarget{data-visualization}{%
-\subsection{Data visualization}\label{data-visualization}}
-
-In this study, we have daily COVID-19 data related to deaths, recovered
-and confirmed cases in all 25 districts in Sri Lanka. This structure
-generates a multiple time series collection. Visualizing these time
-series data is useful to identify similarities and dissimilarities
-between districts and their general trends. There are two approaches to
-visualizing these time series: (i) drawing individual time series plots
-for each district, and (ii) simultaneously plots all time series on a
-single panel. Plotting separate panels for each district is not
-effective. The reason is that it is hard to compare across 25 different
-panels at once. Plotting all time series simultaneously is also not
-possible due to overlapping time series and scale differences. We use
-heatmaps ~to visualize global and local similarities and dissimilarities
-across districts (Peng (2008)) along with Tree Maps, Choropleth Maps,
-and ~Dorling Cartogram. The cumulative number of confirmed deaths, and
-confirmed cases (both on log scale and linear scale) were used to
-compare the status of Sri Lanka against the top-10 countries affected by
-COVID-19. Furthermore, we use color-blind friendly colour palettes for
-the graphics. To represent qualitative data, a diverging colour palette
-was used, and to represent numeric variables, a sequential colour theme
-was used. Table 5 provides an overview of methods that have been used to
-visualize data.
+prepared based on Krispin (2021). Data visualizations are generated
+using ggplot2 (\textbf{ggplot2?}) and plotly (Sievert 2020) packages in
+R. we used color-blind friendly colour palettes for the graphics. To
+represent qualitative data, a diverging colour palette was used, and to
+represent numeric variables, a sequential colour theme was used. Table 5
+provides an overview of methods that have been used to visualize data.
 
@@
 
+The source code to reprouduce the results are available in a public
+GitHub repository and can be access online via
+\url{https://github.com/thiyangt/covid19srilanka}
+
 \hypertarget{results}{%
@@
 The ``Sri Lanka COVID-19 Dashboard'' provides an overview of the
-COVID-19 pandemic for Sri Lanka. This dashboard has eight panels as
-follows.
+COVID-19 pandemic and administration of vaacine information in Sri
+Lanka. Figure 1 shows an screenshot of the dashboard. This dashboard has
+eight panels as listed in Table 6. The deployed dashboard is made
+accessible to public through
+\url{https://thiyanga.netlify.app/post/covid19srilankadb/}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{images/dashboard}
+
+}
+
+\caption{Overview Panel: Screeshot of the Dashboard}\label{fig:unnamed-chunk-1}
+\end{figure}
+
+\newpage
 
@@
 
-Inorder, to compare patterns across districts min-max transformation is
-used to normalize data into 0-1 scale.
-
 \begin{figure}
 \centering
-\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{Distribution of COVID19 Cases by Districts Plotted on a Single
-Panel}
+\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{Distribution of COVID19 Cases by Districts: (A) Plotted on a
+Single Panel, (B) Plotted on Separate Panels.}
 \end{figure}
@@
 
+We now describe the novel visualization approaches we included in out
+dashboard. To effectively distribute the vaccine and to support
+situational awareness and inform policy makers decision making it is
+important to know the district-wise spread of COVID-19 cases. We have
+daily COVID-19 data related to confirmed cases in all 25 districts in
+Sri Lanka. This structure generates a multiple time series collection.
+Visualizing these time series data is useful to identify similarities
+and dissimilarities between districts and their general trends. There
+are two approaches to visualizing these time series: (i) drawing
+individual time series plots for each district (as shown in Figure 2),
+and (ii) simultaneously plots all time series on a single panel (as
+shown in Figure 3). Plotting all time series simultaneously is also not
+possible due to overlapping time series and scale differences. Plotting
+separate panels for each district is not effective. The reason is that
+it is hard to compare across 25 different panels at once. In order to
+overcome these problems in multiple time series visualization, we use
+heat maps ((Peng (2008))) to visualize global and local similarities and
+dissimilarities across districts. The associated results are shown in
+Figure 3. Here two heatmaps are used to show the global variations
+(Figure 3: A) and local variation (Figure 3: B) in the time series
+collection. Figure 3-A cell colours represent the actual counts of the
+COVID-19 confirmed cases. This is useful to get an idea about the
+differences in absolute values. Figure 3-B cell colours represent the
+normalized values created applying the min-max transformation. Min-max
+transformation is applied to each district time series by using the
+corresponding district minimum and maximum value of the time series.
+This helps us to get an idea about patterns within districts. For
+example, according to Figure 3A we can see Colombo, Gampaha and Kalutara
+districs COVID-19 cases are significantly higher than other districts.
+According to Figure 3B all districts show an increasing trend pattern as
+the right-hand side of the cells are lighter than the left-hand side
+cells in the heat map. Furthermore according to Figure 3B all districts
+reported high number of cases on 19, 24, 29 August, 2021. Figure 3B is
+useful for identifying these local outlying behaviours. As shown in
+Figure 4, we also use Choropleth map and Dorling cartogram to visualize
+spatial distribution of COVID-19 cases. The vaccination information are
+visualize through interactive time series plots and bar charts. A
+screenshot of associated panels are shown in Figure 5 and Figure 6.
+
+\begin{figure}
+\centering
 \includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{Spatial DIstribution of COVID1-19 Cases by Districts}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{Images/d3}
+
+}
+
+\caption{Time Series Plot of Distribution of First Dose and Second Dose}\label{fig:unnamed-chunk-6}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{Images/d2}
+
+}
+
+\caption{Distribution of Vaccines by Different Types and Dose}\label{fig:unnamed-chunk-7}
+\end{figure}
 
@@
 
-Bar charts \& line charts are the most frequently used tools for the
-visualization of total cases, daily cases \& comparisons with respect to
-time. Some dashboards contained doughnut shape pie charts to summarize
-the total figures. In almost each and every dashboard, value boxes have
-been used to represent total figures. Some dashboards contained
-interactive maps \& data tables to visualize the distribution of cases
-by country, province, region or state. All dashboards are daily updating
-real time dashboards. Gender, age groups and ethnicity can be identified
-as common breakdowns. The data sets \& related links are available on
-most of the dashboards \& can be downloaded. It can be seen that it is
-very easy, clear and user friendly to identify confirmed, recovered and
-deceased cases in dashboards which include one color theme for the whole
-dashboard. Dashboards with the dark background are more comfortable to
-the eyes than dashboards with light background \& light colors. Also, it
-is better if the dashboard can be fitted on a single screen rather than
-adjusting through a grid overlay.
+Bar charts and line charts are the most frequently used tools for the
+visualization of total cases, daily cases and comparisons with respect
+to time. Some dashboards contained doughnut shape pie charts to
+summarize the total figures. In almost each and every dashboard, value
+boxes have been used to represent total figures. Some dashboards
+contained interactive maps \& data tables to visualize the distribution
+of cases by country, province, region or state. All dashboards are daily
+updating real time dashboards. Gender, age groups and ethnicity can be
+identified as common breakdowns. The data sets \& related links are
+available on most of the dashboards \& can be downloaded. It can be seen
+that it is very easy, clear and user friendly to identify confirmed,
+recovered and deceased cases in dashboards which include one color theme
+for the whole dashboard. Dashboards with the dark background are more
+comfortable to the eyes than dashboards with light background \& light
+colors. Also, it is better if the dashboard can be fitted on a single
+screen rather than adjusting through a grid overlay.
 
@@
 
+\leavevmode\hypertarget{ref-plotly}{}%
+Sievert, Carson. 2020. \emph{Interactive Web-Based Data Visualization
+with r, Plotly, and Shiny}. Chapman; Hall/CRC.
+\url{https://plotly-r.com}.
+
 \leavevmode\hypertarget{ref-talagala}{}%
@@
 \begin{figure}
-\includegraphics[width=8.42in]{Images/1} \caption{Johns Hopkins University COVID-19 Dashboard}\label{fig:unnamed-chunk-6}
+\includegraphics[width=8.42in]{Images/1} \caption{Johns Hopkins University COVID-19 Dashboard}\label{fig:unnamed-chunk-8}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.35in]{Images/2} \caption{WHO COVID-19 Dashboard}\label{fig:unnamed-chunk-7}
+\includegraphics[width=8.35in]{Images/2} \caption{WHO COVID-19 Dashboard}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.42in]{Images/3} \caption{COVID-19 Surveillance Dashboard}\label{fig:unnamed-chunk-8}
+\includegraphics[width=8.42in]{Images/3} \caption{COVID-19 Surveillance Dashboard}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.39in]{Images/4} \caption{Corona cases per Municipality in Belgium Dashboard}\label{fig:unnamed-chunk-9}
+\includegraphics[width=8.39in]{Images/4} \caption{Corona cases per Municipality in Belgium Dashboard}\label{fig:unnamed-chunk-11}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.46in]{Images/5} \caption{COVID-19 dashboard for England}\label{fig:unnamed-chunk-10}
+\includegraphics[width=8.46in]{Images/5} \caption{COVID-19 dashboard for England}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.47in]{Images/6} \caption{New Zealand COVID-19 Dashboard}\label{fig:unnamed-chunk-11}
+\includegraphics[width=8.47in]{Images/6} \caption{New Zealand COVID-19 Dashboard}\label{fig:unnamed-chunk-13}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.33in]{Images/7} \caption{COVID-19 Dashboard for Pakistan}\label{fig:unnamed-chunk-12}
+\includegraphics[width=8.33in]{Images/7} \caption{COVID-19 Dashboard for Pakistan}\label{fig:unnamed-chunk-14}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.47in]{Images/8} \caption{COVID-19 in Canada dashboard}\label{fig:unnamed-chunk-13}
+\includegraphics[width=8.47in]{Images/8} \caption{COVID-19 in Canada dashboard}\label{fig:unnamed-chunk-15}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.42in]{Images/9} \caption{India COVID-19 Dashboard}\label{fig:unnamed-chunk-14}
+\includegraphics[width=8.42in]{Images/9} \caption{India COVID-19 Dashboard}\label{fig:unnamed-chunk-16}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.12in]{Images/10} \caption{Italy COVID-19 Dashboard}\label{fig:unnamed-chunk-15}
+\includegraphics[width=8.12in]{Images/10} \caption{Italy COVID-19 Dashboard}\label{fig:unnamed-chunk-17}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.35in]{Images/11} \caption{Jamaica COVID-19 Dashboard}\label{fig:unnamed-chunk-16}
+\includegraphics[width=8.35in]{Images/11} \caption{Jamaica COVID-19 Dashboard}\label{fig:unnamed-chunk-18}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.25in]{Images/12} \caption{GCI COVID-19 Dashboard for Russia}\label{fig:unnamed-chunk-17}
+\includegraphics[width=8.25in]{Images/12} \caption{GCI COVID-19 Dashboard for Russia}\label{fig:unnamed-chunk-19}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.4in]{Images/13} \caption{COVID-19 Live Situational Analysis Dashboard of Sri Lanka}\label{fig:unnamed-chunk-18}
+\includegraphics[width=8.4in]{Images/13} \caption{COVID-19 Live Situational Analysis Dashboard of Sri Lanka}\label{fig:unnamed-chunk-20}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.46in]{Images/14} \caption{COVID 19 ZA South Africa Dashboard}\label{fig:unnamed-chunk-19}
+\includegraphics[width=8.46in]{Images/14} \caption{COVID 19 ZA South Africa Dashboard}\label{fig:unnamed-chunk-21}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=8.33in]{Images/15} \caption{Robert Koch-Institute COVID-19 dashboard for German}\label{fig:unnamed-chunk-20}
+\includegraphics[width=8.33in]{Images/15} \caption{Robert Koch-Institute COVID-19 dashboard for German}\label{fig:unnamed-chunk-22}
 \end{figure}

### 2022-05-08 manuscript_covid19dashboard.Rmd
Completed introduction

@@
 COVID-19 has expanded over the globe, having a significant impact on our daily lives and work.
-In the fight against COVID-19, early responses and timely decisions and actions are critical to save communities and economy worldwide.
-In order to make effective decisions, data is essential.
-Data-driven information guide the decision-making process and also evaluate the effectiveness of strategies taken.
+Early responses and timely decisions and actions are critical to saving communities and economy worldwide.
+Data is essential in order to make effective decisions.
+Data-driven information guides the decision-making process and also evaluates the effectiveness of strategies taken.
 
-In the response to the COVID-19 pandemic, massive amounts of data are being generated.
-Given such available data, it is critical to create tools for exploratory analysis for policy makers, health officials, and the general public.
-Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemics spread and vaccine administration.
-Dashboards allow user to quickly interact with combination of exploratory visualizations and gain quick overview of the data.
-This paper describes the development and implementation of dashboard for COVID-19 outbreak and vaccine administration data  in Sri Lanka.
+Massive amounts of data are being generated in the response to the COVID-19 pandemic.
+Given this available data, it is critical to create tools for exploratory analysis for policymakers, health officials, and the general public.
+Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemic's spread and vaccine administration.
+Dashboards allow users to quickly interact with a combination of exploratory visualizations and gain a quick overview of the data.
+This paper describes the development and implementation of a dashboard for the COVID-19 outbreak and vaccine administration data in Sri Lanka.
 
@@
 Different software can be used to generate dashboards.
-We explored 15 dashboards designed to visualize COVID-19 data in the global and country levels.
+We explored 15 dashboards designed to visualize COVID-19 data at the global and country levels.
 First, dashboards were compared to identify the various features, visualization approaches, and enhancements that should be implemented.
-Next, we developed an interactive dashboard to visualized COVID-19 outbreak and vaccination information in Sri Lanka.
-This dashboard provides front-line health officers a situational awareness of spread of COVID-19 and status of the vaccination program.
+Next, we developed an interactive dashboard to visualize the COVID-19 outbreak and vaccination information in Sri Lanka.
+This dashboard provides front-line health officers a situational awareness of the spread of COVID-19 and status of the vaccination program.
 
-The rest of the paper is organized as follows: the [Section 2](#litreview) of dashboards created using data related to COVID-19 pandemic.
-[Section 3](#methods) provides the methodology and basic design concept; [Section 4](#results) presents results; and [Section 5](#conclusion) offers some concluding remarks.
+The rest of the paper is organized as follows: [Section 2](#litreview) of dashboards created using data related to the COVID-19 pandemic.
+[Section 3](#methods) presents the methodology and basic design concept; [Section 4](#results) presents the results; and [Section 5](# conclusion) concludes.
 

### 2022-05-08 manuscript_covid19dashboard.tex
Completed introduction

@@
 COVID-19 has expanded over the globe, having a significant impact on our
-daily lives and work. In the fight against COVID-19, early responses and
-timely decisions and actions are critical to save communities and
-economy worldwide. In order to make effective decisions, data is
-essential. Data-driven information guide the decision-making process and
-also evaluate the effectiveness of strategies taken.
+daily lives and work. Early responses and timely decisions and actions
+are critical to saving communities and economy worldwide. Data is
+essential in order to make effective decisions. Data-driven information
+guides the decision-making process and also evaluates the effectiveness
+of strategies taken.
 
-In the response to the COVID-19 pandemic, massive amounts of data are
-being generated. Given such available data, it is critical to create
-tools for exploratory analysis for policy makers, health officials, and
+Massive amounts of data are being generated in the response to the
+COVID-19 pandemic. Given this available data, it is critical to create
+tools for exploratory analysis for policymakers, health officials, and
 the general public. Dashboards are one of the greatest visual
-interpretation methods for tracking the COVID-19 pandemics spread and
-vaccine administration. Dashboards allow user to quickly interact with
-combination of exploratory visualizations and gain quick overview of the
-data. This paper describes the development and implementation of
-dashboard for COVID-19 outbreak and vaccine administration data in Sri
-Lanka.
+interpretation methods for tracking the COVID-19 pandemic's spread and
+vaccine administration. Dashboards allow users to quickly interact with
+a combination of exploratory visualizations and gain a quick overview of
+the data. This paper describes the development and implementation of a
+dashboard for the COVID-19 outbreak and vaccine administration data in
+Sri Lanka.
 
@@
 software can be used to generate dashboards. We explored 15 dashboards
-designed to visualize COVID-19 data in the global and country levels.
+designed to visualize COVID-19 data at the global and country levels.
 First, dashboards were compared to identify the various features,
 visualization approaches, and enhancements that should be implemented.
-Next, we developed an interactive dashboard to visualized COVID-19
+Next, we developed an interactive dashboard to visualize the COVID-19
 outbreak and vaccination information in Sri Lanka. This dashboard
-provides front-line health officers a situational awareness of spread of
-COVID-19 and status of the vaccination program.
+provides front-line health officers a situational awareness of the
+spread of COVID-19 and status of the vaccination program.
 
-The rest of the paper is organized as follows: the
+The rest of the paper is organized as follows:
 \protect\hyperlink{litreview}{Section 2} of dashboards created using
-data related to COVID-19 pandemic. \protect\hyperlink{methods}{Section
-3} provides the methodology and basic design concept;
-\protect\hyperlink{results}{Section 4} presents results; and
-\protect\hyperlink{conclusion}{Section 5} offers some concluding
-remarks.
+data related to the COVID-19 pandemic.
+\protect\hyperlink{methods}{Section 3} presents the methodology and
+basic design concept; \protect\hyperlink{results}{Section 4} presents
+the results; and \protect\hyperlink{ux5cux2520conclusion}{Section 5}
+concludes.
 

### 2022-05-08 manuscript_covid19dashboard.Rmd
Updated literature review

@@
 
-Dashboards are one of the best visual interpretation method of tracking the spread & communication of COVID-19 pandemic.
-Initially 15 dashboards were compared  to determine which data types are most frequently shown in dashboards.
-The 15 dashboards we used in the study are listed below (Table 1).
-We compared dashboards to identify the different features, visualization
-methods and improvements that should be occurred.
+Dashboards are one of the best visual interpretation methods for tracking the spread and communication of the COVID-19 pandemic.
+The 15 dashboards we used in the literature survey are listed in Table 1.
+We compared dashboards to identify data types, plotting techniques, colour themes, and other features such as interactivity on plots and panel numbers.
 
@@
 
-First, dashboards were compared to determine which data types are most frequently shown in dashboards. The results are shown in Table 02.
+Table 02 summarizes the data types that are most frequently shown in dashboards.
+As shown in Table 02, all dashboards which are considered in this paper represent the data related to COVID-19 confirmed cases, recovered cases, and deaths.
+There were 8 dashboards out of 15 dashboards that contained vaccination details.
 
@@
 
-As shown in table 02 all dashboards which are considered in this paper represent the data related to COVID-19 confirmed cases, recovered cases & deaths.
-There have been contained vaccination details, 8 dashboards out of 15 dashboards.
-In almost each & every dashboard, value boxes have been used to represent these total figures.
-Bar charts & line charts (trend lines) are most frequently used to visualize these data (confirmed cases, recovered cases, deaths & vaccination details) with respect to time.
-A Majority of dashboards presented data by daily or weekly.
-The mapping is used to track the spatial distribution of COVID-19 cases by country/ provincial/ regional etc.
-When visualizing the data by map color code system & circles with respect to the size of the cases have been used to visualize the variation in size.
-The considerable number of dashboards has been used doughnut shape pie charts to represent total COVID-19 confirmed cases, recovered cases, active cases & deaths as a percentage.
-Also, region, gender, age group & ethnicity can be identified as common breakdowns of COVID-19 cases.
-In some dashboard has been added data tables for representing cases distribution by province/region.
-Only very few dashboards have been visualized in COVID-19 test details.
+Table 03 summarizes the visualization methods used in the dashboards.
+In almost every dashboard, value boxes have been used to represent these total figures.
+Bar charts and line charts (trend lines) are most frequently used to visualize this data (confirmed cases, recovered cases, deaths, and vaccination details) with respect to time.
+A majority of dashboards presented data either daily or weekly.
+The mapping is used to track the spatial distribution of COVID-19 cases by country, province, regional etc.
+When visualizing the data by the map color code system, circles with respect to the size of the cases have been used to visualize the variation in size.
+The considerable number of dashboards has been used in doughnut-shaped pie charts to represent total COVID-19 confirmed cases, recovered cases, active cases, and deaths as a percentage.
+Also, region, gender, age group, and ethnicity can be identified as common breakdowns of COVID-19 cases.
+Data tables for representing case distribution by province/region have been added to some dashboards.
+Very few dashboards have been visualized in the COVID-19 test details.
 Only 6 dashboards have been compared to global situations.
-In addition, fatality rate, incidence rate, ICU beds, stage of the patients & hospitalize details have been contained in the several dashboards.
-These are the most common data types that have been contained in the dashboards.
-The main tools that have been used for different purpose can be shown in the following table.\hfill\break
+In addition, the fatality rate, incidence rate, ICU beds, stage of the patients, and hospitalized details have been contained in the several dashboards.\hfill\break
 

### 2022-05-08 manuscript_covid19dashboard.tex
Updated literature review

@@
 
-Dashboards are one of the best visual interpretation method of tracking
-the spread \& communication of COVID-19 pandemic. Initially 15
-dashboards were compared to determine which data types are most
-frequently shown in dashboards. The 15 dashboards we used in the study
-are listed below (Table 1). We compared dashboards to identify the
-different features, visualization methods and improvements that should
-be occurred.
+Dashboards are one of the best visual interpretation methods for
+tracking the spread and communication of the COVID-19 pandemic. The 15
+dashboards we used in the literature survey are listed in Table 1. We
+compared dashboards to identify data types, plotting techniques, colour
+themes, and other features such as interactivity on plots and panel
+numbers.
 
@@
 
-First, dashboards were compared to determine which data types are most
-frequently shown in dashboards. The results are shown in Table 02.
+Table 02 summarizes the data types that are most frequently shown in
+dashboards. As shown in Table 02, all dashboards which are considered in
+this paper represent the data related to COVID-19 confirmed cases,
+recovered cases, and deaths. There were 8 dashboards out of 15
+dashboards that contained vaccination details.
 
@@
 
-As shown in table 02 all dashboards which are considered in this paper
-represent the data related to COVID-19 confirmed cases, recovered cases
-\& deaths. There have been contained vaccination details, 8 dashboards
-out of 15 dashboards. In almost each \& every dashboard, value boxes
-have been used to represent these total figures. Bar charts \& line
-charts (trend lines) are most frequently used to visualize these data
-(confirmed cases, recovered cases, deaths \& vaccination details) with
-respect to time. A Majority of dashboards presented data by daily or
-weekly. The mapping is used to track the spatial distribution of
-COVID-19 cases by country/ provincial/ regional etc. When visualizing
-the data by map color code system \& circles with respect to the size of
-the cases have been used to visualize the variation in size. The
-considerable number of dashboards has been used doughnut shape pie
-charts to represent total COVID-19 confirmed cases, recovered cases,
-active cases \& deaths as a percentage. Also, region, gender, age group
-\& ethnicity can be identified as common breakdowns of COVID-19 cases.
-In some dashboard has been added data tables for representing cases
-distribution by province/region. Only very few dashboards have been
-visualized in COVID-19 test details. Only 6 dashboards have been
-compared to global situations. In addition, fatality rate, incidence
-rate, ICU beds, stage of the patients \& hospitalize details have been
-contained in the several dashboards. These are the most common data
-types that have been contained in the dashboards. The main tools that
-have been used for different purpose can be shown in the following
-table.\hfill\break
+Table 03 summarizes the visualization methods used in the dashboards. In
+almost every dashboard, value boxes have been used to represent these
+total figures. Bar charts and line charts (trend lines) are most
+frequently used to visualize this data (confirmed cases, recovered
+cases, deaths, and vaccination details) with respect to time. A majority
+of dashboards presented data either daily or weekly. The mapping is used
+to track the spatial distribution of COVID-19 cases by country,
+province, regional etc. When visualizing the data by the map color code
+system, circles with respect to the size of the cases have been used to
+visualize the variation in size. The considerable number of dashboards
+has been used in doughnut-shaped pie charts to represent total COVID-19
+confirmed cases, recovered cases, active cases, and deaths as a
+percentage. Also, region, gender, age group, and ethnicity can be
+identified as common breakdowns of COVID-19 cases. Data tables for
+representing case~distribution by province/region have been added to
+some dashboards. Very few dashboards have been visualized in the
+COVID-19 test details. Only 6 dashboards have been compared to global
+situations. In addition, the fatality rate, incidence rate, ICU beds,
+stage of the patients, and hospitalized details have been contained in
+the several dashboards.\hfill\break
 

### 2022-05-15 manuscript_covid19dashboard.Rmd
Updated page with images

@@
 
-The source code to reprouduce the results are available in a public GitHub repository and can be access online via https://github.com/thiyangt/covid19srilanka
-
-# Results {#results}
-
-The "Sri Lanka COVID-19 Dashboard" provides an overview of the COVID-19 pandemic and administration of vaacine information in Sri Lanka.
-Figure 1 shows an screenshot of the dashboard.
-This dashboard has eight panels as listed in Table 6.
-The deployed dashboard is made accessible to public through https://thiyanga.netlify.app/post/covid19srilankadb/
-
-```{r, echo=FALSE, fig.cap="Overview Panel: Screeshot of the Dashboard", out.width="80%", fig.align='center'}
-knitr::include_graphics("images/dashboard.png")
-```
-
-\newpage
-
-**Table 6: Description of the panels**
-
-| **Name of the Panel**|**Description of the Panel**| 
-|:------|:----------------------------------------------|
-|Overview          |Total count of COVID-19 confirmed, recovered, deaths, active cases & total vaccine doses administered.\hfill\break Provide an overview of daily COVID-19 confirmed, recovered & deaths by plots.\hfill\break|
-|Cases by Wave| There are three tabs in this panel.\hfill\break  * Total COVID-19 confirmed cases - Cumulative count of COVID-19 confirmed cases with specific dates\hfill\break * COVID-19 Cases Distribution by Wave - Daily confirmed cases by wave\hfill\break * COVID-19 Deaths Distribution by Wave - Daily deaths by wave\hfill\break|
-|COVID-19 Patients Distribution | Provide an overview of confirmed cases district wise distribution.
-There are four tabs in this panel.\hfill\break * Total COVID-19 Patients Distribution in Sri Lanka - Total confirmed counts for each district is represented by tree map\hfill\break * Country Map - Total confirmed cases in each district represented by Sri Lanka country map\hfill\break * Distribution of Daily COVID-19 Patients for Last 30 Days - Visualize the daily confirmed cases distribution by districts in last 30 days\hfill\break * By Applying Min-Max Transformation - Visualize the details in the third tab by applying min-max transformation for each district\hfill\break|
-|Vaccination Details | Provide an overview of COVID-19 vaccination in Sri Lanka.
-There are two tabs.\hfill\break * Total Vaccine Doses - Visualize the administered vaccine doses as first dose only & fully vaccinated\hfill\break * Total Administered Doses by Vaccine Name - Visualize the vaccination by vaccine names\hfill\break|
-| Top 10 Countries| In this panel, compare the Sri Lanka confirmed & deaths with top 10 countries in the world (top 10 countries - The countries which have been reported highest number of confirmed cases as 31st of August 2021).\hfill\break There are two tabs.\hfill\break * Comparison of Cumulative Cases in Sri Lanka with Top 10 Countries - Compare the confirmed and deaths in Sri Lanka with top 10 countries by cumulative time series plots\hfill\break * Comparison of Log of Cumulative Cases in Sri Lanka with Top 10 Countries -  Compare the confirmed and deaths in Sri Lanka with top 10 countries by log cumulative time series plots\hfill\break (The data has been pulled from WHO)\hfill\break|
-|Global Comparison | Compare the total confirmed & deaths in Sri lanka with top 10 countries in Global & Asia.
-There are two tabs.\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the World - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the world & compare the case fatality ratios\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the Asia - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the Asia & compare the case fatality ratios\hfill\break|
-|Global Map| Visualize the distribution of confirmed, recovered & deaths in the world by world map.\hfill\break|
-|About | This panel contains the details about the dashboard.|
+We now describe the novel visualization approaches we included in out dashboard.
+To effectively distribute the vaccine and to support situational awareness and inform policy makers decision making it is important to know the district-wise spread of COVID-19 cases.
+We have daily COVID-19 data related to confirmed cases in all 25 districts in Sri Lanka.
+This structure generates a multiple time series collection.
+Visualizing these time series data is useful to identify similarities and dissimilarities between districts and their general trends.
+There are two approaches to visualizing these time series: (i) drawing individual time series plots for each district (as shown in Figure 2), and (ii) simultaneously plots all time series on a single panel (as shown in Figure 3).
+Plotting all time series simultaneously is also not possible due to overlapping time series and scale differences.
+Plotting separate panels for each district is not effective.
+The reason is that it is hard to compare across 25 different panels at once.
+In order to overcome these problems in multiple time series visualization,  we use heat maps ((@peng2008method)) to visualize global and local similarities and dissimilarities across districts.
+The associated results are shown in Figure 3.
+Here two heatmaps are used to show the global variations (Figure 3: A) and local variation (Figure 3: B) in the time series collection.
+Figure 3-A cell colours represent the actual counts of the COVID-19 confirmed cases.
+This is useful to get an idea about the differences in absolute values.
+Figure 3-B cell colours represent the normalized values created applying the min-max transformation.
+Min-max transformation is applied to each district time series by using the corresponding district minimum and maximum value of the time series.
+This helps us to get an idea about patterns within districts.
+For example, according to Figure 3A we can see Colombo, Gampaha and Kalutara districs COVID-19 cases are significantly higher than other districts.
+According to Figure 3B all districts show an increasing trend pattern as the right-hand side of the cells are lighter than the left-hand side cells in the heat map.
+Furthermore according to Figure 3B all districts reported high number of cases on  19, 24, 29 August, 2021.
+Figure 3B is useful for identifying these local outlying behaviours.
+As shown in Figure 4, we also use Choropleth map and Dorling cartogram to visualize spatial distribution of COVID-19 cases.
+The vaccination information are visualize through interactive time series plots and bar charts.
+A screenshot of associated panels are shown in Figure 5 and Figure 6.
 
@@
 
-We now describe the novel visualization approaches we included in out dashboard.
-To effectively distribute the vaccine and to support situational awareness and inform policy makers decision making it is important to know the district-wise spread of COVID-19 cases.
-We have daily COVID-19 data related to confirmed cases in all 25 districts in Sri Lanka.
-This structure generates a multiple time series collection.
-Visualizing these time series data is useful to identify similarities and dissimilarities between districts and their general trends.
-There are two approaches to visualizing these time series: (i) drawing individual time series plots for each district (as shown in Figure 2), and (ii) simultaneously plots all time series on a single panel (as shown in Figure 3).
-Plotting all time series simultaneously is also not possible due to overlapping time series and scale differences.
-Plotting separate panels for each district is not effective.
-The reason is that it is hard to compare across 25 different panels at once.
-In order to overcome these problems in multiple time series visualization,  we use heat maps ((@peng2008method)) to visualize global and local similarities and dissimilarities across districts.
-The associated results are shown in Figure 3.
-Here two heatmaps are used to show the global variations (Figure 3: A) and local variation (Figure 3: B) in the time series collection.
-Figure 3-A cell colours represent the actual counts of the COVID-19 confirmed cases.
-This is useful to get an idea about the differences in absolute values.
-Figure 3-B cell colours represent the normalized values created applying the min-max transformation.
-Min-max transformation is applied to each district time series by using the corresponding district minimum and maximum value of the time series.
-This helps us to get an idea about patterns within districts.
-For example, according to Figure 3A we can see Colombo, Gampaha and Kalutara districs COVID-19 cases are significantly higher than other districts.
-According to Figure 3B all districts show an increasing trend pattern as the right-hand side of the cells are lighter than the left-hand side cells in the heat map.
-Furthermore according to Figure 3B all districts reported high number of cases on  19, 24, 29 August, 2021.
-Figure 3B is useful for identifying these local outlying behaviours.
-As shown in Figure 4, we also use Choropleth map and Dorling cartogram to visualize spatial distribution of COVID-19 cases.
-The vaccination information are visualize through interactive time series plots and bar charts.
-A screenshot of associated panels are shown in Figure 5 and Figure 6.
-
-```{r, comment=NA, message=FALSE, warning=FALSE, fig.cap="Spatial DIstribution of COVID1-19 Cases by Districts"}
+```{r, comment=NA, message=FALSE, warning=FALSE, fig.cap="Spatial DIstribution of COVID1-19 Cases by Districts", fig.height=8}
 
@@
 
-```{r, echo=FALSE, fig.cap="Time Series Plot of Distribution of First Dose and Second Dose", out.width="80%", fig.align='center'}
-knitr::include_graphics("Images/d3.png")
+The source code to reprouduce the results are available in a public GitHub repository and can be access online via https://github.com/thiyangt/covid19srilanka
+
+# Results {#results}
+
+The "Sri Lanka COVID-19 Dashboard" provides an overview of the COVID-19 pandemic and administration of vaacine information in Sri Lanka.  This dashboard has eight panels as listed in Table 6.
+
+**Table 6: Description of the panels**
+
+| **Name of the Panel**|**Description of the Panel**| 
+|:------|:----------------------------------------------|
+|Overview          |Total count of COVID-19 confirmed, recovered, deaths, active cases & total vaccine doses administered.\hfill\break Provide an overview of daily COVID-19 confirmed, recovered & deaths by plots.\hfill\break|
+|Cases by Wave| There are three tabs in this panel.\hfill\break  * Total COVID-19 confirmed cases - Cumulative count of COVID-19 confirmed cases with specific dates\hfill\break * COVID-19 Cases Distribution by Wave - Daily confirmed cases by wave\hfill\break * COVID-19 Deaths Distribution by Wave - Daily deaths by wave\hfill\break|
+|COVID-19 Patients Distribution | Provide an overview of confirmed cases district wise distribution.
+There are four tabs in this panel.\hfill\break * Total COVID-19 Patients Distribution in Sri Lanka - Total confirmed counts for each district is represented by tree map\hfill\break * Country Map - Total confirmed cases in each district represented by Sri Lanka country map\hfill\break * Distribution of Daily COVID-19 Patients for Last 30 Days - Visualize the daily confirmed cases distribution by districts in last 30 days\hfill\break * By Applying Min-Max Transformation - Visualize the details in the third tab by applying min-max transformation for each district\hfill\break|
+|Vaccination Details | Provide an overview of COVID-19 vaccination in Sri Lanka.
+There are two tabs.\hfill\break * Total Vaccine Doses - Visualize the administered vaccine doses as first dose only & fully vaccinated\hfill\break * Total Administered Doses by Vaccine Name - Visualize the vaccination by vaccine names\hfill\break|
+| Top 10 Countries| In this panel, compare the Sri Lanka confirmed & deaths with top 10 countries in the world (top 10 countries - The countries which have been reported highest number of confirmed cases as 31st of August 2021).\hfill\break There are two tabs.\hfill\break * Comparison of Cumulative Cases in Sri Lanka with Top 10 Countries - Compare the confirmed and deaths in Sri Lanka with top 10 countries by cumulative time series plots\hfill\break * Comparison of Log of Cumulative Cases in Sri Lanka with Top 10 Countries -  Compare the confirmed and deaths in Sri Lanka with top 10 countries by log cumulative time series plots\hfill\break (The data has been pulled from WHO)\hfill\break|
+|Global Comparison | Compare the total confirmed & deaths in Sri lanka with top 10 countries in Global & Asia.
+There are two tabs.\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the World - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the world & compare the case fatality ratios\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the Asia - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the Asia & compare the case fatality ratios\hfill\break|
+|Global Map| Visualize the distribution of confirmed, recovered & deaths in the world by world map.\hfill\break|
+|About | This panel contains the details about the dashboard.|
+
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 1", out.width="80%", fig.align='center'}
+knitr::include_graphics("Images/image1.png")
 ```
 
-```{r, echo=FALSE, fig.cap="Distribution of Vaccines by Different Types and Dose", out.width="80%", fig.align='center'}
-knitr::include_graphics("Images/d2.png")
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 2", out.width="80%", fig.align='center'}
+knitr::include_graphics("Images/image2.png")
 ```
+
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 3", out.width="80%", fig.align='center'}
+knitr::include_graphics("Images/image3.png")
+```
+
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 4", out.width="80%", fig.align='center'}
+knitr::include_graphics("Images/image4.png")
+```
+
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 5", out.width="80%", fig.align='center'}
+knitr::include_graphics("Images/image5.png")
+```
+
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 6", out.width="80%", fig.align='center'}
+knitr::include_graphics("Images/image6.png")
+```
+
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 7", out.width="80%", fig.align='center'}
+knitr::include_graphics("Images/image7.png")
+```
+
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 8", out.width="80%", fig.align='center'}
+knitr::include_graphics("Images/image8.png")
+```
+
+\newpage
 
@@
 <div id="refs"></div>
-
-\newpage
-
-# Appendix: Overview of reviewed dashboard
-
-Real time updated COVID-19 dashboard created by the John Hopkins University Center for Systems Science & Engineering (JHU CSSE) has included total confirmed cases & total deaths by country, province/region/state (not for all countries)[6].
-Weekly & daily global confirmed cases, deaths & vaccine doses have been visualized using bar plots related to the date in the side panels.
-Also, an interactive world map has included in the dashboard to represent total confirmed cases, incidence rate, fatality ratio & vaccine doses administered by country & state of US.
-All the links related to the data sets has been included in the dashboard.
-In this dashboard red, white, green colors has been used to indicate confirmed cases, deaths & vaccine doses respectively.
-The dashboard has fitted on a single screen.
-
-```{r, fig.cap="Johns Hopkins University COVID-19 Dashboard"}
-knitr::include_graphics("Images/1.PNG")
-```
-
-Live updated “WHO COVID-19 dashboard” contained four panels. The interactive world
-map in the dashboard has visualized the distribution of COVID-19 confirmed cases, deaths &
-vaccination by country [15].
-Line chart & bar charts have been used as visualization plots.
-Blue, red, green colors have been used to represent confirmed cases, deaths & vaccination respectively.
-The links related to the underlying data was available in the dashboard.
-The dashboard was not fitted on a single screen.
-
-```{r, fig.cap="WHO COVID-19 Dashboard"}
-knitr::include_graphics("Images/2.PNG")
-```
-
-The “COVID-19 surveillance dashboard” created by the University of Virginia in
-collaboration with Bio complexity Institute has included interactive world map with a time slide to visualize the confirmed cases, recovered cases, deaths, active cases & vaccine details [7].
-Line charts, bar charts & data tables have been visualized in the dashboard.
-Red, blue, green & yellow colors have been used to identify confirmed, deceased, recovered & active cases respectively.
-The data behind the dashboard also available on the dashboard & can be downloaded.
-The dashboard has fitted on a single screen.
-
-```{r, fig.cap="COVID-19 Surveillance Dashboard"}
-knitr::include_graphics("Images/3.PNG")
-```
-
-Corona cases (COVID-19) per municipality in Belgium dashboard has included interactive
-map for distribution of total confirmed cases per municipality, bar chart for number of cases per province, line charts for number of cases per day by region & municipality, dodge bar chart for deaths per region, per age & pie chart for deaths per age as visualization methods[3].
-Also it included several value boxes for total cases, total deaths & hospital situations.
-There was no specific color theme used in this dashboard.
-This dashboard also fitted on a single screen.
-
-```{r, fig.cap="Corona cases per Municipality in Belgium Dashboard"}
-knitr::include_graphics("Images/4.PNG")
-```
-
-The COVID-19 dashboard for England created by NHS providers included daily totals, daily changes & weekly changes in new cases, hospital admission, patients in hospital, and patients
-in ICU beds, deaths & vaccinations in England as counts [2].
-Bar plots for daily new cases, hospital admissions & deaths for last 30 days have been added in this dashboard.
-Also it has included data table for total count, daily changes (as increase or decrease), weekly changes (as increase or decrease) of new cases & hospital admission by region.
-It can be seen that, three different colors have been used in bar plot as orange for new cases, green for hospital admissions & red for deaths.
-These plots were not interactive plots & there was only one panel.
-The dashboard was not fitted on a single screen.
-
-```{r, fig.cap="COVID-19 dashboard for England"}
-knitr::include_graphics("Images/5.PNG")
-```
-
-The Environmental Science and Research Institute COVID-19 dashboard for New Zealand (was developed & is maintained in collaboration with Epi-interactive Ltd) contained five panels
-as overview, outbreak, source, international & ESR reports [10].
-In this real time updated dashboard, they have included value boxes for total confirmed cases, recovered cases, deaths & interactive map has been added for distribution of confirmed & probable cases as incidence & count.
-Also they have included bar charts for confirmed cases distribution by gender, age group, & line charts for confirmed & deaths as daily & cumulative.
-The dashboard has not fitted on a single screen.
-
-```{r, fig.cap="New Zealand COVID-19 Dashboard"}
-knitr::include_graphics("Images/6.PNG")
-```
-
-Pakistan’s official COVID-19 dashboard is a real time updated dashboard with one panel. This dashboard contains value boxes for total confirmed cases, active cases, deaths, recovered
-cases by country.
-Province wise cases have been represented by map & data table [11].
-Line charts & bar charts have been added to visualize the distribution of cases related to time & province.
-In the value boxes blue, orange, pink & green colors have been applied for confirmed cases, active cases, deaths & recoveries respectively, but there was no specific color theme used when plotting the graphs.
-The dashboard cannot be seen on a single screen without grid overlay.
-
-```{r, fig.cap="COVID-19 Dashboard for Pakistan"}
-knitr::include_graphics("Images/7.PNG")
-```
-
-COVID-19 Canada live dashboard has been developed with three panels for provincial data, hospital resources & projections [14].
-This dashboard contained interactive map for represented total reported cases & deaths by province.
-Data tables, line charts, bar charts were the most frequently used visualization types of this dashboard.
-More details were presented on this dashboard by province like ventilator counts, hospital beds per 100 residents, resource capacity based on critical case rate, effect of social distance, forecast of deaths & cases.
-There was no specific color theme applied for plotting the graphs.
-The dashboard also was not fitted on a single screen.
-
-```{r, fig.cap="COVID-19 in Canada dashboard"}
-knitr::include_graphics("Images/8.PNG")
-```
-
-“Zoho analytics COVID-19 live dashboard” for India contained three panels as insight by state, trend analysis & vaccination [1].
-Interactive country map has been used to visualize the distribution of confirmed cases, active cases, deaths, recovered cases & vaccination by states in India.
-Line charts, bar charts & doughnut shape pie charts have been used as other visualization tools.
-Also this dashboard contained data table for overview of COVID-19 in India.
-This dashboard also was not fitted on a single screen.
-
-```{r, fig.cap="India COVID-19 Dashboard"}
-knitr::include_graphics("Images/9.PNG")
-```
-
-Italy COVID-19 dashboard created by the COVID-19 task force of the department of infection diseases and the IT service Istituto Superiore Sanita, contained interactive map, bar charts, doughnut shape pie charts & heat maps as visualization tools [4].
-It provides data related to COVID-19 cases, cases distribution by age, clinical status, and weekly number of cases notified in Italy by region/province.
-This dashboard also was not fitted on a single screen.
-
-```{r, fig.cap="Italy COVID-19 Dashboard"}
-knitr::include_graphics("Images/10.PNG")
-```
-
-Live COVID-19 dashboard developed & contributed by an amber innovations for the government of Jamaica has added interactive country map & data table to represent the COVID-
-19 cases by parish [9].
-Line charts & bar charts have been added to the dashboard for cumulative & daily COVID-19 cases.
-Yellow, green & red colors have been used to represent confirmed, recovered & deaths respectively.
-Doughnut shape pie chart has been added to represent the distribution of recovered, confirmed & deaths by age group.
-This dashboard was not fitted on a single screen.
-
-```{r, fig.cap="Jamaica COVID-19 Dashboard"}
-knitr::include_graphics("Images/11.PNG")
-```
-
-The GCI COVID-19 dashboard for Russia includes an interactive map, bar charts & line charts to visualize the COVID-19 data [13]. Yellow, green, red & blue colors have been added to
-identify the confirmed, recovered, deaths & active case respectively in plotting the charts.
-
-```{r, fig.cap="GCI COVID-19 Dashboard for Russia"}
-knitr::include_graphics("Images/12.PNG")
-```
-
-COVID-19 live situation analysis dashboard of Sri Lanka contains total & daily figures in Sri Lanka as counts [5].
-Total cases & active cases have been visualized by using a line chart.
-Bar charts have been used to summarize the daily confirmed cases and recovered cases.
-As well fatality rate, recovery rate & daily investigations using PCR tests & rapid antigen tests have been compared by using the bar charts.
-Doughnut shape pie chart has been used to summarize the total cases.
-Blue, green & red colors have been applied to active, recovered & deaths respectively.
-This dashboard was also not fitted on a single screen.
-
-```{r, fig.cap="COVID-19 Live Situational Analysis Dashboard of Sri Lanka"}
-knitr::include_graphics("Images/13.PNG")
-```
-
-COVID-19 live situation analysis dashboard of Sri Lanka contains total & daily figures in Sri Lanka as counts [5].
-Total cases & active cases have been visualized by using a line chart.
-Bar charts have been used to summarize the daily confirmed cases and recovered cases.
-As well fatality rate, recovery rate & daily investigations using PCR tests & rapid antigen tests have been compared by using the bar charts.
-Doughnut shape pie chart has been used to summarize the total cases.
-Blue, green & red colors have been applied to active, recovered & deaths respectively.
-This dashboard was also not fitted on a single screen.
-
-```{r, fig.cap="COVID 19 ZA South Africa Dashboard"}
-knitr::include_graphics("Images/14.PNG")
-```
-
-Robert Koch-Institute COVID-19 dashboard for German contains interactive map, line charts & bar charts as visualization tools of COVID-19 data [12]. The dashboard has been fitted
-on a single screen. This dashboard format is similar to the John Hopkins dashboard.
-
-```{r, fig.cap="Robert Koch-Institute COVID-19 dashboard for German"}
-knitr::include_graphics("Images/15.PNG")
-```

### 2022-05-15 manuscript_covid19dashboard.tex
Updated page with images

@@
 identified as common breakdowns of COVID-19 cases. Data tables for
-representing case~distribution by province/region have been added to
+representing case distribution by province/region have been added to
 some dashboards. Very few dashboards have been visualized in the
@@
 
+We now describe the novel visualization approaches we included in out
+dashboard. To effectively distribute the vaccine and to support
+situational awareness and inform policy makers decision making it is
+important to know the district-wise spread of COVID-19 cases. We have
+daily COVID-19 data related to confirmed cases in all 25 districts in
+Sri Lanka. This structure generates a multiple time series collection.
+Visualizing these time series data is useful to identify similarities
+and dissimilarities between districts and their general trends. There
+are two approaches to visualizing these time series: (i) drawing
+individual time series plots for each district (as shown in Figure 2),
+and (ii) simultaneously plots all time series on a single panel (as
+shown in Figure 3). Plotting all time series simultaneously is also not
+possible due to overlapping time series and scale differences. Plotting
+separate panels for each district is not effective. The reason is that
+it is hard to compare across 25 different panels at once. In order to
+overcome these problems in multiple time series visualization, we use
+heat maps ((Peng (2008))) to visualize global and local similarities and
+dissimilarities across districts. The associated results are shown in
+Figure 3. Here two heatmaps are used to show the global variations
+(Figure 3: A) and local variation (Figure 3: B) in the time series
+collection. Figure 3-A cell colours represent the actual counts of the
+COVID-19 confirmed cases. This is useful to get an idea about the
+differences in absolute values. Figure 3-B cell colours represent the
+normalized values created applying the min-max transformation. Min-max
+transformation is applied to each district time series by using the
+corresponding district minimum and maximum value of the time series.
+This helps us to get an idea about patterns within districts. For
+example, according to Figure 3A we can see Colombo, Gampaha and Kalutara
+districs COVID-19 cases are significantly higher than other districts.
+According to Figure 3B all districts show an increasing trend pattern as
+the right-hand side of the cells are lighter than the left-hand side
+cells in the heat map. Furthermore according to Figure 3B all districts
+reported high number of cases on 19, 24, 29 August, 2021. Figure 3B is
+useful for identifying these local outlying behaviours. As shown in
+Figure 4, we also use Choropleth map and Dorling cartogram to visualize
+spatial distribution of COVID-19 cases. The vaccination information are
+visualize through interactive time series plots and bar charts. A
+screenshot of associated panels are shown in Figure 5 and Figure 6.
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{Distribution of COVID19 Cases by Districts: (A) Plotted on a
+Single Panel, (B) Plotted on Separate Panels.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{(A) Global View of Distribution of COVID19 Cases by DIstricts,
+(B) Local View of Distribution of COVID19 Cases by Districts}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{Spatial DIstribution of COVID1-19 Cases by Districts}
+\end{figure}
+
 The source code to reprouduce the results are available in a public
@@
 COVID-19 pandemic and administration of vaacine information in Sri
-Lanka. Figure 1 shows an screenshot of the dashboard. This dashboard has
-eight panels as listed in Table 6. The deployed dashboard is made
-accessible to public through
-\url{https://thiyanga.netlify.app/post/covid19srilankadb/}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.8\linewidth]{images/dashboard}
-
-}
-
-\caption{Overview Panel: Screeshot of the Dashboard}\label{fig:unnamed-chunk-1}
-\end{figure}
-
-\newpage
+Lanka. This dashboard has eight panels as listed in Table 6.
 
@@
 \begin{figure}
-\centering
-\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-3-1.pdf}
-\caption{Distribution of COVID19 Cases by Districts: (A) Plotted on a
-Single Panel, (B) Plotted on Separate Panels.}
-\end{figure}
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{(A) Global View of Distribution of COVID19 Cases by DIstricts,
-(B) Local View of Distribution of COVID19 Cases by Districts}
-\end{figure}
+{\centering \includegraphics[width=0.8\linewidth]{Images/image1}
 
-We now describe the novel visualization approaches we included in out
-dashboard. To effectively distribute the vaccine and to support
-situational awareness and inform policy makers decision making it is
-important to know the district-wise spread of COVID-19 cases. We have
-daily COVID-19 data related to confirmed cases in all 25 districts in
-Sri Lanka. This structure generates a multiple time series collection.
-Visualizing these time series data is useful to identify similarities
-and dissimilarities between districts and their general trends. There
-are two approaches to visualizing these time series: (i) drawing
-individual time series plots for each district (as shown in Figure 2),
-and (ii) simultaneously plots all time series on a single panel (as
-shown in Figure 3). Plotting all time series simultaneously is also not
-possible due to overlapping time series and scale differences. Plotting
-separate panels for each district is not effective. The reason is that
-it is hard to compare across 25 different panels at once. In order to
-overcome these problems in multiple time series visualization, we use
-heat maps ((Peng (2008))) to visualize global and local similarities and
-dissimilarities across districts. The associated results are shown in
-Figure 3. Here two heatmaps are used to show the global variations
-(Figure 3: A) and local variation (Figure 3: B) in the time series
-collection. Figure 3-A cell colours represent the actual counts of the
-COVID-19 confirmed cases. This is useful to get an idea about the
-differences in absolute values. Figure 3-B cell colours represent the
-normalized values created applying the min-max transformation. Min-max
-transformation is applied to each district time series by using the
-corresponding district minimum and maximum value of the time series.
-This helps us to get an idea about patterns within districts. For
-example, according to Figure 3A we can see Colombo, Gampaha and Kalutara
-districs COVID-19 cases are significantly higher than other districts.
-According to Figure 3B all districts show an increasing trend pattern as
-the right-hand side of the cells are lighter than the left-hand side
-cells in the heat map. Furthermore according to Figure 3B all districts
-reported high number of cases on 19, 24, 29 August, 2021. Figure 3B is
-useful for identifying these local outlying behaviours. As shown in
-Figure 4, we also use Choropleth map and Dorling cartogram to visualize
-spatial distribution of COVID-19 cases. The vaccination information are
-visualize through interactive time series plots and bar charts. A
-screenshot of associated panels are shown in Figure 5 and Figure 6.
+}
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{Spatial DIstribution of COVID1-19 Cases by Districts}
+\caption{Screeshot of Panel 1}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{Images/d3}
+{\centering \includegraphics[width=0.8\linewidth]{Images/image2}
 
@@
 
-\caption{Time Series Plot of Distribution of First Dose and Second Dose}\label{fig:unnamed-chunk-6}
+\caption{Screeshot of Panel 2}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{Images/d2}
+{\centering \includegraphics[width=0.8\linewidth]{Images/image3}
 
@@
 
-\caption{Distribution of Vaccines by Different Types and Dose}\label{fig:unnamed-chunk-7}
+\caption{Screeshot of Panel 3}\label{fig:unnamed-chunk-7}
 \end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{Images/image4}
+
+}
+
+\caption{Screeshot of Panel 4}\label{fig:unnamed-chunk-8}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{Images/image5}
+
+}
+
+\caption{Screeshot of Panel 5}\label{fig:unnamed-chunk-9}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{Images/image6}
+
+}
+
+\caption{Screeshot of Panel 6}\label{fig:unnamed-chunk-10}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{Images/image7}
+
+}
+
+\caption{Screeshot of Panel 7}\label{fig:unnamed-chunk-11}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{Images/image8}
+
+}
+
+\caption{Screeshot of Panel 8}\label{fig:unnamed-chunk-12}
+\end{figure}
+
+\newpage
 
@@
 
-\newpage
-
-\hypertarget{appendix-overview-of-reviewed-dashboard}{%
-\section{Appendix: Overview of reviewed
-dashboard}\label{appendix-overview-of-reviewed-dashboard}}
-
-Real time updated COVID-19 dashboard created by the John Hopkins
-University Center for Systems Science \& Engineering (JHU CSSE) has
-included total confirmed cases \& total deaths by country,
-province/region/state (not for all countries){[}6{]}. Weekly \& daily
-global confirmed cases, deaths \& vaccine doses have been visualized
-using bar plots related to the date in the side panels. Also, an
-interactive world map has included in the dashboard to represent total
-confirmed cases, incidence rate, fatality ratio \& vaccine doses
-administered by country \& state of US. All the links related to the
-data sets has been included in the dashboard. In this dashboard red,
-white, green colors has been used to indicate confirmed cases, deaths \&
-vaccine doses respectively. The dashboard has fitted on a single screen.
-
-\begin{figure}
-\includegraphics[width=8.42in]{Images/1} \caption{Johns Hopkins University COVID-19 Dashboard}\label{fig:unnamed-chunk-8}
-\end{figure}
-
-Live updated ``WHO COVID-19 dashboard'' contained four panels. The
-interactive world map in the dashboard has visualized the distribution
-of COVID-19 confirmed cases, deaths \& vaccination by country {[}15{]}.
-Line chart \& bar charts have been used as visualization plots. Blue,
-red, green colors have been used to represent confirmed cases, deaths \&
-vaccination respectively. The links related to the underlying data was
-available in the dashboard. The dashboard was not fitted on a single
-screen.
-
-\begin{figure}
-\includegraphics[width=8.35in]{Images/2} \caption{WHO COVID-19 Dashboard}\label{fig:unnamed-chunk-9}
-\end{figure}
-
-The ``COVID-19 surveillance dashboard'' created by the University of
-Virginia in collaboration with Bio complexity Institute has included
-interactive world map with a time slide to visualize the confirmed
-cases, recovered cases, deaths, active cases \& vaccine details {[}7{]}.
-Line charts, bar charts \& data tables have been visualized in the
-dashboard. Red, blue, green \& yellow colors have been used to identify
-confirmed, deceased, recovered \& active cases respectively. The data
-behind the dashboard also available on the dashboard \& can be
-downloaded. The dashboard has fitted on a single screen.
-
-\begin{figure}
-\includegraphics[width=8.42in]{Images/3} \caption{COVID-19 Surveillance Dashboard}\label{fig:unnamed-chunk-10}
-\end{figure}
-
-Corona cases (COVID-19) per municipality in Belgium dashboard has
-included interactive map for distribution of total confirmed cases per
-municipality, bar chart for number of cases per province, line charts
-for number of cases per day by region \& municipality, dodge bar chart
-for deaths per region, per age \& pie chart for deaths per age as
-visualization methods{[}3{]}. Also it included several value boxes for
-total cases, total deaths \& hospital situations. There was no specific
-color theme used in this dashboard. This dashboard also fitted on a
-single screen.
-
-\begin{figure}
-\includegraphics[width=8.39in]{Images/4} \caption{Corona cases per Municipality in Belgium Dashboard}\label{fig:unnamed-chunk-11}
-\end{figure}
-
-The COVID-19 dashboard for England created by NHS providers included
-daily totals, daily changes \& weekly changes in new cases, hospital
-admission, patients in hospital, and patients in ICU beds, deaths \&
-vaccinations in England as counts {[}2{]}. Bar plots for daily new
-cases, hospital admissions \& deaths for last 30 days have been added in
-this dashboard. Also it has included data table for total count, daily
-changes (as increase or decrease), weekly changes (as increase or
-decrease) of new cases \& hospital admission by region. It can be seen
-that, three different colors have been used in bar plot as orange for
-new cases, green for hospital admissions \& red for deaths. These plots
-were not interactive plots \& there was only one panel. The dashboard
-was not fitted on a single screen.
-
-\begin{figure}
-\includegraphics[width=8.46in]{Images/5} \caption{COVID-19 dashboard for England}\label{fig:unnamed-chunk-12}
-\end{figure}
-
-The Environmental Science and Research Institute COVID-19 dashboard for
-New Zealand (was developed \& is maintained in collaboration with
-Epi-interactive Ltd) contained five panels as overview, outbreak,
-source, international \& ESR reports {[}10{]}. In this real time updated
-dashboard, they have included value boxes for total confirmed cases,
-recovered cases, deaths \& interactive map has been added for
-distribution of confirmed \& probable cases as incidence \& count. Also
-they have included bar charts for confirmed cases distribution by
-gender, age group, \& line charts for confirmed \& deaths as daily \&
-cumulative. The dashboard has not fitted on a single screen.
-
-\begin{figure}
-\includegraphics[width=8.47in]{Images/6} \caption{New Zealand COVID-19 Dashboard}\label{fig:unnamed-chunk-13}
-\end{figure}
-
-Pakistan's official COVID-19 dashboard is a real time updated dashboard
-with one panel. This dashboard contains value boxes for total confirmed
-cases, active cases, deaths, recovered cases by country. Province wise
-cases have been represented by map \& data table {[}11{]}. Line charts
-\& bar charts have been added to visualize the distribution of cases
-related to time \& province. In the value boxes blue, orange, pink \&
-green colors have been applied for confirmed cases, active cases, deaths
-\& recoveries respectively, but there was no specific color theme used
-when plotting the graphs. The dashboard cannot be seen on a single
-screen without grid overlay.
-
-\begin{figure}
-\includegraphics[width=8.33in]{Images/7} \caption{COVID-19 Dashboard for Pakistan}\label{fig:unnamed-chunk-14}
-\end{figure}
-
-COVID-19 Canada live dashboard has been developed with three panels for
-provincial data, hospital resources \& projections {[}14{]}. This
-dashboard contained interactive map for represented total reported cases
-\& deaths by province. Data tables, line charts, bar charts were the
-most frequently used visualization types of this dashboard. More details
-were presented on this dashboard by province like ventilator counts,
-hospital beds per 100 residents, resource capacity based on critical
-case rate, effect of social distance, forecast of deaths \& cases. There
-was no specific color theme applied for plotting the graphs. The
-dashboard also was not fitted on a single screen.
-
-\begin{figure}
-\includegraphics[width=8.47in]{Images/8} \caption{COVID-19 in Canada dashboard}\label{fig:unnamed-chunk-15}
-\end{figure}
-
-``Zoho analytics COVID-19 live dashboard'' for India contained three
-panels as insight by state, trend analysis \& vaccination {[}1{]}.
-Interactive country map has been used to visualize the distribution of
-confirmed cases, active cases, deaths, recovered cases \& vaccination by
-states in India. Line charts, bar charts \& doughnut shape pie charts
-have been used as other visualization tools. Also this dashboard
-contained data table for overview of COVID-19 in India. This dashboard
-also was not fitted on a single screen.
-
-\begin{figure}
-\includegraphics[width=8.42in]{Images/9} \caption{India COVID-19 Dashboard}\label{fig:unnamed-chunk-16}
-\end{figure}
-
-Italy COVID-19 dashboard created by the COVID-19 task force of the
-department of infection diseases and the IT service Istituto Superiore
-Sanita, contained interactive map, bar charts, doughnut shape pie charts
-\& heat maps as visualization tools {[}4{]}. It provides data related to
-COVID-19 cases, cases distribution by age, clinical status, and weekly
-number of cases notified in Italy by region/province. This dashboard
-also was not fitted on a single screen.
-
-\begin{figure}
-\includegraphics[width=8.12in]{Images/10} \caption{Italy COVID-19 Dashboard}\label{fig:unnamed-chunk-17}
-\end{figure}
-
-Live COVID-19 dashboard developed \& contributed by an amber innovations
-for the government of Jamaica has added interactive country map \& data
-table to represent the COVID- 19 cases by parish {[}9{]}. Line charts \&
-bar charts have been added to the dashboard for cumulative \& daily
-COVID-19 cases. Yellow, green \& red colors have been used to represent
-confirmed, recovered \& deaths respectively. Doughnut shape pie chart
-has been added to represent the distribution of recovered, confirmed \&
-deaths by age group. This dashboard was not fitted on a single screen.
-
-\begin{figure}
-\includegraphics[width=8.35in]{Images/11} \caption{Jamaica COVID-19 Dashboard}\label{fig:unnamed-chunk-18}
-\end{figure}
-
-The GCI COVID-19 dashboard for Russia includes an interactive map, bar
-charts \& line charts to visualize the COVID-19 data {[}13{]}. Yellow,
-green, red \& blue colors have been added to identify the confirmed,
-recovered, deaths \& active case respectively in plotting the charts.
-
-\begin{figure}
-\includegraphics[width=8.25in]{Images/12} \caption{GCI COVID-19 Dashboard for Russia}\label{fig:unnamed-chunk-19}
-\end{figure}
-
-COVID-19 live situation analysis dashboard of Sri Lanka contains total
-\& daily figures in Sri Lanka as counts {[}5{]}. Total cases \& active
-cases have been visualized by using a line chart. Bar charts have been
-used to summarize the daily confirmed cases and recovered cases. As well
-fatality rate, recovery rate \& daily investigations using PCR tests \&
-rapid antigen tests have been compared by using the bar charts. Doughnut
-shape pie chart has been used to summarize the total cases. Blue, green
-\& red colors have been applied to active, recovered \& deaths
-respectively. This dashboard was also not fitted on a single screen.
-
-\begin{figure}
-\includegraphics[width=8.4in]{Images/13} \caption{COVID-19 Live Situational Analysis Dashboard of Sri Lanka}\label{fig:unnamed-chunk-20}
-\end{figure}
-
-COVID-19 live situation analysis dashboard of Sri Lanka contains total
-\& daily figures in Sri Lanka as counts {[}5{]}. Total cases \& active
-cases have been visualized by using a line chart. Bar charts have been
-used to summarize the daily confirmed cases and recovered cases. As well
-fatality rate, recovery rate \& daily investigations using PCR tests \&
-rapid antigen tests have been compared by using the bar charts. Doughnut
-shape pie chart has been used to summarize the total cases. Blue, green
-\& red colors have been applied to active, recovered \& deaths
-respectively. This dashboard was also not fitted on a single screen.
-
-\begin{figure}
-\includegraphics[width=8.46in]{Images/14} \caption{COVID 19 ZA South Africa Dashboard}\label{fig:unnamed-chunk-21}
-\end{figure}
-
-Robert Koch-Institute COVID-19 dashboard for German contains interactive
-map, line charts \& bar charts as visualization tools of COVID-19 data
-{[}12{]}. The dashboard has been fitted on a single screen. This
-dashboard format is similar to the John Hopkins dashboard.
-
-\begin{figure}
-\includegraphics[width=8.33in]{Images/15} \caption{Robert Koch-Institute COVID-19 dashboard for German}\label{fig:unnamed-chunk-22}
-\end{figure}
-
 \end{document}

### 2022-05-15 manuscript_covid19dashboard.Rmd
Updated

@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 1", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 1: Overview", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image1.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 2", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 2: Cases by Wave", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image2.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 3", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 3: Distribution of COVID-19 Patients", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image3.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 4", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 4: Vaccination Details", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image4.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 5", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 5: Comparison with Top 10 Countries", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image5.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 6", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 6: Global Comparison", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image6.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 7", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 7: World Map", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image7.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 8", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 8: About", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image8.png")

### 2022-05-15 manuscript_covid19dashboard.tex
Updated

@@
 
-\caption{Screeshot of Panel 1}\label{fig:unnamed-chunk-5}
+\caption{Screeshot of Panel 1: Overview}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 2}\label{fig:unnamed-chunk-6}
+\caption{Screeshot of Panel 2: Cases by Wave}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 3}\label{fig:unnamed-chunk-7}
+\caption{Screeshot of Panel 3: Distribution of COVID-19 Patients}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 4}\label{fig:unnamed-chunk-8}
+\caption{Screeshot of Panel 4: Vaccination Details}\label{fig:unnamed-chunk-8}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 5}\label{fig:unnamed-chunk-9}
+\caption{Screeshot of Panel 5: Comparison with Top 10 Countries}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 6}\label{fig:unnamed-chunk-10}
+\caption{Screeshot of Panel 6: Global Comparison}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 7}\label{fig:unnamed-chunk-11}
+\caption{Screeshot of Panel 7: World Map}\label{fig:unnamed-chunk-11}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 8}\label{fig:unnamed-chunk-12}
+\caption{Screeshot of Panel 8: About}\label{fig:unnamed-chunk-12}
 \end{figure}

### 2022-05-15 manuscript_covid19dashboard.Rmd
Completed literature review

@@
 ---
-title: "Interactive dashboard to monitor the COVID-19 outbreak and Vaccine Administration"
+title: "Interactive Dashboard to Monitor the COVID-19 Outbreak and Vaccine Administration"
 output: 
@@
 COVID-19 has expanded over the globe, having a significant impact on our daily lives and work.
-Early responses and timely decisions and actions are critical to saving communities and economy worldwide.
+Early responses and timely decisions and actions are critical to saving communities and economies worldwide.
 Data is essential in order to make effective decisions.
@@
 Next, we developed an interactive dashboard to visualize the COVID-19 outbreak and vaccination information in Sri Lanka.
-This dashboard provides front-line health officers a situational awareness of the spread of COVID-19 and status of the vaccination program.
+This dashboard provides front-line health officers a situational awareness of the spread of COVID-19 and the status of the vaccination program.
 
@@
 
-Table 03 summarizes the visualization methods used in the dashboards.
-In almost every dashboard, value boxes have been used to represent these total figures.
-Bar charts and line charts (trend lines) are most frequently used to visualize this data (confirmed cases, recovered cases, deaths, and vaccination details) with respect to time.
-A majority of dashboards presented data either daily or weekly.
-The mapping is used to track the spatial distribution of COVID-19 cases by country, province, regional etc.
+Table 03 highlights the dashboard visualization techniques.
+Value boxes have been utilized to display total figures on practically every dashboard.
+The most common ways of visualizing confirmed cases, recovered cases, deaths, and immunization details are bar charts and line charts (trend lines).
+The majority of dashboards displayed data on a daily or weekly basis.
+The spatial distribution of COVID-19 cases by country, province, regional, and other factors is tracked using choropleth maps.
 When visualizing the data by the map color code system, circles with respect to the size of the cases have been used to visualize the variation in size.
-The considerable number of dashboards has been used in doughnut-shaped pie charts to represent total COVID-19 confirmed cases, recovered cases, active cases, and deaths as a percentage.
-Also, region, gender, age group, and ethnicity can be identified as common breakdowns of COVID-19 cases.
-Data tables for representing case distribution by province/region have been added to some dashboards.
+Several dashboards use doughnut-shaped pie charts to indicate total COVID-19 confirmed cases, recovered cases, active cases, and deaths as a proportion.
+Furthermore, region, gender, age group, and ethnicity can be identified as common breakdowns of COVID-19 cases.
+Data tables for representing cases' distribution by province/region have been added to some dashboards.
 Very few dashboards have been visualized in the COVID-19 test details.
@@
 
-## 2.2 Comparison of Dashboards
+## Comparison of Dashboards
 
-Before developing a dashboard, it is necessary to think about which visualization tools & features that should be contained in the dashboards.
-What are the most suitable plots, how many panels in the dashboard, what data should be included, how to fit dashboard on a screen, colors & is it real time updated or not are the common things that should be considered before develop the dashboards.
-In table 02 has been compared these visualization tools & features under the following categories.
+Before developing a dashboard, it is necessary to think about which visualization tools and features should be contained in the dashboard.
+What are the most suitable plots, how many panels in the dashboard, what data should be included, how to fit the dashboard on a screen, colors, and is it real time updated or not are the common things that should be considered before developing the dashboards.
+Table 4 summarizes information under the following categories:
 
-  * Number of panels - How many panels which are included in the dashboard.
-  * Visualization tools – what are the graphical representations of data which are contained in the dashboards. 
-  * Fitted on a single screen – whether the dashboard fits on a single screen or not (users can see the whole dashboard on a single screen without adjusting through grid overlay or not).
-  * Color theme – is there a unique color used for one data type in the whole dashboard (i.e.: one color scale for one data type everywhere on the dashboard). 
-  * Dark background – background color of the dashboard is dark or light.
-  * Data available – whether users can be downloaded/available the data set which has been visualized on the dashboard. 
-  * Real time updated – whether the dashboard is updated daily/ specific time (live dashboard) or not.
+  i. Number of panels - How many panels are included in the dashboard.
+  ii. Visualization tools – what are the graphical representations of data. 
+  iii. Fitted on a single screen – whether the dashboard fits on a single screen or not (users can see the whole dashboard on a single screen without adjusting through grid overlay or not).
+  iv. Color theme – is there a unique color used for one data type in the whole dashboard (i.e.: one color scale for one data type everywhere on the dashboard). 
+  v. Dark background – the background color of the dashboard is dark or light.
+  vi. Data available – whether users can downloaded or whether data is available to reproduce the results. 
+  vii. Real time updated – whether the dashboard is updated daily/ specific time (live dashboard) or not.
 
@@
 
-As shown in Table 4 almost each & every dashboard, line charts & bar charts have been used to visualize the data.
-Heat map & dot plot has been used only one dashboard.
+As listed in Table 4, on almost each and every dashboard, line charts and bar charts have been used to visualize the data.
+The heat map and dot plot were only used on one dashboard.
 Only four dashboards have been fitted with a single screen.
-The majority of dashboards have used color theme on the whole dashboard.
-That is, dashboards have been applied different colors for different type of data (i.e.
-One specific color for confirmed cases, another color for deaths, etc.) in the whole dashboard.
-The data set & related links have available on some dashboards & users can download these data sets.
-There are 6 dashboards with dark background while others have been used light background.
-Last updated time & date of the latest available data  has been reported at the top or bottom of the first panel in the dashboard.
-Like, half of dashboards included all data in one panel.
+The majority of dashboards use a color theme on the whole dashboard.
+That is, dashboards have been applied with different colors for different types of data (i.e., one specific color for confirmed cases, another color for deaths, etc.) across the whole dashboard.
+The data set and related links are available on some dashboards, and users can download these data sets.
+There are 6 dashboards with a dark background, while others have a light background.
+The last updated time and date of the latest available data has been reported at the top or bottom of the first panel in the dashboard.
+Half of the dashboards included all the information in a single panel.
 
@@
 
-We obtained data from COVID-19 situation reports published by Epidemiology Unit, Ministry of Health Sri Lanka.
-We obtained data corresponds to number of death cases, number of hospitalized cases, number of recovered cases and COVID-19 vaccinated counts in Sri Lanka.
-The data are made available through an open-source R package covid19srilanka [@talagala].
+We obtained data from COVID-19 situation reports published by the Epidemiology Unit, Ministry of Health Sri Lanka.
+The data includes the number of death cases, number of hospitalized cases, number of recovered cases, and COVID-19 vaccinated counts in Sri Lanka.
+The data is made available through an open-source R package covid19srilanka [@talagala].
 

### 2022-05-15 manuscript_covid19dashboard.tex
Completed literature review

@@
 \hypersetup{
-  pdftitle={Interactive dashboard to monitor the COVID-19 outbreak and Vaccine Administration},
+  pdftitle={Interactive Dashboard to Monitor the COVID-19 Outbreak and Vaccine Administration},
   hidelinks,
@@
 
-\title{Interactive dashboard to monitor the COVID-19 outbreak and
+\title{Interactive Dashboard to Monitor the COVID-19 Outbreak and
 Vaccine Administration}
@@
 daily lives and work. Early responses and timely decisions and actions
-are critical to saving communities and economy worldwide. Data is
+are critical to saving communities and economies worldwide. Data is
 essential in order to make effective decisions. Data-driven information
@@
 provides front-line health officers a situational awareness of the
-spread of COVID-19 and status of the vaccination program.
+spread of COVID-19 and the status of the vaccination program.
 
@@
 
-Table 03 summarizes the visualization methods used in the dashboards. In
-almost every dashboard, value boxes have been used to represent these
-total figures. Bar charts and line charts (trend lines) are most
-frequently used to visualize this data (confirmed cases, recovered
-cases, deaths, and vaccination details) with respect to time. A majority
-of dashboards presented data either daily or weekly. The mapping is used
-to track the spatial distribution of COVID-19 cases by country,
-province, regional etc. When visualizing the data by the map color code
-system, circles with respect to the size of the cases have been used to
-visualize the variation in size. The considerable number of dashboards
-has been used in doughnut-shaped pie charts to represent total COVID-19
-confirmed cases, recovered cases, active cases, and deaths as a
-percentage. Also, region, gender, age group, and ethnicity can be
-identified as common breakdowns of COVID-19 cases. Data tables for
-representing case distribution by province/region have been added to
-some dashboards. Very few dashboards have been visualized in the
-COVID-19 test details. Only 6 dashboards have been compared to global
-situations. In addition, the fatality rate, incidence rate, ICU beds,
-stage of the patients, and hospitalized details have been contained in
-the several dashboards.\hfill\break
+Table 03 highlights the dashboard visualization techniques. Value boxes
+have been utilized to display total figures on practically every
+dashboard. The most common ways of visualizing confirmed cases,
+recovered cases, deaths, and immunization details are bar charts and
+line charts (trend lines). The majority of dashboards displayed data on
+a daily or weekly basis. The spatial distribution of COVID-19 cases by
+country, province, regional, and other factors is tracked using
+choropleth maps. When visualizing the data by the map color code system,
+circles with respect to the size of the cases have been used to
+visualize the variation in size. Several dashboards use doughnut-shaped
+pie charts to indicate total COVID-19 confirmed cases, recovered cases,
+active cases, and deaths as a proportion. Furthermore, region, gender,
+age group, and ethnicity can be identified as common breakdowns of
+COVID-19 cases. Data tables for representing cases' distribution by
+province/region have been added to some dashboards. Very few dashboards
+have been visualized in the COVID-19 test details. Only 6 dashboards
+have been compared to global situations. In addition, the fatality rate,
+incidence rate, ICU beds, stage of the patients, and hospitalized
+details have been contained in the several dashboards.\hfill\break
 
@@
 \hypertarget{comparison-of-dashboards}{%
-\subsection{2.2 Comparison of
-Dashboards}\label{comparison-of-dashboards}}
+\subsection{Comparison of Dashboards}\label{comparison-of-dashboards}}
 
 Before developing a dashboard, it is necessary to think about which
-visualization tools \& features that should be contained in the
-dashboards. What are the most suitable plots, how many panels in the
-dashboard, what data should be included, how to fit dashboard on a
-screen, colors \& is it real time updated or not are the common things
-that should be considered before develop the dashboards. In table 02 has
-been compared these visualization tools \& features under the following
-categories.
+visualization tools and features should be contained in the dashboard.
+What are the most suitable plots, how many panels in the dashboard, what
+data should be included, how to fit the dashboard on a screen, colors,
+and is it real time updated or not are the common things that should be
+considered before developing the dashboards. Table 4 summarizes
+information under the following categories:
 
-\begin{itemize}
+\begin{enumerate}
+\def\labelenumi{\roman{enumi}.}
 \tightlist
 \item
-  Number of panels - How many panels which are included in the
-  dashboard.
+  Number of panels - How many panels are included in the dashboard.
 \item
-  Visualization tools -- what are the graphical representations of data
-  which are contained in the dashboards.
+  Visualization tools -- what are the graphical representations of data.
 \item
@@
 \item
-  Dark background -- background color of the dashboard is dark or light.
+  Dark background -- the background color of the dashboard is dark or
+  light.
 \item
-  Data available -- whether users can be downloaded/available the data
-  set which has been visualized on the dashboard.
+  Data available -- whether users can downloaded or whether data is
+  available to reproduce the results.
 \item
@@
   time (live dashboard) or not.
-\end{itemize}
+\end{enumerate}
 
@@
 
-As shown in Table 4 almost each \& every dashboard, line charts \& bar
-charts have been used to visualize the data. Heat map \& dot plot has
-been used only one dashboard. Only four dashboards have been fitted with
-a single screen. The majority of dashboards have used color theme on the
-whole dashboard. That is, dashboards have been applied different colors
-for different type of data (i.e.~One specific color for confirmed cases,
-another color for deaths, etc.) in the whole dashboard. The data set \&
-related links have available on some dashboards \& users can download
-these data sets. There are 6 dashboards with dark background while
-others have been used light background. Last updated time \& date of the
-latest available data has been reported at the top or bottom of the
-first panel in the dashboard. Like, half of dashboards included all data
-in one panel.
+As listed in Table 4, on almost each and every dashboard, line charts
+and bar charts have been used to visualize the data. The heat map and
+dot plot were only used on one dashboard. Only four dashboards have been
+fitted with a single screen. The majority of dashboards use a color
+theme on the whole dashboard. That is, dashboards have been applied with
+different colors for different types of data (i.e., one specific color
+for confirmed cases, another color for deaths, etc.) across the whole
+dashboard. The data set and related links are available on some
+dashboards, and users can download these data sets. There are 6
+dashboards with a dark background, while others have a light background.
+The last updated time and date of the latest available data has been
+reported at the top or bottom of the first panel in the dashboard. Half
+of the dashboards included all the information in a single panel.
 
@@
 
-We obtained data from COVID-19 situation reports published by
-Epidemiology Unit, Ministry of Health Sri Lanka. We obtained data
-corresponds to number of death cases, number of hospitalized cases,
-number of recovered cases and COVID-19 vaccinated counts in Sri Lanka.
-The data are made available through an open-source R package
-covid19srilanka (Talagala 2021).
+We obtained data from COVID-19 situation reports published by the
+Epidemiology Unit, Ministry of Health Sri Lanka. The data includes the
+number of death cases, number of hospitalized cases, number of recovered
+cases, and COVID-19 vaccinated counts in Sri Lanka. The data is made
+available through an open-source R package covid19srilanka (Talagala
+2021).
 

### 2022-05-15 manuscript_covid19dashboard.Rmd
90% completed

@@
 The initial layout for the dashboard was prepared based on @rami.
-Data visualizations are generated using ggplot2 [@ggplot2] and plotly [@plotly] packages in R.
-we used color-blind friendly colour palettes for the graphics.
-To represent qualitative data, a diverging colour palette was used, and to represent numeric variables, a sequential colour theme was used.
+Data visualizations are generated using the ggplot2 [@ggplot2] and plotly [@plotly] packages in R.
+We used color-blind friendly color palettes for the graphics.
+A diverging colour palette was used to represent qualitative data, and to represent numeric variables, a sequential colour theme was used.
 Table 5 provides an overview of methods that have been used to visualize data.
@@
 
-We now describe the novel visualization approaches we included in out dashboard.
-To effectively distribute the vaccine and to support situational awareness and inform policy makers decision making it is important to know the district-wise spread of COVID-19 cases.
+We now describe the novel visualization approaches we included in our dashboard.
+To effectively distribute the vaccine and to support situational awareness and inform policymakers' decision-making, it is important to know the district-wise spread of COVID-19 cases.
  We have daily COVID-19 data related to confirmed cases in all 25 districts in Sri Lanka.
 This structure generates a multiple time series collection.
-Visualizing these time series data is useful to identify similarities and dissimilarities between districts and their general trends.
-There are two approaches to visualizing these time series: (i) drawing individual time series plots for each district (as shown in Figure 2), and (ii) simultaneously plots all time series on a single panel (as shown in Figure 3).
+Visualizing this time series data is useful to identify similarities and dissimilarities between districts and their general trends.
+There are two approaches to visualizing these time series: (i) creating individual time series plots for each district (as shown in Figure 1-A), and (ii) plotting all time series on a single panel at the same time (as shown in Figure 1-B).
 Plotting all time series simultaneously is also not possible due to overlapping time series and scale differences.
@@
 The reason is that it is hard to compare across 25 different panels at once.
-In order to overcome these problems in multiple time series visualization,  we use heat maps ((@peng2008method)) to visualize global and local similarities and dissimilarities across districts.
-The associated results are shown in Figure 3.
-Here two heatmaps are used to show the global variations (Figure 3: A) and local variation (Figure 3: B) in the time series collection.
-Figure 3-A cell colours represent the actual counts of the COVID-19 confirmed cases.
-This is useful to get an idea about the differences in absolute values.
-Figure 3-B cell colours represent the normalized values created applying the min-max transformation.
-Min-max transformation is applied to each district time series by using the corresponding district minimum and maximum value of the time series.
+In order to overcome these problems in multiple time series visualization, we use heat maps (@peng2008method) to visualize global and local similarities and dissimilarities across districts.
+The associated results are shown in Figure 2.
+Here, two heatmaps are used to show the global variations (Figure 2: A) and local variations (Figure 2: B) in the time series collection.
+Figure 2-A cell colours represent the actual counts of the COVID-19 confirmed cases.
+This is useful to get an idea of the differences in absolute values.
+Figure 2-B cell colors represent the normalized values created by applying the min-max transformation.
+A Min-max transformation is applied to each district's time series by using the corresponding district's minimum and maximum value.
 This helps us to get an idea about patterns within districts.
-For example, according to Figure 3A we can see Colombo, Gampaha and Kalutara districs COVID-19 cases are significantly higher than other districts.
-According to Figure 3B all districts show an increasing trend pattern as the right-hand side of the cells are lighter than the left-hand side cells in the heat map.
-Furthermore according to Figure 3B all districts reported high number of cases on  19, 24, 29 August, 2021.
-Figure 3B is useful for identifying these local outlying behaviours.
-As shown in Figure 4, we also use Choropleth map and Dorling cartogram to visualize spatial distribution of COVID-19 cases.
-The vaccination information are visualize through interactive time series plots and bar charts.
-A screenshot of associated panels are shown in Figure 5 and Figure 6.
+For example, according to Figure 2A, we can see that in Colombo, Gampaha, and Kalutara districts, COVID-19 cases are significantly higher than in other districts.
+According to Figure 2B, all districts show an increasing trend pattern as the right-hand side of the cells are lighter than the left-hand side cells in the heat map.
+Furthermore, according to Figure 2B, all districts reported a high number of cases on August 19, 24, and 29, 2021.
+Figure 3B is useful for identifying these local outlying behaviors.
+As shown in Figure 4, we also use a Choropleth map and Dorling cartogram to visualize the spatial distribution of COVID-19 cases.
+The vaccination information is visualized through interactive time series plots and bar charts.
 
@@
 
-| **Name of the Panel**|**Description of the Panel**| 
-|:------|:----------------------------------------------|
-|Overview          |Total count of COVID-19 confirmed, recovered, deaths, active cases & total vaccine doses administered.\hfill\break Provide an overview of daily COVID-19 confirmed, recovered & deaths by plots.\hfill\break|
-|Cases by Wave| There are three tabs in this panel.\hfill\break  * Total COVID-19 confirmed cases - Cumulative count of COVID-19 confirmed cases with specific dates\hfill\break * COVID-19 Cases Distribution by Wave - Daily confirmed cases by wave\hfill\break * COVID-19 Deaths Distribution by Wave - Daily deaths by wave\hfill\break|
+| **Name of the Panel**|**Description of the Panel**| **Figure** |
+|:------|:----------------------------------------------|:----|
+|Overview          |Total count of COVID-19 confirmed, recovered, deaths, active cases and total vaccine doses administered.\hfill\break Provide an overview of daily COVID-19 confirmed, recovered & deaths by plots.\hfill\break| Figure 4 |
+|Cases by Wave| There are three tabs in this panel.\hfill\break  * Total COVID-19 confirmed cases - Cumulative count of COVID-19 confirmed cases with specific dates\hfill\break * COVID-19 Cases Distribution by Wave - Daily confirmed cases by wave\hfill\break * COVID-19 Deaths Distribution by Wave - Daily deaths by wave\hfill\break| Figure 5 |
 |COVID-19 Patients Distribution | Provide an overview of confirmed cases district wise distribution.
-There are four tabs in this panel.\hfill\break * Total COVID-19 Patients Distribution in Sri Lanka - Total confirmed counts for each district is represented by tree map\hfill\break * Country Map - Total confirmed cases in each district represented by Sri Lanka country map\hfill\break * Distribution of Daily COVID-19 Patients for Last 30 Days - Visualize the daily confirmed cases distribution by districts in last 30 days\hfill\break * By Applying Min-Max Transformation - Visualize the details in the third tab by applying min-max transformation for each district\hfill\break|
+There are four tabs in this panel.\hfill\break * Total COVID-19 Patients Distribution in Sri Lanka - Total confirmed counts for each district is represented by tree map\hfill\break * Country Map - Total confirmed cases in each district represented by Sri Lanka country map\hfill\break * Distribution of Daily COVID-19 Patients for Last 30 Days - Visualize the daily confirmed cases distribution by districts in last 30 days\hfill\break * By Applying Min-Max Transformation - Visualize the details in the third tab by applying min-max transformation for each district\hfill\break| Figure 6 |
 |Vaccination Details | Provide an overview of COVID-19 vaccination in Sri Lanka.
-There are two tabs.\hfill\break * Total Vaccine Doses - Visualize the administered vaccine doses as first dose only & fully vaccinated\hfill\break * Total Administered Doses by Vaccine Name - Visualize the vaccination by vaccine names\hfill\break|
-| Top 10 Countries| In this panel, compare the Sri Lanka confirmed & deaths with top 10 countries in the world (top 10 countries - The countries which have been reported highest number of confirmed cases as 31st of August 2021).\hfill\break There are two tabs.\hfill\break * Comparison of Cumulative Cases in Sri Lanka with Top 10 Countries - Compare the confirmed and deaths in Sri Lanka with top 10 countries by cumulative time series plots\hfill\break * Comparison of Log of Cumulative Cases in Sri Lanka with Top 10 Countries -  Compare the confirmed and deaths in Sri Lanka with top 10 countries by log cumulative time series plots\hfill\break (The data has been pulled from WHO)\hfill\break|
+There are two tabs.\hfill\break * Total Vaccine Doses - Visualize the administered vaccine doses as first dose only & fully vaccinated\hfill\break * Total Administered Doses by Vaccine Name - Visualize the vaccination by vaccine names\hfill\break| Figure 7 |
+| Top 10 Countries| In this panel, compare the Sri Lanka confirmed & deaths with top 10 countries in the world (top 10 countries - The countries which have been reported highest number of confirmed cases as 31st of August 2021).\hfill\break There are two tabs.\hfill\break * Comparison of Cumulative Cases in Sri Lanka with Top 10 Countries - Compare the confirmed and deaths in Sri Lanka with top 10 countries by cumulative time series plots\hfill\break * Comparison of Log of Cumulative Cases in Sri Lanka with Top 10 Countries -  Compare the confirmed and deaths in Sri Lanka with top 10 countries by log cumulative time series plots\hfill\break (The data has been pulled from WHO)\hfill\break| Figure 8 |
 |Global Comparison | Compare the total confirmed & deaths in Sri lanka with top 10 countries in Global & Asia.
-There are two tabs.\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the World - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the world & compare the case fatality ratios\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the Asia - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the Asia & compare the case fatality ratios\hfill\break|
-|Global Map| Visualize the distribution of confirmed, recovered & deaths in the world by world map.\hfill\break|
-|About | This panel contains the details about the dashboard.|
+There are two tabs.\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the World - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the world & compare the case fatality ratios\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the Asia - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the Asia & compare the case fatality ratios\hfill\break| Figure 9 |
+|Global Map| Visualize the distribution of confirmed, recovered & deaths in the world by world map.\hfill\break| Figure 10 |
+|About | This panel contains the details about the dashboard.| Figure 11 |
 
@@
 
-Bar charts and line charts are the most frequently used tools for the visualization of total cases, daily cases and comparisons with respect to time.
-Some dashboards contained doughnut shape pie charts to summarize the total figures.
-In almost each and every dashboard, value boxes have been used to represent total figures.
-Some dashboards contained interactive maps & data tables to visualize the distribution of cases by country, province, region or state.
-All dashboards are daily updating real time dashboards.
-Gender, age groups and ethnicity can be identified as common breakdowns.
-The data sets & related links are available on most of the dashboards & can be downloaded.
-It can be seen that it is very easy, clear and user friendly to identify confirmed, recovered and deceased cases in dashboards which include one color theme for the whole dashboard.
-Dashboards with the dark background are more comfortable to the eyes than dashboards with light background & light colors.
-Also, it is better if the dashboard can be fitted on a single screen rather than adjusting through a grid overlay.
+Bar charts and line charts are the most frequently used tools for the visualization of total cases, daily cases, and weekly cases.
+and comparisons with respect to time. Some dashboards contained doughnut-shaped pie charts to summarize
+the total figures. In almost every dashboard, value boxes have been used to represent total figures.
+Some dashboards contained interactive maps and data tables to visualize the distribution of cases by country,
+province, region, or state. All dashboards are updated daily in real time. Gender, age groups, and ethnicity can be identified as common breakdowns. Most data sets and related links are available.
+
+When building our dashboard we use a colour-blind friendly theme.
+Our dashboard is fully reproducible.
+The source code to reprouduce the results are available in a public GitHub repository and can be accessonline via https://github.com/thiyangt/covid19srilanka
 

### 2022-05-15 manuscript_covid19dashboard.tex
90% completed

@@
 prepared based on Krispin (2021). Data visualizations are generated
-using ggplot2 (\textbf{ggplot2?}) and plotly (Sievert 2020) packages in
-R. we used color-blind friendly colour palettes for the graphics. To
-represent qualitative data, a diverging colour palette was used, and to
+using the ggplot2 (\textbf{ggplot2?}) and plotly (Sievert 2020) packages
+in R. We used color-blind friendly color palettes for the graphics. A
+diverging colour palette was used to represent qualitative data, and to
 represent numeric variables, a sequential colour theme was used. Table 5
@@
 
-We now describe the novel visualization approaches we included in out
+We now describe the novel visualization approaches we included in our
 dashboard. To effectively distribute the vaccine and to support
-situational awareness and inform policy makers decision making it is
-important to know the district-wise spread of COVID-19 cases. We have
+situational awareness and inform policymakers' decision-making, it is
+important to know the district-wise spread of COVID-19 cases. ~We have
 daily COVID-19 data related to confirmed cases in all 25 districts in
 Sri Lanka. This structure generates a multiple time series collection.
-Visualizing these time series data is useful to identify similarities
-and dissimilarities between districts and their general trends. There
-are two approaches to visualizing these time series: (i) drawing
-individual time series plots for each district (as shown in Figure 2),
-and (ii) simultaneously plots all time series on a single panel (as
-shown in Figure 3). Plotting all time series simultaneously is also not
+Visualizing this time series data is useful to identify similarities and
+dissimilarities between districts and their general trends. There are
+two approaches to visualizing these time series: (i) creating individual
+time series plots for each district (as shown in Figure 1-A), and (ii)
+plotting all time series on a single panel at the same time (as shown in
+Figure 1-B). Plotting all time series simultaneously is also not
 possible due to overlapping time series and scale differences. Plotting
@@
 overcome these problems in multiple time series visualization, we use
-heat maps ((Peng (2008))) to visualize global and local similarities and
+heat maps (Peng (2008)) to visualize global and local similarities and
 dissimilarities across districts. The associated results are shown in
-Figure 3. Here two heatmaps are used to show the global variations
-(Figure 3: A) and local variation (Figure 3: B) in the time series
-collection. Figure 3-A cell colours represent the actual counts of the
-COVID-19 confirmed cases. This is useful to get an idea about the
-differences in absolute values. Figure 3-B cell colours represent the
-normalized values created applying the min-max transformation. Min-max
-transformation is applied to each district time series by using the
-corresponding district minimum and maximum value of the time series.
-This helps us to get an idea about patterns within districts. For
-example, according to Figure 3A we can see Colombo, Gampaha and Kalutara
-districs COVID-19 cases are significantly higher than other districts.
-According to Figure 3B all districts show an increasing trend pattern as
-the right-hand side of the cells are lighter than the left-hand side
-cells in the heat map. Furthermore according to Figure 3B all districts
-reported high number of cases on 19, 24, 29 August, 2021. Figure 3B is
-useful for identifying these local outlying behaviours. As shown in
-Figure 4, we also use Choropleth map and Dorling cartogram to visualize
-spatial distribution of COVID-19 cases. The vaccination information are
-visualize through interactive time series plots and bar charts. A
-screenshot of associated panels are shown in Figure 5 and Figure 6.
+Figure 2. Here, two heatmaps are used to show the global variations
+(Figure 2: A) and local variations (Figure 2: B) in the time series
+collection. Figure 2-A cell colours represent the actual counts of the
+COVID-19 confirmed cases. This is useful to get an idea of the
+differences in absolute values. Figure 2-B cell colors represent the
+normalized values created by applying the min-max transformation. A
+Min-max transformation is applied to each district's time series by
+using the corresponding district's minimum and maximum value. This helps
+us to get an idea about patterns within districts. For example,
+according to Figure 2A, we can see that in Colombo, Gampaha, and
+Kalutara districts, COVID-19 cases are significantly higher than in
+other districts. According to Figure 2B, all districts show an
+increasing trend pattern as the right-hand side of the cells are lighter
+than the left-hand side cells in the heat map. Furthermore, according to
+Figure 2B, all districts reported a high number of cases on August 19,
+24, and 29, 2021. Figure 3B is useful for identifying these local
+outlying behaviors. As shown in Figure 4, we also use a Choropleth map
+and Dorling cartogram to visualize the spatial distribution of COVID-19
+cases. The vaccination information is visualized through interactive
+time series plots and bar charts.
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.13}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.87}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.12}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.80}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.08}}@{}}
 \toprule
-\textbf{Name of the Panel} & \textbf{Description of the Panel} \\
+\textbf{Name of the Panel} & \textbf{Description of the Panel} &
+\textbf{Figure} \\
 \midrule
@@
 Overview & Total count of COVID-19 confirmed, recovered, deaths, active
-cases \& total vaccine doses administered.\hfill\break Provide an
+cases and total vaccine doses administered.\hfill\break Provide an
 overview of daily COVID-19 confirmed, recovered \& deaths by
-plots.\hfill\break \\
+plots.\hfill\break & Figure 4 \\
 Cases by Wave & There are three tabs in this panel.\hfill\break  * Total
@@
 Daily confirmed cases by wave\hfill\break * COVID-19 Deaths Distribution
-by Wave - Daily deaths by wave\hfill\break \\
+by Wave - Daily deaths by wave\hfill\break & Figure 5 \\
 COVID-19 Patients Distribution & Provide an overview of confirmed cases
@@
 Min-Max Transformation - Visualize the details in the third tab by
-applying min-max transformation for each district\hfill\break \\
+applying min-max transformation for each district\hfill\break & Figure
+6 \\
 Vaccination Details & Provide an overview of COVID-19 vaccination in Sri
@@
 vaccinated\hfill\break * Total Administered Doses by Vaccine Name -
-Visualize the vaccination by vaccine names\hfill\break \\
+Visualize the vaccination by vaccine names\hfill\break & Figure 7 \\
 Top 10 Countries & In this panel, compare the Sri Lanka confirmed \&
@@
 cumulative time series plots\hfill\break (The data has been pulled from
-WHO)\hfill\break \\
+WHO)\hfill\break & Figure 8 \\
 Global Comparison & Compare the total confirmed \& deaths in Sri lanka
@@
 Asia - Compare the total confirmed \& deaths in Sri Lanka with top 10
-countries in the Asia \& compare the case fatality ratios\hfill\break \\
+countries in the Asia \& compare the case fatality ratios\hfill\break &
+Figure 9 \\
 Global Map & Visualize the distribution of confirmed, recovered \&
-deaths in the world by world map.\hfill\break \\
-About & This panel contains the details about the dashboard. \\
+deaths in the world by world map.\hfill\break & Figure 10 \\
+About & This panel contains the details about the dashboard. & Figure
+11 \\
 \bottomrule
@@
 Bar charts and line charts are the most frequently used tools for the
-visualization of total cases, daily cases and comparisons with respect
-to time. Some dashboards contained doughnut shape pie charts to
-summarize the total figures. In almost each and every dashboard, value
-boxes have been used to represent total figures. Some dashboards
-contained interactive maps \& data tables to visualize the distribution
-of cases by country, province, region or state. All dashboards are daily
-updating real time dashboards. Gender, age groups and ethnicity can be
-identified as common breakdowns. The data sets \& related links are
-available on most of the dashboards \& can be downloaded. It can be seen
-that it is very easy, clear and user friendly to identify confirmed,
-recovered and deceased cases in dashboards which include one color theme
-for the whole dashboard. Dashboards with the dark background are more
-comfortable to the eyes than dashboards with light background \& light
-colors. Also, it is better if the dashboard can be fitted on a single
-screen rather than adjusting through a grid overlay.
+visualization of total cases, daily cases, and weekly cases. and
+comparisons with respect to time. Some dashboards contained
+doughnut-shaped pie charts to summarize the total figures. In almost
+every dashboard, value boxes have been used to represent total figures.
+Some dashboards contained interactive maps and data tables to visualize
+the distribution of cases by country, province, region, or state. All
+dashboards are updated daily in real time. Gender, age groups, and
+ethnicity can be identified as common breakdowns. Most data sets and
+related links are available.
+
+When building our dashboard we use a colour-blind friendly theme. Our
+dashboard is fully reproducible. The source code to reprouduce the
+results are available in a public GitHub repository and can be
+accessonline via \url{https://github.com/thiyangt/covid19srilanka}
 

### 2022-05-15 manuscript_covid19dashboard.Rmd
Completed paper

@@
 
-As of September 20th, 2021, 221 countries and territories are infected by the COVID-19 worldwide pandemic.
-Dashboards are the most often used visualization method for visualizing COVID-19 data and informing the public.
-We explored 15 different dashboards.
+Dashboards are the most common visualization method for displaying COVID-19 data and informing the public.
+We examined 15 different dashboards to see how various visualization techniques were used.
+This paper describes the creation and implementation of a dashboard for COVID-19 epidemic and vaccination administration data in Sri Lanka.
 
@@
 
-**Table 01: Labeled of the Dashboards**
+**Table 01: Labels of the dashboards**
 
@@
 
-**Table 04: Comparison of visualization tools & features of dashboard**
+**Table 04: Comparison of visualization tools and features of dashboard**
 
@@
 
-```{r, comment=NA, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of COVID19 Cases by Districts: (A) Plotted on a Single Panel, (B) Plotted on Separate Panels.", fig.height=12, fig.width=10}
+```{r, comment=NA, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of COVID-19 cases by districts: (A) Plotted on a single panel, (B) Plotted on separate panels.", fig.height=12, fig.width=10}
 library(covid19srilanka)
@@
 
-```{r, echo=FALSE,fig.cap="(A) Global View of Distribution of COVID19 Cases by DIstricts, (B) Local View of Distribution of COVID19 Cases by Districts", fig.width=10, fig.height=10}
+```{r, echo=FALSE,fig.cap="(A) Global view of distribution of COVID-19 cases by districts, (B) Local view of distribution of COVID-19 cases by districts", fig.width=10, fig.height=10}
 
@@
 
-```{r, comment=NA, message=FALSE, warning=FALSE, fig.cap="Spatial DIstribution of COVID1-19 Cases by Districts", fig.height=8}
+```{r, comment=NA, message=FALSE, warning=FALSE, fig.cap="Spatial distribution of COVID1-19 cases by districts", fig.height=8}
 
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 1: Overview", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of panel 1: Overview", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image1.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 2: Cases by Wave", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of panel 2: Cases by Wave", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image2.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 3: Distribution of COVID-19 Patients", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of panel 3: Distribution of COVID-19 Patients", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image3.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 4: Vaccination Details", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of panel 4: Vaccination Details", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image4.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 5: Comparison with Top 10 Countries", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of panel 5: Comparison with Top 10 Countries", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image5.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 6: Global Comparison", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of Panel 6: Global comparison", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image6.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 7: World Map", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of panel 7: World Map", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image7.png")
@@
 
-```{r, echo=FALSE, fig.cap="Screeshot of Panel 8: About", out.width="80%", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Screeshot of panel 8: About", out.width="80%", fig.align='center'}
 knitr::include_graphics("Images/image8.png")
@@
 
-When building our dashboard we use a colour-blind friendly theme.
-Our dashboard is fully reproducible.
-The source code to reprouduce the results are available in a public GitHub repository and can be accessonline via https://github.com/thiyangt/covid19srilanka
+We used a color-blind-friendly theme when creating our dashboard.
+The dashboard includes both static and interactive charts that can be used for explanation, exploration, and display.
+The data speaks to us, but it is not always easy to understand in what language they are doing it.
+Hence, we examine the same data under different angles using different types of plots.
+We do not have district-wise vaccination details.
+We do not have vaccination information per district.
+We'll save this for later.
+Our dashboard is completely replicable.
+The source code for reproducing the results is accessible in a public GitHub repository at https://github.com/thiyangt/covid19srilanka.
 

### 2022-05-15 manuscript_covid19dashboard.tex
Completed paper

@@
 
-As of September 20th, 2021, 221 countries and territories are infected
-by the COVID-19 worldwide pandemic. Dashboards are the most often used
-visualization method for visualizing COVID-19 data and informing the
-public. We explored 15 different dashboards.
+Dashboards are the most common visualization method for displaying
+COVID-19 data and informing the public. We examined 15 different
+dashboards to see how various visualization techniques were used. This
+paper describes the creation and implementation of a dashboard for
+COVID-19 epidemic and vaccination administration data in Sri Lanka.
 
@@
 
-\textbf{Table 01: Labeled of the Dashboards}
+\textbf{Table 01: Labels of the dashboards}
 
@@
 
-\textbf{Table 04: Comparison of visualization tools \& features of
+\textbf{Table 04: Comparison of visualization tools and features of
 dashboard}
@@
 \includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{Distribution of COVID19 Cases by Districts: (A) Plotted on a
-Single Panel, (B) Plotted on Separate Panels.}
+\caption{Distribution of COVID-19 cases by districts: (A) Plotted on a
+single panel, (B) Plotted on separate panels.}
 \end{figure}
@@
 \includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-3-1.pdf}
-\caption{(A) Global View of Distribution of COVID19 Cases by DIstricts,
-(B) Local View of Distribution of COVID19 Cases by Districts}
+\caption{(A) Global view of distribution of COVID-19 cases by districts,
+(B) Local view of distribution of COVID-19 cases by districts}
 \end{figure}
@@
 \includegraphics{manuscript_covid19dashboard_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{Spatial DIstribution of COVID1-19 Cases by Districts}
+\caption{Spatial distribution of COVID1-19 cases by districts}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 1: Overview}\label{fig:unnamed-chunk-5}
+\caption{Screeshot of panel 1: Overview}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 2: Cases by Wave}\label{fig:unnamed-chunk-6}
+\caption{Screeshot of panel 2: Cases by Wave}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 3: Distribution of COVID-19 Patients}\label{fig:unnamed-chunk-7}
+\caption{Screeshot of panel 3: Distribution of COVID-19 Patients}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 4: Vaccination Details}\label{fig:unnamed-chunk-8}
+\caption{Screeshot of panel 4: Vaccination Details}\label{fig:unnamed-chunk-8}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 5: Comparison with Top 10 Countries}\label{fig:unnamed-chunk-9}
+\caption{Screeshot of panel 5: Comparison with Top 10 Countries}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 6: Global Comparison}\label{fig:unnamed-chunk-10}
+\caption{Screeshot of Panel 6: Global comparison}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 7: World Map}\label{fig:unnamed-chunk-11}
+\caption{Screeshot of panel 7: World Map}\label{fig:unnamed-chunk-11}
 \end{figure}
@@
 
-\caption{Screeshot of Panel 8: About}\label{fig:unnamed-chunk-12}
+\caption{Screeshot of panel 8: About}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 
-When building our dashboard we use a colour-blind friendly theme. Our
-dashboard is fully reproducible. The source code to reprouduce the
-results are available in a public GitHub repository and can be
-accessonline via \url{https://github.com/thiyangt/covid19srilanka}
+We used a color-blind-friendly theme when creating our dashboard. The
+dashboard includes both static and interactive charts that can be used
+for explanation, exploration, and display. The data speaks to us, but it
+is not always easy to understand in what language they are doing it.
+Hence, we examine the same data under different angles using different
+types of plots. We do not have district-wise vaccination details. We do
+not have vaccination information per district. We'll save this for
+later. Our dashboard is completely replicable. The source code for
+reproducing the results is accessible in a public GitHub repository at
+\url{https://github.com/thiyangt/covid19srilanka}.
 

### 2022-05-15 manuscript_covid19dashboard.Rmd
Updated

@@
 
-**Thiyanga S. Talagala**^[Corresponding author, email: ttalagala@gmail.com]
+**Thiyanga S. Talagala**^[Corresponding author, email: ttalagala@sjp.ac.lk]
 
@@
 
-**Keywords:** Visualization, multiple time series, heatmap, COVID-19 vaccine, flexdashboard
+**Keywords:** Visualization, multiple time series, heat map, COVID-19 vaccine, flexdashboard
 
@@
 Massive amounts of data are being generated in the response to the COVID-19 pandemic.
-Given this available data, it is critical to create tools for exploratory analysis for policymakers, health officials, and the general public.
-Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemic's spread and vaccine administration.
+Given this available data, it is critical to create tools for exploratory analysis for policy-makers, health officials, and the general public.
+Dashboards are one of the greatest visual interpretation methods for tracking the COVID-19 pandemics spread and vaccine administration.
 Dashboards allow users to quickly interact with a combination of exploratory visualizations and gain a quick overview of the data.
@@
 
-There are a plethora of COVID-19 visualization dashboards that have been designed to visualize the pandemic's global and local status.
+There are a plethora of COVID-19 visualization dashboards that have been designed to visualize the pandemics global and local status.
 Different software can be used to generate dashboards.
@@
 |:--:|:-------------------|------|
-|1 | COVID-19 dashboard created by the John Hopkins University Center for Systems Science & Engineering (JHU CSSE) |@jh|
+|1 | COVID-19 dashboard created by the John Hopkins University Centre for Systems Science & Engineering (JHU CSSE) |@jh|
 |2 | WHO  COVID-19 Dashboard |@who|
@@
 The spatial distribution of COVID-19 cases by country, province, regional, and other factors is tracked using choropleth maps.
-When visualizing the data by the map color code system, circles with respect to the size of the cases have been used to visualize the variation in size.
+When visualizing the data by the map colour code system, circles with respect to the size of the cases have been used to visualize the variation in size.
 Several dashboards use doughnut-shaped pie charts to indicate total COVID-19 confirmed cases, recovered cases, active cases, and deaths as a proportion.
@@
 Before developing a dashboard, it is necessary to think about which visualization tools and features should be contained in the dashboard.
-What are the most suitable plots, how many panels in the dashboard, what data should be included, how to fit the dashboard on a screen, colors, and is it real time updated or not are the common things that should be considered before developing the dashboards.
+What are the most suitable plots, how many panels in the dashboard, what data should be included, how to fit the dashboard on a screen, colours, and is it real time updated or not are the common things that should be considered before developing the dashboards.
 Table 4 summarizes information under the following categories:
@@
   iii. Fitted on a single screen – whether the dashboard fits on a single screen or not (users can see the whole dashboard on a single screen without adjusting through grid overlay or not).
-  iv. Color theme – is there a unique color used for one data type in the whole dashboard (i.e.: one color scale for one data type everywhere on the dashboard). 
-  v. Dark background – the background color of the dashboard is dark or light.
+  iv. Colour theme – is there a unique colour used for one data type in the whole dashboard (i.e.: one colour scale for one data type everywhere on the dashboard). 
+  v. Dark background – the background colour of the dashboard is dark or light.
   vi. Data available – whether users can downloaded or whether data is available to reproduce the results. 
@@
 
-| Name of the Dashboard | Number of panels|  Visualization tools | Fitted on a single screen | Color theme | Dark background | Data available | Real time updated|
+| Name of the Dashboard | Number of panels|  Visualization tools | Fitted on a single screen | Colour theme | Dark background | Data available | Real time updated|
 |:-------: |:-------: |:-------:  |:-------:|:-------:   |:-------:  |:-------:   |:-------:  |
@@
 Only four dashboards have been fitted with a single screen.
-The majority of dashboards use a color theme on the whole dashboard.
-That is, dashboards have been applied with different colors for different types of data (i.e., one specific color for confirmed cases, another color for deaths, etc.) across the whole dashboard.
+The majority of dashboards use a colour theme on the whole dashboard.
+That is, dashboards have been applied with different colours for different types of data (i.e., one specific colour for confirmed cases, another colour for deaths, etc.) across the whole dashboard.
 The data set and related links are available on some dashboards, and users can download these data sets.
@@
 Data visualizations are generated using the ggplot2 [@ggplot2] and plotly [@plotly] packages in R.
-We used color-blind friendly color palettes for the graphics.
+We used colour-blind friendly colour palettes for the graphics.
 A diverging colour palette was used to represent qualitative data, and to represent numeric variables, a sequential colour theme was used.
@@
 |Total COVID-19 death cases by time and wave         | Histogram \hfill\break
-|Distribution of COVID-19 patients by districts        | Tree map, Choropleth maps, Dorling Cartogram, heatmaps \hfill\break
+|Distribution of COVID-19 patients by districts        | Tree map, Choropleth maps, Dorling Cartogram, heat maps \hfill\break
 |Total vaccination by first dose and second dose        | Time series plot \hfill\break
@@
 |Total administrated does by vaccine name       | Stacked bar chart \hfill\break
-|Comparison of cases with in Sri Lanka with Top 10 countries       | Cummulative cases by time, Log of cummulative cases by time, stacked bar chart \hfill\break
+|Comparison of cases with in Sri Lanka with Top 10 countries       | Cumulative cases by time, Log of cumulative cases by time, stacked bar chart \hfill\break
 |Spread of COVID-19 around the world      | Choropleth maps \hfill\break
@@
 We now describe the novel visualization approaches we included in our dashboard.
-To effectively distribute the vaccine and to support situational awareness and inform policymakers' decision-making, it is important to know the district-wise spread of COVID-19 cases.
+To effectively distribute the vaccine and to support situational awareness and inform policy-makers' decision-making, it is important to know the district-wise spread of COVID-19 cases.
  We have daily COVID-19 data related to confirmed cases in all 25 districts in Sri Lanka.
@@
 The associated results are shown in Figure 2.
-Here, two heatmaps are used to show the global variations (Figure 2: A) and local variations (Figure 2: B) in the time series collection.
+Here, two heat maps are used to show the global variations (Figure 2: A) and local variations (Figure 2: B) in the time series collection.
 Figure 2-A cell colours represent the actual counts of the COVID-19 confirmed cases.
 This is useful to get an idea of the differences in absolute values.
-Figure 2-B cell colors represent the normalized values created by applying the min-max transformation.
-A Min-max transformation is applied to each district's time series by using the corresponding district's minimum and maximum value.
+Figure 2-B cell colours represent the normalized values created by applying the min-max transformation.
+A min-max transformation is applied to each district's time series by using the corresponding district's minimum and maximum value.
 This helps us to get an idea about patterns within districts.
@@
 Furthermore, according to Figure 2B, all districts reported a high number of cases on August 19, 24, and 29, 2021.
-Figure 3B is useful for identifying these local outlying behaviors.
+Figure 3B is useful for identifying these local outlying behaviours.
 As shown in Figure 4, we also use a Choropleth map and Dorling cartogram to visualize the spatial distribution of COVID-19 cases.
@@
 
-The source code to reprouduce the results are available in a public GitHub repository and can be access online via https://github.com/thiyangt/covid19srilanka
-
 # Results {#results}
 
-The "Sri Lanka COVID-19 Dashboard" provides an overview of the COVID-19 pandemic and administration of vaacine information in Sri Lanka.  This dashboard has eight panels as listed in Table 6.
+The "Sri Lanka COVID-19 Dashboard" provides an overview of the COVID-19 pandemic and administration of vaccine information in Sri Lanka.  This dashboard has eight panels as listed in Table 6.
 
@@
 | Top 10 Countries| In this panel, compare the Sri Lanka confirmed & deaths with top 10 countries in the world (top 10 countries - The countries which have been reported highest number of confirmed cases as 31st of August 2021).\hfill\break There are two tabs.\hfill\break * Comparison of Cumulative Cases in Sri Lanka with Top 10 Countries - Compare the confirmed and deaths in Sri Lanka with top 10 countries by cumulative time series plots\hfill\break * Comparison of Log of Cumulative Cases in Sri Lanka with Top 10 Countries -  Compare the confirmed and deaths in Sri Lanka with top 10 countries by log cumulative time series plots\hfill\break (The data has been pulled from WHO)\hfill\break| Figure 8 |
-|Global Comparison | Compare the total confirmed & deaths in Sri lanka with top 10 countries in Global & Asia.
-There are two tabs.\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the World - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the world & compare the case fatality ratios\hfill\break * Comparison of the sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the Asia - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the Asia & compare the case fatality ratios\hfill\break| Figure 9 |
+|Global Comparison | Compare the total confirmed & deaths in Sri Lanka with top 10 countries in Global & Asia.
+There are two tabs.\hfill\break * Comparison of the Sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the World - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the world & compare the case fatality ratios\hfill\break * Comparison of the Sri Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the Asia - Compare the total confirmed & deaths in Sri Lanka with top 10 countries in the Asia & compare the case fatality ratios\hfill\break| Figure 9 |
 |Global Map| Visualize the distribution of confirmed, recovered & deaths in the world by world map.\hfill\break| Figure 10 |
@@
 
-We used a color-blind-friendly theme when creating our dashboard.
+We used a colour-blind-friendly theme when creating our dashboard.
 The dashboard includes both static and interactive charts that can be used for explanation, exploration, and display.

### 2022-05-15 manuscript_covid19dashboard.tex
Updated

@@
 \textbf{Thiyanga S. Talagala}\footnote{Corresponding author, email:
-  \href{mailto:ttalagala@gmail.com}{\nolinkurl{ttalagala@gmail.com}}}
+  \href{mailto:ttalagala@sjp.ac.lk}{\nolinkurl{ttalagala@sjp.ac.lk}}}
 
@@
 
-\textbf{Keywords:} Visualization, multiple time series, heatmap,
+\textbf{Keywords:} Visualization, multiple time series, heat map,
 COVID-19 vaccine, flexdashboard
@@
 COVID-19 pandemic. Given this available data, it is critical to create
-tools for exploratory analysis for policymakers, health officials, and
+tools for exploratory analysis for policy-makers, health officials, and
 the general public. Dashboards are one of the greatest visual
-interpretation methods for tracking the COVID-19 pandemic's spread and
+interpretation methods for tracking the COVID-19 pandemics spread and
 vaccine administration. Dashboards allow users to quickly interact with
@@
 There are a plethora of COVID-19 visualization dashboards that have been
-designed to visualize the pandemic's global and local status. Different
+designed to visualize the pandemics global and local status. Different
 software can be used to generate dashboards. We explored 15 dashboards
@@
 \endhead
-1 & COVID-19 dashboard created by the John Hopkins University Center for
+1 & COVID-19 dashboard created by the John Hopkins University Centre for
 Systems Science \& Engineering (JHU CSSE) & {``{COVID-19 Map -- Johns
@@
 country, province, regional, and other factors is tracked using
-choropleth maps. When visualizing the data by the map color code system,
-circles with respect to the size of the cases have been used to
+choropleth maps. When visualizing the data by the map colour code
+system, circles with respect to the size of the cases have been used to
 visualize the variation in size. Several dashboards use doughnut-shaped
@@
 What are the most suitable plots, how many panels in the dashboard, what
-data should be included, how to fit the dashboard on a screen, colors,
+data should be included, how to fit the dashboard on a screen, colours,
 and is it real time updated or not are the common things that should be
@@
 \item
-  Color theme -- is there a unique color used for one data type in the
-  whole dashboard (i.e.: one color scale for one data type everywhere on
-  the dashboard).
+  Colour theme -- is there a unique colour used for one data type in the
+  whole dashboard (i.e.: one colour scale for one data type everywhere
+  on the dashboard).
 \item
-  Dark background -- the background color of the dashboard is dark or
+  Dark background -- the background colour of the dashboard is dark or
   light.
@@
 Name of the Dashboard & Number of panels & Visualization tools & Fitted
-on a single screen & Color theme & Dark background & Data available &
+on a single screen & Colour theme & Dark background & Data available &
 Real time updated \\
@@
 dot plot were only used on one dashboard. Only four dashboards have been
-fitted with a single screen. The majority of dashboards use a color
+fitted with a single screen. The majority of dashboards use a colour
 theme on the whole dashboard. That is, dashboards have been applied with
-different colors for different types of data (i.e., one specific color
-for confirmed cases, another color for deaths, etc.) across the whole
+different colours for different types of data (i.e., one specific colour
+for confirmed cases, another colour for deaths, etc.) across the whole
 dashboard. The data set and related links are available on some
@@
 using the ggplot2 (\textbf{ggplot2?}) and plotly (Sievert 2020) packages
-in R. We used color-blind friendly color palettes for the graphics. A
+in R. We used colour-blind friendly colour palettes for the graphics. A
 diverging colour palette was used to represent qualitative data, and to
@@
 Distribution of COVID-19 patients by districts & Tree map, Choropleth
-maps, Dorling Cartogram, heatmaps \hfill\break \\
+maps, Dorling Cartogram, heat maps \hfill\break \\
 Total vaccination by first dose and second dose & Time series plot
@@
 \hfill\break \\
-Comparison of cases with in Sri Lanka with Top 10 countries &
-Cummulative cases by time, Log of cummulative cases by time, stacked bar
-chart \hfill\break \\
+Comparison of cases with in Sri Lanka with Top 10 countries & Cumulative
+cases by time, Log of cumulative cases by time, stacked bar chart
+\hfill\break \\
 Spread of COVID-19 around the world & Choropleth maps \hfill\break \\
@@
 dashboard. To effectively distribute the vaccine and to support
-situational awareness and inform policymakers' decision-making, it is
+situational awareness and inform policy-makers' decision-making, it is
 important to know the district-wise spread of COVID-19 cases. ~We have
@@
 dissimilarities across districts. The associated results are shown in
-Figure 2. Here, two heatmaps are used to show the global variations
+Figure 2. Here, two heat maps are used to show the global variations
 (Figure 2: A) and local variations (Figure 2: B) in the time series
@@
 COVID-19 confirmed cases. This is useful to get an idea of the
-differences in absolute values. Figure 2-B cell colors represent the
+differences in absolute values. Figure 2-B cell colours represent the
 normalized values created by applying the min-max transformation. A
-Min-max transformation is applied to each district's time series by
+min-max transformation is applied to each district's time series by
 using the corresponding district's minimum and maximum value. This helps
@@
 24, and 29, 2021. Figure 3B is useful for identifying these local
-outlying behaviors. As shown in Figure 4, we also use a Choropleth map
+outlying behaviours. As shown in Figure 4, we also use a Choropleth map
 and Dorling cartogram to visualize the spatial distribution of COVID-19
@@
 
-The source code to reprouduce the results are available in a public
-GitHub repository and can be access online via
-\url{https://github.com/thiyangt/covid19srilanka}
-
 \hypertarget{results}{%
@@
 The ``Sri Lanka COVID-19 Dashboard'' provides an overview of the
-COVID-19 pandemic and administration of vaacine information in Sri
+COVID-19 pandemic and administration of vaccine information in Sri
 Lanka. This dashboard has eight panels as listed in Table 6.
@@
 WHO)\hfill\break & Figure 8 \\
-Global Comparison & Compare the total confirmed \& deaths in Sri lanka
+Global Comparison & Compare the total confirmed \& deaths in Sri Lanka
 with top 10 countries in Global \& Asia. There are two
-tabs.\hfill\break * Comparison of the sri Lanka with Top 10 Countries
+tabs.\hfill\break * Comparison of the Sri Lanka with Top 10 Countries
 Reporting the Most COVID-19 Cases in the World - Compare the total
 confirmed \& deaths in Sri Lanka with top 10 countries in the world \&
-compare the case fatality ratios\hfill\break * Comparison of the sri
+compare the case fatality ratios\hfill\break * Comparison of the Sri
 Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the
@@
 
-We used a color-blind-friendly theme when creating our dashboard. The
+We used a colour-blind-friendly theme when creating our dashboard. The
 dashboard includes both static and interactive charts that can be used

### 2022-05-15 manuscript_covid19dashboard.Rmd
Updated

@@
 title: "Interactive Dashboard to Monitor the COVID-19 Outbreak and Vaccine Administration"
+header-includes:
+   - \usepackage{url}
+   - \usepackage{hyperref}
 output: 
@@
 
-Bar charts and line charts are the most frequently used tools for the visualization of total cases, daily cases, and weekly cases.
-and comparisons with respect to time. Some dashboards contained doughnut-shaped pie charts to summarize
-the total figures. In almost every dashboard, value boxes have been used to represent total figures.
-Some dashboards contained interactive maps and data tables to visualize the distribution of cases by country,
-province, region, or state. All dashboards are updated daily in real time. Gender, age groups, and ethnicity can be identified as common breakdowns. Most data sets and related links are available.
+Bar charts and line charts are the most frequently used tools for the visualization of total cases, daily cases, and weekly cases, and comparisons with respect to time.
+Some dashboards contained doughnut-shaped pie charts to summarize the total figures.
+In almost every dashboard, value boxes have been used to represent total figures.
+Some dashboards contained interactive maps and data tables to visualize the distribution of cases by country, province, region, or state.
+All dashboards are updated daily in real time.
+Gender, age groups, and ethnicity can be identified as common breakdowns.
+Most data sets and related links are available.
 
 We used a colour-blind-friendly theme when creating our dashboard.
-The dashboard includes both static and interactive charts that can be used for explanation, exploration, and display.
+The dashboard includes both static and interactive charts that can be used for explanatory (telling you something), exploratory (give user the chance to interact with the plots), and exhibitory (showing data visually) purposes.
 The data speaks to us, but it is not always easy to understand in what language they are doing it.
@@
 We do not have district-wise vaccination details.
-We do not have vaccination information per district.
-We'll save this for later.
-Our dashboard is completely replicable.
+This is a potential future direction to think about.
+Our dashboard is completely reproducible.
 The source code for reproducing the results is accessible in a public GitHub repository at https://github.com/thiyangt/covid19srilanka.

### 2022-05-15 manuscript_covid19dashboard.tex
Updated

@@
 \setcounter{secnumdepth}{5}
+\usepackage{url}
+\usepackage{hyperref}
 \ifluatex
@@
 Bar charts and line charts are the most frequently used tools for the
-visualization of total cases, daily cases, and weekly cases. and
+visualization of total cases, daily cases, and weekly cases, and
 comparisons with respect to time. Some dashboards contained
@@
 dashboard includes both static and interactive charts that can be used
-for explanation, exploration, and display. The data speaks to us, but it
-is not always easy to understand in what language they are doing it.
-Hence, we examine the same data under different angles using different
-types of plots. We do not have district-wise vaccination details. We do
-not have vaccination information per district. We'll save this for
-later. Our dashboard is completely replicable. The source code for
-reproducing the results is accessible in a public GitHub repository at
-\url{https://github.com/thiyangt/covid19srilanka}.
+for explanatory (telling you something), exploratory (give user the
+chance to interact with the plots), and exhibitory (showing data
+visually) purposes. The data speaks to us, but it is not always easy to
+understand in what language they are doing it. Hence, we examine the
+same data under different angles using different types of plots. We do
+not have district-wise vaccination details. This is a potential future
+direction to think about. Our dashboard is completely reproducible. The
+source code for reproducing the results is accessible in a public GitHub
+repository at \url{https://github.com/thiyangt/covid19srilanka}.
 
